### 2016-04-08 Paper/Infovis Paper.tex

\documentclass{sigchi}
% Use this command to override the default ACM copyright statement
% (e.g. for preprints).  Consult the conference website for the
% camera-ready copyright statement.

%% HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
%% Please note you need to make sure the copy for your specific
%% license is used here!
% \toappear{
% Permission to make digital or hard copies of all or part of this work
% for personal or classroom use is granted without fee provided that
% copies are not made or distributed for profit or commercial advantage
% and that copies bear this notice and the full citation on the first
% page. Copyrights for components of this work owned by others than ACM
% must be honored. Abstracting with credit is permitted. To copy
% otherwise, or republish, to post on servers or to redistribute to
% lists, requires prior specific permission and/or a fee. Request
% permissions from \href{mailto:Permissions@acm.org}{Permissions@acm.org}. \\
% \emph{CHI '16},  May 07--12, 2016, San Jose, CA, USA \\
% ACM xxx-x-xxxx-xxxx-x/xx/xx\ldots \$15.00 \\
% DOI: \url{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
% }

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy
% \pagenumbering{arabic}

% Load basic packages
\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning
% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
\usepackage{ccicons}          % Cite your images correctly!
% \usepackage[utf8]{inputenc} % for a UTF8 editor only

% If you want to use todo notes, marginpars etc. during creation of
% your draft document, you have to enable the "chi_draft" option for
% the document class. To do this, change the very first line to:
% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
% by using the "\todo{...}"  command. Make sure to disable the draft
% option again before submitting your final document.
\usepackage{todonotes}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Visualizing NBA teams across multiple seasons}
\def\plainauthor[#1]{#1}
\def\plainauthors[#1]#2#3{#1, #2, #3}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; required.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
  pdfauthor={\plainauthors[Yannick Boesmans]{Yannick Laevaert}{Yves Langeraert}},
%  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  pdfdisplaydoctitle=true, % For Accessibility
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{\plaintitle}

\numberofauthors{3}
\author{%
  \alignauthor{\plainauthor[Yannick Boesmans]\\
    \affaddr{KU Leuven}\\
    \email{yannick.boesmans}\\
    \email{@student.kuleuven.be}}\\
  \alignauthor{\plainauthor[Yannick Laevaert]\\
    \affaddr{KU Leuven}\\
    \email{yannick.laevaert}\\
    \email{@student.kuleuven.be}}\\
  \alignauthor{\plainauthor[Yves Langeraert]\\
    \affaddr{KU Leuven}\\
    \email{yves.langeraert}\\
    \email{@student.kuleuven.be}}\\
}

\maketitle

\begin{abstract}
  UPDATED---\today. This sample paper describes the formatting
  requirements for SIGCHI conference proceedings, and offers
  recommendations on writing for the worldwide SIGCHI
  readership. Please review this document even if you have submitted
  to SIGCHI conferences before, as some format details have changed
  relative to previous years. Abstracts should be about 150 words and
  are required.
\end{abstract}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section{Introduction}

This format is to be used for submissions that are published in the
conference proceedings. We wish to give this volume a consistent,
high-quality appearance. We therefore ask that authors follow some
simple guidelines. You should format your paper exactly like this
document. The easiest way to do this is to replace the content with
your own material.  This document describes how to prepare your
submissions using \LaTeX.

\section{Page Size and Columns}
On each page your material should fit within a rectangle of 7 $\times$
9.15 inches (18 $\times$ 23.2 cm), centered on a US Letter page (8.5
$\times$ 11 inches), beginning 0.85 inches (1.9 cm) from the top of
the page, with a 0.3 inches (0.85 cm) space between two 3.35 inches
(8.4 cm) columns. Right margins should be justified, not
ragged. Please be sure your document and PDF are US letter and not A4.

\section{Typeset Text}
The styles contained in this document have been modified from the
default styles to reflect ACM formatting conventions. For example,
content paragraphs like this one are formatted using the Normal style.

\LaTeX\ sometimes will create overfull lines that extend into columns.
To attempt to combat this, the \texttt{.cls} file has a command,
\texttt{{\textbackslash}sloppy}, that essentially asks \LaTeX\ to
prefer underfull lines with extra whitespace.  For more details on
this, and info on how to control it more finely, check out
{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.

\subsection{Title and Authors}

Your paper's title, authors and affiliations should run across the
full width of the page in a single column 17.8 cm (7 in.) wide.  The
title should be in Helvetica or Arial 18-point bold.  Authors' names
should be in Times New Roman or Times Roman 12-point bold, and
affiliations in 12-point regular.

See \texttt{{\textbackslash}author} section of this template for
instructions on how to format the authors. For more than three
authors, you may have to place some address information in a footnote,
or in a named section at the end of your paper. Names may optionally
be placed in a single centered row instead of at the top of each
column. Leave one 10-point line of white space below the last line of
affiliations.

\subsection{Abstract and Keywords}

Every submission should begin with an abstract of about 150 words,
followed by a set of Author Keywords and ACM Classification
Keywords. The abstract and keywords should be placed in the left
column of the first page under the left half of the title. The
abstract should be a concise statement of the problem, approach, and
conclusions of the work described. It should clearly state the paper's
contribution to the field of HCI\@.

\subsection{Normal or Body Text}

Please use a 10-point Times New Roman or Times Roman font or, if this
is unavailable, another proportional font with serifs, as close as
possible in appearance to Times Roman 10-point. Other than Helvetica
or Arial headings, please use sans-serif or non-proportional fonts
only for special purposes, such as source code text.

\subsection{First Page Copyright Notice}
This template include a sample ACM copyright notice at the bottom of
page 1, column 1.  Upon acceptance, you will be provided with the
appropriate copyright statement and unique DOI string for publication.
Accepted papers will be distributed in the conference
publications. They will also be placed in the ACM Digital Library,
where they will remain accessible to thousands of researchers and
practitioners worldwide. See
\url{http://acm.org/publications/policies/copyright_policy} for the
ACM's copyright and permissions policy.

\subsection{Subsequent Pages}

On pages beyond the first, start at the top of the page and continue
in double-column format.  The two columns on the last page should be
of equal length.

\begin{figure}
\centering
  \includegraphics[width=0.9\columnwidth]{figures/sigchi-logo}
  \caption{Insert a caption below each figure. Do not alter the
    Caption style.  One-line captions should be centered; multi-line
    should be justified. }~\label{fig:figure1}
\end{figure}

\subsection{References and Citations}

Use a numbered list of references at the end of the article, ordered
alphabetically by last name of first author, and referenced by numbers
in
brackets~\cite{acm_categories,ethics,Klemmer:2002:WSC:503376.503378}.
Your references should be published materials accessible to the
public. Internal technical reports may be cited only if they are
easily accessible (i.e., you provide the address for obtaining the
report within your citation) and may be obtained by any reader for a
nominal fee. Proprietary information may not be cited. Private
communications should be acknowledged in the main text, not referenced
(e.g., ``[Borriello, personal communication]'').

References should be in ACM citation format:
\url{http://acm.org/publications/submissions/latex_style}. This
includes citations to internet
resources~\cite{acm_categories,cavender:writing,CHINOSAUR:venue,psy:gangnam}
according to ACM format, although it is often appropriate to include
URLs directly in the text, as above.

% Use a numbered list of references at the end of the article, ordered
% alphabetically by first author, and referenced by numbers in
% brackets~\cite{ethics, Klemmer:2002:WSC:503376.503378,
%   Mather:2000:MUT, Zellweger:2001:FAO:504216.504224}. For papers from
% conference proceedings, include the title of the paper and an
% abbreviated name of the conference (e.g., for Interact 2003
% proceedings, use \textit{Proc. Interact 2003}). Do not include the
% location of the conference or the exact date; do include the page
% numbers if available. See the examples of citations at the end of this
% document. Within this template file, use the \texttt{References} style
% for the text of your citation.

% Your references should be published materials accessible to the
% public.  Internal technical reports may be cited only if they are
% easily accessible (i.e., you provide the address for obtaining the
% report within your citation) and may be obtained by any reader for a
% nominal fee.  Proprietary information may not be cited. Private
% communications should be acknowledged in the main text, not referenced
% (e.g., ``[Robertson, personal communication]'').

\begin{table}
  \centering
  \begin{tabular}{l r r r}
    % \toprule
    & & \multicolumn{2}{c}{\small{\textbf{Test Conditions}}} \\
    \cmidrule(r){3-4}
    {\small\textit{Name}}
    & {\small \textit{First}}
      & {\small \textit{Second}}
    & {\small \textit{Final}} \\
    \midrule
    Marsden & 223.0 & 44 & 432,321 \\
    Nass & 22.2 & 16 & 234,333 \\
    Borriello & 22.9 & 11 & 93,123 \\
    Karat & 34.9 & 2200 & 103,322 \\
    % \bottomrule
  \end{tabular}
  \caption{Table captions should be placed below the table. We
    recommend table lines be 1 point, 25\% black. Minimize use of
    table grid lines.}~\label{tab:table1}
\end{table}

\section{Sections}

The heading of a section should be in Helvetica or Arial 9-point bold,
all in capitals. Sections should \textit{not} be numbered.

\subsection{Subsections}

Headings of subsections should be in Helvetica or Arial 9-point bold
with initial letters capitalized.  For sub-sections and
sub-subsections, a word like \emph{the} or \emph{of} is not
capitalized unless it is the first word of the heading.

\subsubsection{Sub-subsections}

Headings for sub-subsections should be in Helvetica or Arial 9-point
italic with initial letters capitalized.  Standard
\texttt{{\textbackslash}section}, \texttt{{\textbackslash}subsection},
and \texttt{{\textbackslash}subsubsection} commands will work fine in
this template.

\section{Figures/Captions}

Place figures and tables at the top or bottom of the appropriate
column or columns, on the same page as the relevant text (see
Figure~\ref{fig:figure1}). A figure or table may extend across both
columns to a maximum width of 17.78 cm (7 in.).

\begin{figure*}
  \centering
  \includegraphics[width=1.75\columnwidth]{figures/map}
  \caption{In this image, the map maximizes use of space. You can make
    figures as wide as you need, up to a maximum of the full width of
    both columns. Note that \LaTeX\ tends to render large figures on a
    dedicated page. Image: \ccbynd~ayman on
    Flickr.}~\label{fig:figure2}
\end{figure*}

Captions should be Times New Roman or Times Roman 9-point bold.  They
should be numbered (e.g., ``Table~\ref{tab:table1}'' or
``Figure~\ref{fig:figure1}''), centered and placed beneath the figure
or table.  Please note that the words ``Figure'' and ``Table'' should
be spelled out (e.g., ``Figure'' rather than ``Fig.'') wherever they
occur. Figures, like Figure~\ref{fig:figure2}, may span columns and
all figures should also include alt text for improved accessibility.
Papers and notes may use color figures, which are included in the page
limit; the figures must be usable when printed in black-and-white in
the proceedings.

The paper may be accompanied by a short video figure up to five
minutes in length. However, the paper should stand on its own without
the video figure, as the video may not be available to everyone who
reads the paper.

\subsection{Inserting Images}
When possible, include a vector formatted graphic (i.e. PDF or EPS).
When including bitmaps,  use an image editing tool to resize the image
at the appropriate printing resolution (usually 300 dpi).

\section{Quotations}
Quotations may be italicized when \textit{``placed inline''} (Anab,
23F).

\begin{quote}
Longer quotes, when placed in their own paragraph, need not be
italicized or in quotation marks when indented (Ramon, 39M).  
\end{quote}

\section{Language, Style, and Content}

The written and spoken language of SIGCHI is English. Spelling and
punctuation may use any dialect of English (e.g., British, Canadian,
US, etc.) provided this is done consis- tently. Hyphenation is
optional. To ensure suitability for an international audience, please
pay attention to the following:

\begin{itemize}
\item Write in a straightforward style.
\item Try to avoid long or complex sentence structures.
\item Briefly define or explain all technical terms that may be
  unfamiliar to readers.
\item Explain all acronyms the first time they are used in your
  text---e.g., ``Digital Signal Processing (DSP)''.
\item Explain local references (e.g., not everyone knows all city
  names in a particular country).
\item Explain ``insider'' comments. Ensure that your whole audience
  understands any reference whose meaning you do not describe (e.g.,
  do not assume that everyone has used a Macintosh or a particular
  application).
\item Explain colloquial language and puns. Understanding phrases like
  ``red herring'' may require a local knowledge of English.  Humor and
  irony are difficult to translate.
\item Use unambiguous forms for culturally localized concepts, such as
  times, dates, currencies, and numbers (e.g., ``1--5--97'' or
  ``5/1/97'' may mean 5 January or 1 May, and ``seven o'clock'' may
  mean 7:00 am or 19:00). For currencies, indicate equivalences:
  ``Participants were paid {\fontfamily{txr}\selectfont \textwon}
  25,000, or roughly US \$22.''
\item Be careful with the use of gender-specific pronouns (he, she)
  and other gendered words (chairman, manpower, man-months). Use
  inclusive language that is gender-neutral (e.g., she or he, they,
  s/he, chair, staff, staff-hours, person-years). See the
  \textit{Guidelines for Bias-Free Writing} for further advice and
  examples regarding gender and other personal
  attributes~\cite{Schwartz:1995:GBF}. Be particularly aware of
  considerations around writing about people with disabilities.
\item If possible, use the full (extended) alphabetic character set
  for names of persons, institutions, and places (e.g.,
  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Nguy{\~{\^{e}}}n,
  Universit{\"a}t, Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).
  These characters are already included in most versions and variants
  of Times, Helvetica, and Arial fonts.
\end{itemize}

\section{Accessibility}
The Executive Council of SIGCHI has committed to making SIGCHI
conferences more inclusive for researchers, practitioners, and
educators with disabilities. As a part of this goal, the all authors
are asked to work on improving the accessibility of their
submissions. Specifically, we encourage authors to carry out the
following five steps:
\begin{enumerate}
\item Add alternative text to all figures
\item Mark table headings
\item Add tags to the PDF
\item Verify the default language
\item Set the tab order to ``Use Document Structure''
\end{enumerate}
For more information and links to instructions and resources, please
see: \url{http://chi2016.acm.org/accessibility}.  The
\texttt{{\textbackslash}hyperref} package allows you to create well tagged PDF files,
please see the preamble of this template for an example.

\section{Page Numbering, Headers and Footers}
Your final submission should not contain footer or header information
at the top or bottom of each page. Specifically, your final submission
should not include page numbers. Initial submissions may include page
numbers, but these must be removed for camera-ready. Page numbers will
be added to the PDF when the proceedings are assembled.

\section{Producing and Testing PDF Files}

We recommend that you produce a PDF version of your submission well
before the final deadline.  Your PDF file must be ACM DL
Compliant. The requirements for an ACM Compliant PDF are available at:
{\url{http://www.sheridanprinting.com/typedept/ACM-distilling-settings.htm}}.

Test your PDF file by viewing or printing it with the same software we
will use when we receive it, Adobe Acrobat Reader Version 10. This is
widely available at no cost. Note that most
reviewers will use a North American/European version of Acrobat
reader, so please check your PDF accordingly.

When creating your PDF from Word, ensure that you generate a tagged
PDF from improved accessibility. This can be done by using the Adobe
PDF add-in, also called PDFMaker. Select Acrobat | Preferences from
the ribbon and ensure that ``Enable Accessibility and Reflow with
tagged Adobe PDF'' is selected. You can then generate a tagged PDF by
selecting ``Create PDF'' from the Acrobat ribbon.

\section{Conclusion}

It is important that you write for the SIGCHI audience. Please read
previous years' proceedings to understand the writing style and
conventions that successful authors have used. It is particularly
important that you state clearly what you have done, not merely what
you plan to do, and explain how your work is different from previously
published work, i.e., the unique contribution that your work makes to
the field. Please consider what the reader will learn from your
submission, and how they will find your work useful. If you write with
these questions in mind, your work is more likely to be successful,
both in being accepted into the conference, and in influencing the
work of our field.

\section{Acknowledgments}

Sample text: We thank all the volunteers, and all publications support
and staff, who wrote and provided helpful comments on previous
versions of this document. Authors 1, 2, and 3 gratefully acknowledge
the grant from NSF (\#1234--2012--ABC). \textit{This whole paragraph is
  just an example.}

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance{}

\section{References Format}
Your references should be published materials accessible to the
public. Internal technical reports may be cited only if they are
easily accessible and may be obtained by any reader for a nominal
fee. Proprietary information may not be cited. Private communications
should be acknowledged in the main text, not referenced (e.g.,
[Golovchinsky, personal communication]). References must be the same
font size as other body text. References should be in alphabetical
order by last name of first author. Use a numbered list of references
at the end of the article, ordered alphabetically by last name of
first author, and referenced by numbers in brackets. For papers from
conference proceedings, include the title of the paper and the name of
the conference. Do not include the location of the conference or the
exact date; do include the page numbers if available.

References should be in ACM citation format:
\url{http://www.acm.org/publications/submissions/latex_style}.  This
includes citations to Internet
resources~\cite{CHINOSAUR:venue,cavender:writing,psy:gangnam}
according to ACM format, although it is often appropriate to include
URLs directly in the text, as above. Example reference formatting for
individual journal articles~\cite{ethics}, articles in conference
proceedings~\cite{Klemmer:2002:WSC:503376.503378},
books~\cite{Schwartz:1995:GBF}, theses~\cite{sutherland:sketchpad},
book chapters~\cite{winner:politics}, an entire journal
issue~\cite{kaye:puc},
websites~\cite{acm_categories,cavender:writing},
tweets~\cite{CHINOSAUR:venue}, patents~\cite{heilig:sensorama}, 
games~\cite{supermetroid:snes}, and
online videos~\cite{psy:gangnam} is given here.  See the examples of
citations at the end of this document and in the accompanying
\texttt{BibTeX} document. This formatting is a edited version of the
format automatically generated by the ACM Digital Library
(\url{http://dl.acm.org}) as ``ACM Ref.'' DOI and/or URL links are
optional but encouraged as are full first names. Note that the
Hyperlink style used throughout this document uses blue links;
however, URLs in the references section may optionally appear in
black.

% BALANCE COLUMNS
\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{sample}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-04-11 Paper/Infovis Paper.tex

@@
 % }
+\toappear{}
 
@@
 \begin{abstract}
-  UPDATED---\today. This sample paper describes the formatting
-  requirements for SIGCHI conference proceedings, and offers
-  recommendations on writing for the worldwide SIGCHI
-  readership. Please review this document even if you have submitted
-  to SIGCHI conferences before, as some format details have changed
-  relative to previous years. Abstracts should be about 150 words and
-  are required.
+    This paper details the conceptualization and implementation of an
+    interactive visualization of data about NBA seasons including teams, game
+    results and player statistics. The d3.js framework was used to develop this
+    visualization.
+    %TODO add results
 \end{abstract}
 
+%TODO keywords
 \category{H.5.m.}{Information Interfaces and Presentation
@@
 
-This format is to be used for submissions that are published in the
-conference proceedings. We wish to give this volume a consistent,
-high-quality appearance. We therefore ask that authors follow some
-simple guidelines. You should format your paper exactly like this
-document. The easiest way to do this is to replace the content with
-your own material.  This document describes how to prepare your
-submissions using \LaTeX.
+This paper describes the visualization of NBA data made by the ``The Tufters''
+team for the course ``Information Visualization''. In section~\ref{sec:goal}
+we describe the goal of the visualization and the target audience. In
+section~\ref{sec:data} we describe the data used, including its origins,
+advantages and limitations. In section~\ref{sec:literature} we give an overview
+of related literature and web resources, including related visualizations.
+Related visualizations include both visualizations of NBA or other sports data,
+as well as visualizations tackling a problem we encountered during the
+development of our visualization. In section~\ref{sec:visualization} we describe
+the visualization itself. We give an overview of the different stages of
+development of the visualization, as well as the major design decisions made
+during the development process. Section~\ref{sec:discussion} discusses potential
+improvements of the final visualization and lessons learned from the project. We
+conclude in section~\ref{sec:conclusion} by giving a short overview.
 
-\section{Page Size and Columns}
-On each page your material should fit within a rectangle of 7 $\times$
-9.15 inches (18 $\times$ 23.2 cm), centered on a US Letter page (8.5
-$\times$ 11 inches), beginning 0.85 inches (1.9 cm) from the top of
-the page, with a 0.3 inches (0.85 cm) space between two 3.35 inches
-(8.4 cm) columns. Right margins should be justified, not
-ragged. Please be sure your document and PDF are US letter and not A4.
+\section{Goal and Target Audience}\ref{sec:goal}
+The visualization's goal is allowing exploration of NBA data by lay persons.
+More specifically, the visualization does not try to offer premade explanations
+for phenomenons visible in the NBA data. By providing easy and intuitive access
+to the data, users can draw their own conclusions. 
+Secondly, the visualization's target audience are lay persons. This means people
+not professionally active in the field of basketball. Specifically, fans of NBA
+are the core of our target audience.
 
-\section{Typeset Text}
-The styles contained in this document have been modified from the
-default styles to reflect ACM formatting conventions. For example,
-content paragraphs like this one are formatted using the Normal style.
+\section{Data}\ref{sec:data}
+The data visualized is a subset of the data available on basketball statistics
+site basketball-reference\cite{basketball-reference}. A wide range of data is
+available on this site. In our visualization we only use data from 1971 onwards.
+The data we use includes league standings and playoff rankings for each team,
+the team's roster and individual player statistics for each year, including the
+PER~(Player Efficiency Rating).
 
-\LaTeX\ sometimes will create overfull lines that extend into columns.
-To attempt to combat this, the \texttt{.cls} file has a command,
-\texttt{{\textbackslash}sloppy}, that essentially asks \LaTeX\ to
-prefer underfull lines with extra whitespace.  For more details on
-this, and info on how to control it more finely, check out
-{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.
-
-\subsection{Title and Authors}
-
-Your paper's title, authors and affiliations should run across the
-full width of the page in a single column 17.8 cm (7 in.) wide.  The
-title should be in Helvetica or Arial 18-point bold.  Authors' names
-should be in Times New Roman or Times Roman 12-point bold, and
-affiliations in 12-point regular.
-
-See \texttt{{\textbackslash}author} section of this template for
-instructions on how to format the authors. For more than three
-authors, you may have to place some address information in a footnote,
-or in a named section at the end of your paper. Names may optionally
-be placed in a single centered row instead of at the top of each
-column. Leave one 10-point line of white space below the last line of
-affiliations.
-
-\subsection{Abstract and Keywords}
-
-Every submission should begin with an abstract of about 150 words,
-followed by a set of Author Keywords and ACM Classification
-Keywords. The abstract and keywords should be placed in the left
-column of the first page under the left half of the title. The
-abstract should be a concise statement of the problem, approach, and
-conclusions of the work described. It should clearly state the paper's
-contribution to the field of HCI\@.
-
-\subsection{Normal or Body Text}
-
-Please use a 10-point Times New Roman or Times Roman font or, if this
-is unavailable, another proportional font with serifs, as close as
-possible in appearance to Times Roman 10-point. Other than Helvetica
-or Arial headings, please use sans-serif or non-proportional fonts
-only for special purposes, such as source code text.
-
-\subsection{First Page Copyright Notice}
-This template include a sample ACM copyright notice at the bottom of
-page 1, column 1.  Upon acceptance, you will be provided with the
-appropriate copyright statement and unique DOI string for publication.
-Accepted papers will be distributed in the conference
-publications. They will also be placed in the ACM Digital Library,
-where they will remain accessible to thousands of researchers and
-practitioners worldwide. See
-\url{http://acm.org/publications/policies/copyright_policy} for the
-ACM's copyright and permissions policy.
-
-\subsection{Subsequent Pages}
-
-On pages beyond the first, start at the top of the page and continue
-in double-column format.  The two columns on the last page should be
-of equal length.
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.9\columnwidth]{figures/sigchi-logo}
-  \caption{Insert a caption below each figure. Do not alter the
-    Caption style.  One-line captions should be centered; multi-line
-    should be justified. }~\label{fig:figure1}
-\end{figure}
-
-\subsection{References and Citations}
-
-Use a numbered list of references at the end of the article, ordered
-alphabetically by last name of first author, and referenced by numbers
-in
-brackets~\cite{acm_categories,ethics,Klemmer:2002:WSC:503376.503378}.
-Your references should be published materials accessible to the
-public. Internal technical reports may be cited only if they are
-easily accessible (i.e., you provide the address for obtaining the
-report within your citation) and may be obtained by any reader for a
-nominal fee. Proprietary information may not be cited. Private
-communications should be acknowledged in the main text, not referenced
-(e.g., ``[Borriello, personal communication]'').
-
-References should be in ACM citation format:
-\url{http://acm.org/publications/submissions/latex_style}. This
-includes citations to internet
-resources~\cite{acm_categories,cavender:writing,CHINOSAUR:venue,psy:gangnam}
-according to ACM format, although it is often appropriate to include
-URLs directly in the text, as above.
-
-% Use a numbered list of references at the end of the article, ordered
-% alphabetically by first author, and referenced by numbers in
-% brackets~\cite{ethics, Klemmer:2002:WSC:503376.503378,
-%   Mather:2000:MUT, Zellweger:2001:FAO:504216.504224}. For papers from
-% conference proceedings, include the title of the paper and an
-% abbreviated name of the conference (e.g., for Interact 2003
-% proceedings, use \textit{Proc. Interact 2003}). Do not include the
-% location of the conference or the exact date; do include the page
-% numbers if available. See the examples of citations at the end of this
-% document. Within this template file, use the \texttt{References} style
-% for the text of your citation.
-
-% Your references should be published materials accessible to the
-% public.  Internal technical reports may be cited only if they are
-% easily accessible (i.e., you provide the address for obtaining the
-% report within your citation) and may be obtained by any reader for a
-% nominal fee.  Proprietary information may not be cited. Private
-% communications should be acknowledged in the main text, not referenced
-% (e.g., ``[Robertson, personal communication]'').
-
-\begin{table}
-  \centering
-  \begin{tabular}{l r r r}
-    % \toprule
-    & & \multicolumn{2}{c}{\small{\textbf{Test Conditions}}} \\
-    \cmidrule(r){3-4}
-    {\small\textit{Name}}
-    & {\small \textit{First}}
-      & {\small \textit{Second}}
-    & {\small \textit{Final}} \\
-    \midrule
-    Marsden & 223.0 & 44 & 432,321 \\
-    Nass & 22.2 & 16 & 234,333 \\
-    Borriello & 22.9 & 11 & 93,123 \\
-    Karat & 34.9 & 2200 & 103,322 \\
-    % \bottomrule
-  \end{tabular}
-  \caption{Table captions should be placed below the table. We
-    recommend table lines be 1 point, 25\% black. Minimize use of
-    table grid lines.}~\label{tab:table1}
-\end{table}
-
-\section{Sections}
-
-The heading of a section should be in Helvetica or Arial 9-point bold,
-all in capitals. Sections should \textit{not} be numbered.
-
-\subsection{Subsections}
-
-Headings of subsections should be in Helvetica or Arial 9-point bold
-with initial letters capitalized.  For sub-sections and
-sub-subsections, a word like \emph{the} or \emph{of} is not
-capitalized unless it is the first word of the heading.
-
-\subsubsection{Sub-subsections}
-
-Headings for sub-subsections should be in Helvetica or Arial 9-point
-italic with initial letters capitalized.  Standard
-\texttt{{\textbackslash}section}, \texttt{{\textbackslash}subsection},
-and \texttt{{\textbackslash}subsubsection} commands will work fine in
-this template.
-
-\section{Figures/Captions}
-
-Place figures and tables at the top or bottom of the appropriate
-column or columns, on the same page as the relevant text (see
-Figure~\ref{fig:figure1}). A figure or table may extend across both
-columns to a maximum width of 17.78 cm (7 in.).
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=1.75\columnwidth]{figures/map}
-  \caption{In this image, the map maximizes use of space. You can make
-    figures as wide as you need, up to a maximum of the full width of
-    both columns. Note that \LaTeX\ tends to render large figures on a
-    dedicated page. Image: \ccbynd~ayman on
-    Flickr.}~\label{fig:figure2}
-\end{figure*}
-
-Captions should be Times New Roman or Times Roman 9-point bold.  They
-should be numbered (e.g., ``Table~\ref{tab:table1}'' or
-``Figure~\ref{fig:figure1}''), centered and placed beneath the figure
-or table.  Please note that the words ``Figure'' and ``Table'' should
-be spelled out (e.g., ``Figure'' rather than ``Fig.'') wherever they
-occur. Figures, like Figure~\ref{fig:figure2}, may span columns and
-all figures should also include alt text for improved accessibility.
-Papers and notes may use color figures, which are included in the page
-limit; the figures must be usable when printed in black-and-white in
-the proceedings.
-
-The paper may be accompanied by a short video figure up to five
-minutes in length. However, the paper should stand on its own without
-the video figure, as the video may not be available to everyone who
-reads the paper.
-
-\subsection{Inserting Images}
-When possible, include a vector formatted graphic (i.e. PDF or EPS).
-When including bitmaps,  use an image editing tool to resize the image
-at the appropriate printing resolution (usually 300 dpi).
-
-\section{Quotations}
-Quotations may be italicized when \textit{``placed inline''} (Anab,
-23F).
-
-\begin{quote}
-Longer quotes, when placed in their own paragraph, need not be
-italicized or in quotation marks when indented (Ramon, 39M).  
-\end{quote}
-
-\section{Language, Style, and Content}
-
-The written and spoken language of SIGCHI is English. Spelling and
-punctuation may use any dialect of English (e.g., British, Canadian,
-US, etc.) provided this is done consis- tently. Hyphenation is
-optional. To ensure suitability for an international audience, please
-pay attention to the following:
-
-\begin{itemize}
-\item Write in a straightforward style.
-\item Try to avoid long or complex sentence structures.
-\item Briefly define or explain all technical terms that may be
-  unfamiliar to readers.
-\item Explain all acronyms the first time they are used in your
-  text---e.g., ``Digital Signal Processing (DSP)''.
-\item Explain local references (e.g., not everyone knows all city
-  names in a particular country).
-\item Explain ``insider'' comments. Ensure that your whole audience
-  understands any reference whose meaning you do not describe (e.g.,
-  do not assume that everyone has used a Macintosh or a particular
-  application).
-\item Explain colloquial language and puns. Understanding phrases like
-  ``red herring'' may require a local knowledge of English.  Humor and
-  irony are difficult to translate.
-\item Use unambiguous forms for culturally localized concepts, such as
-  times, dates, currencies, and numbers (e.g., ``1--5--97'' or
-  ``5/1/97'' may mean 5 January or 1 May, and ``seven o'clock'' may
-  mean 7:00 am or 19:00). For currencies, indicate equivalences:
-  ``Participants were paid {\fontfamily{txr}\selectfont \textwon}
-  25,000, or roughly US \$22.''
-\item Be careful with the use of gender-specific pronouns (he, she)
-  and other gendered words (chairman, manpower, man-months). Use
-  inclusive language that is gender-neutral (e.g., she or he, they,
-  s/he, chair, staff, staff-hours, person-years). See the
-  \textit{Guidelines for Bias-Free Writing} for further advice and
-  examples regarding gender and other personal
-  attributes~\cite{Schwartz:1995:GBF}. Be particularly aware of
-  considerations around writing about people with disabilities.
-\item If possible, use the full (extended) alphabetic character set
-  for names of persons, institutions, and places (e.g.,
-  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Nguy{\~{\^{e}}}n,
-  Universit{\"a}t, Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).
-  These characters are already included in most versions and variants
-  of Times, Helvetica, and Arial fonts.
-\end{itemize}
-
-\section{Accessibility}
-The Executive Council of SIGCHI has committed to making SIGCHI
-conferences more inclusive for researchers, practitioners, and
-educators with disabilities. As a part of this goal, the all authors
-are asked to work on improving the accessibility of their
-submissions. Specifically, we encourage authors to carry out the
-following five steps:
-\begin{enumerate}
-\item Add alternative text to all figures
-\item Mark table headings
-\item Add tags to the PDF
-\item Verify the default language
-\item Set the tab order to ``Use Document Structure''
-\end{enumerate}
-For more information and links to instructions and resources, please
-see: \url{http://chi2016.acm.org/accessibility}.  The
-\texttt{{\textbackslash}hyperref} package allows you to create well tagged PDF files,
-please see the preamble of this template for an example.
-
-\section{Page Numbering, Headers and Footers}
-Your final submission should not contain footer or header information
-at the top or bottom of each page. Specifically, your final submission
-should not include page numbers. Initial submissions may include page
-numbers, but these must be removed for camera-ready. Page numbers will
-be added to the PDF when the proceedings are assembled.
-
-\section{Producing and Testing PDF Files}
-
-We recommend that you produce a PDF version of your submission well
-before the final deadline.  Your PDF file must be ACM DL
-Compliant. The requirements for an ACM Compliant PDF are available at:
-{\url{http://www.sheridanprinting.com/typedept/ACM-distilling-settings.htm}}.
-
-Test your PDF file by viewing or printing it with the same software we
-will use when we receive it, Adobe Acrobat Reader Version 10. This is
-widely available at no cost. Note that most
-reviewers will use a North American/European version of Acrobat
-reader, so please check your PDF accordingly.
-
-When creating your PDF from Word, ensure that you generate a tagged
-PDF from improved accessibility. This can be done by using the Adobe
-PDF add-in, also called PDFMaker. Select Acrobat | Preferences from
-the ribbon and ensure that ``Enable Accessibility and Reflow with
-tagged Adobe PDF'' is selected. You can then generate a tagged PDF by
-selecting ``Create PDF'' from the Acrobat ribbon.
-
-\section{Conclusion}
-
-It is important that you write for the SIGCHI audience. Please read
-previous years' proceedings to understand the writing style and
-conventions that successful authors have used. It is particularly
-important that you state clearly what you have done, not merely what
-you plan to do, and explain how your work is different from previously
-published work, i.e., the unique contribution that your work makes to
-the field. Please consider what the reader will learn from your
-submission, and how they will find your work useful. If you write with
-these questions in mind, your work is more likely to be successful,
-both in being accepted into the conference, and in influencing the
-work of our field.
-
-\section{Acknowledgments}
-
-Sample text: We thank all the volunteers, and all publications support
-and staff, who wrote and provided helpful comments on previous
-versions of this document. Authors 1, 2, and 3 gratefully acknowledge
-the grant from NSF (\#1234--2012--ABC). \textit{This whole paragraph is
-  just an example.}
+\section{Related Work}\ref{sec:literature}
+\section{Visualization}\ref{sec:visualization}
+\section{Discussion}\ref{sec:discussion}
+\section{Conclusion}\ref{sec:conclusion}
 
@@
 
-\section{References Format}
-Your references should be published materials accessible to the
-public. Internal technical reports may be cited only if they are
-easily accessible and may be obtained by any reader for a nominal
-fee. Proprietary information may not be cited. Private communications
-should be acknowledged in the main text, not referenced (e.g.,
-[Golovchinsky, personal communication]). References must be the same
-font size as other body text. References should be in alphabetical
-order by last name of first author. Use a numbered list of references
-at the end of the article, ordered alphabetically by last name of
-first author, and referenced by numbers in brackets. For papers from
-conference proceedings, include the title of the paper and the name of
-the conference. Do not include the location of the conference or the
-exact date; do include the page numbers if available.
-
-References should be in ACM citation format:
-\url{http://www.acm.org/publications/submissions/latex_style}.  This
-includes citations to Internet
-resources~\cite{CHINOSAUR:venue,cavender:writing,psy:gangnam}
-according to ACM format, although it is often appropriate to include
-URLs directly in the text, as above. Example reference formatting for
-individual journal articles~\cite{ethics}, articles in conference
-proceedings~\cite{Klemmer:2002:WSC:503376.503378},
-books~\cite{Schwartz:1995:GBF}, theses~\cite{sutherland:sketchpad},
-book chapters~\cite{winner:politics}, an entire journal
-issue~\cite{kaye:puc},
-websites~\cite{acm_categories,cavender:writing},
-tweets~\cite{CHINOSAUR:venue}, patents~\cite{heilig:sensorama}, 
-games~\cite{supermetroid:snes}, and
-online videos~\cite{psy:gangnam} is given here.  See the examples of
-citations at the end of this document and in the accompanying
-\texttt{BibTeX} document. This formatting is a edited version of the
-format automatically generated by the ACM Digital Library
-(\url{http://dl.acm.org}) as ``ACM Ref.'' DOI and/or URL links are
-optional but encouraged as are full first names. Note that the
-Hyperlink style used throughout this document uses blue links;
-however, URLs in the references section may optionally appear in
-black.
-
-% BALANCE COLUMNS
-\balance{}
-
 % REFERENCES FORMAT

### 2016-04-11 bubbleview.js
Update bubbleview.js
Nieuwe functie met curved lines, selectie mogelijk die blijft overheen de tijd en aanzet naar nieuwe lijnenstructuur.
Ook al begin voor highlights (team geselecteerd is gekleurd, de rest fade wat uit)

@@ -1,8 +1,13 @@
-function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xpos, ypos, width, height){
+function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id){
+		console.log(dataInput);
+		
 		var playOffs = dataInput[0].playoffs;
 		var dataInput = dataInput[0].teams;
 		var east = dataInput.filter(function(data) { return data.region == "east"});
 		var west = dataInput.filter(function(data) { return data.region == "west"});
+		
+		var height = 600,
+			width = document.body.clientWidth*0.8;
 		
 		var yMaxData = d3.max(dataInput, function(data) {return parseInt(data[outlineVariable]);});
 		
@@ -12,7 +17,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			positionArray[i] = [0, dataInput.filter(function(data) {return data[outlineVariable] == i+1 }).length]
 		}
 		positionArray[yMaxData] = [0, dataInput.filter(function(data) {return data[outlineVariable] == null }).length]
-		
+
 		function rgbcolor(strokeVariable, region) {
 			if (region == 'west')
 			    return (d3.rgb(colorScale(strokeVariable),colorScale(strokeVariable), 255)); 
@@ -28,24 +33,70 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 		
 		function yPosition(outlineVariable) {
 			if (outlineVariable === undefined){
-				return (ypos + height-40)}
+				return (height-40)}
 			else	
-				return (ypos + 20+(parseInt(outlineVariable))*(height/(yMaxData+2)))
+				return (20+(parseInt(outlineVariable)*2-1)*(height/(yMaxData*2+2)))
 		}
 		
-		function xPosition(outlineVariable) {
+		function xPosition(outlineVariable, team) {
+			var lastMatch = playOffs.filter(function(data) {return data.loser == team});
+			try {
+				var winner = dataInput.filter(function(data) { return data.team == lastMatch[0].winner});
+				var lastMatchWinner = playOffs.filter(function(data) {return data.loser == winner[0].team});
+				var winnerofwinner = dataInput.filter(function(data) { return data.team == lastMatchWinner[0].winner})
+				var loser = dataInput.filter(function(data) { return data.team == lastMatch[0].loser});}
+				catch(err){
+					var winnerofwinner = null;
[…]

### 2016-04-19 bubbleview.js
Quite some stuff

@@ -211,8 +211,8 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 								})
       						.on('mouseout', function(data) {
       								tip.hide(data);
-      								d3.select(this)
-										.style('fill',rgbcolor(data[strokeVariable], data["region"]));
+                                    d3.select(this)
+                                        .style('fill',rgbcolor(data[strokeVariable], data["region"]));
       								if (visibleBoolean){
       									d3.selectAll(".arc").style("visibility", "visible");
       									console.log("#"+visibleClass.split(".").join("_"))

### 2016-04-19 scraper/combine.py

@@ -10,22 +10,10 @@ for yeardata in data:
     year = yeardata['year']
     for team in yeardata['teams']:
         teamname = team['team']
-        teamfile = teamfolder + str(year) + "_" + teamname.replace(" ", "_").replace("/", "-") + ".csv"
-        with open(teamfile, 'r') as csvfile:
-            reader = csv.reader(csvfile, delimiter=',')
-            keys = []
-            players = []
-            first = True;
-            for row in reader:
-                if first:
-                    keys = row
-                    first = False;
-                else:
-                    player = {}
-                    for i in range(0, len(row)):
-                        player[keys[i]] = row[i]
-                    players.append(player)
-            team['players'] = players
+        teamfile = teamfolder + str(year) + "_" + teamname.replace(" ", "_").replace("/", "-") + ".json"
+        with open(teamfile, 'r') as jsonfile:
+            f = json.load(jsonfile)
+            team['players'] = f
 
 with open('combined.json', 'w') as outjsonfile:
     json.dump(data, outjsonfile, indent=4)

### 2016-04-19 teamstats.js

@@ -2,10 +2,10 @@ function drawTs(data, team, year, div, x, y, width, height) {
 
     d3.select(".smallmultiple").remove();
 
-    var nPlots = 5;
+    var bases = ['srs', 'srs', 'srs'];
 
     var x = d3.scale.linear()
-        .range([0, width/nPlots])
+    .range([0, width/bases.length])
         .domain([1971,2015]);
 
     var srsScale = d3.scale.linear()
@@ -26,34 +26,36 @@ function drawTs(data, team, year, div, x, y, width, height) {
         }
     });
 
+    var svg = div.selectAll("svg")
+                    .data(bases)
+                    .enter().append("svg")
+                        .attr("width", width/bases.length)
+                        .attr("height", height)
+                        .attr("class", "smallmultiple");
 
-    var srs = div.append("svg")
-                 .attr("class", "smallmultiple")
-                 .attr("width", width/nPlots)
-                 .attr("height", height);
+    console.log(svg);
 
-    srs.append("path")
+    svg.append("path")
         .datum(srsData)
         .attr("class", "line")
         .attr("d", srsline);
 
-    srs.append("text")
-        .attr("x", (width / (2 * nPlots)))
+    svg.append("text")
+        .attr("x", (width / (2 * bases.length)))
         .attr("y", 20)
         .attr("text-anchor", "middle")
         .text("SRS");
 
-
     var highlightYear = year;
 
-    srs.on("mouseout", function() {
+    div.on("mouseout", function() {
         highlightYear = year;
         fixcyline();
         drawBubbles(miniBubbleSvg, 0, 0, miniw, minih, highlightYear, true);
         drawZoom(zoomSvg, 0, 0, width, height - 280, highlightYear, team);
     });
 
-    srs.on("mousemove", function() {
+    svg.on("mousemove", function() {
         highlightYear = Math.round(x.invert(d3.mouse(this)[0]));
         fixcyline();
         drawBubbles(miniBubbleSvg, 0, 0, miniw, minih, highlightYear, true);
@@ -79,15 +81,14 @@ function drawTs(data, team, year, div, x, y, width, height) {
 
     }
 
-    
-    srs.append("line")
+    svg.selectAll("svg").append("line")
         .attr("class", "curYearLine");
 
-    srs.append("text")
+    svg.append("text")
         .attr("class", "curYearVal")
         .attr("text-anchor", "middle");
 
-    srs.append("text")
+    svg.append("text")
         .attr("class", "curYearName")
         .attr("text-anchor", "middle");
         

### 2016-04-19 teamstats.js
Some more changes

@@ -1,30 +1,45 @@
 function drawTs(data, team, year, div, x, y, width, height) {
 
-    d3.select(".smallmultiple").remove();
-
-    var bases = ['srs', 'srs', 'srs'];
-
-    var x = d3.scale.linear()
-    .range([0, width/bases.length])
-        .domain([1971,2015]);
+    d3.selectAll(".smallmultiple").remove();
 
     var srsScale = d3.scale.linear()
         .range([height, 0])
         .domain([-10,10]);
 
-    var srsline = d3.svg.line()
-        .x(function(d) { return x(d.year); })
-        .y(function(d) { return srsScale(d.val); });
+    var rankScale = d3.scale.linear()
+        .range([height, 0])
+        .domain([16,1]);
 
-    console.log(data);
-    srsData = []
-    data.forEach(function (d) {
-        var theTeam = d.teams.filter(function (d) { return d.team == team; }); 
-        if (theTeam.length > 0) {
-            srsData.push({year : d.year,
-                val : theTeam[0].srs });
-        }
+    var poRankScale = d3.scale.linear()
+        .range([height, 0])
+        .domain([6,1]);
+
+    var bases = [{ name: 'SRS', dataSelector : function(d) { return d.team.srs; }, scaler: srsScale}
+            ,{ name: 'League Rank', dataSelector: function(d) {return d.team.leaguerank; }, scaler: rankScale}
+            ,{ name: 'Playoff Rank', dataSelector: function(d) {if (d.team.playoffrank != undefined) {return d.team.playoffrank;} else { return 6;} }, scaler: poRankScale}];
+
+    var x = d3.scale.linear()
+    .range([0, width/bases.length])
+        .domain([1984,2015]);
+
+    var xAxis = d3.svg.axis()
+        .scale(x)
+        .orient("top");
+
+    function makeLine(b) {
+        return d3.svg.line()
+                .x(function(d) { return x(d.year); })
+                .y(function(d) { return b.scaler(b.dataSelector(d)); });
+    }
+
+    filteredData = data.map(function (d) { 
+            var year = d.year; 
+            var thisTeam = d.teams.find(function (d) { return d.team === team; }); 
+            return {year: year, team: thisTeam}; 
+    }).filter(function(d) {
+        return d.team != undefined;
     });
+
 
     var svg = div.selectAll("svg")
                     .data(bases)
@@ -33,55 +48,30 @@ function drawTs(data, team, year, div, x, y, width, height) {
                         .attr("height", height)
[…]

### 2016-04-19 zoomview.js

@@ -47,7 +47,7 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
         CharlotteBobcats : {shirt: "000080", edge: "add8e6"}
     }
 
-    console.log(teamWanted, yearWanted, arrowVariable);
+    //console.log(teamWanted, yearWanted, arrowVariable);
     d3.select(".zoomchart").remove();
 
     var margin = {top: 20, right: 20, bottom: 20, left: 20};
@@ -204,7 +204,7 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
     var rad = scaleCircle(teamSRS, w, h, minMaxSRSAllYears[0], minMaxSRSAllYears[1]);
     var maxRad = scaleCircle(minMaxSRSAllYears[1], w, h, minMaxSRSAllYears[0], minMaxSRSAllYears[1]);
     var minRad = scaleCircle(minMaxSRSAllYears[0], w, h, minMaxSRSAllYears[0], minMaxSRSAllYears[1]);
-    console.log(minRad);
+    //console.log(minRad);
     chart.append('g')
         .attr('class', 'circleTeam')
         .append("circle")
@@ -721,6 +721,6 @@ function PERChangesPosition(players1, players2, position) {
     currTemp = parseFloat(currTemp) / currentPlayers.length;
     result = parseFloat(currTemp) - parseFloat(prevTemp);
 
-    console.log(result)
+    //console.log(result)
     return result;
 }

### 2016-04-19 bubbleview.js
All new data gathered, fixed some bugs

@@ -1,5 +1,5 @@
 function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
-		console.log(dataInput);
+		//console.log(dataInput);
 		
 		var playOffs = dataInput[0].playoffs;
 		var dataInput = dataInput[0].teams;
@@ -89,7 +89,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 		// Returns an event handler for fading a given chord group.
 		function fade(opacity) {
 		  return function(g, i) {
-		  	console.log(svg.selectAll(".circles"));
+		  	//console.log(svg.selectAll(".circles"));
 		    svg.selectAll(".circles")
 		        //.filter(function(d) { return d.source.index != i && d.target.index != i; })
 		      .transition()
@@ -215,8 +215,8 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
                                         .style('fill',rgbcolor(data[strokeVariable], data["region"]));
       								if (visibleBoolean){
       									d3.selectAll(".arc").style("visibility", "visible");
-      									console.log("#"+visibleClass.split(".").join("_"))
-      									console.log(d3.select("#"+visibleClass.split(".").join("_")));
+      									//console.log("#"+visibleClass.split(".").join("_"))
+      									//console.log(d3.select("#"+visibleClass.split(".").join("_")));
       									d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
 									}
 
@@ -266,7 +266,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 										.style('fill',rgbcolor(data[strokeVariable], data["region"]));
       								if (visibleBoolean){
       									d3.selectAll(".arc").style("visibility", "visible");
-      									console.log("."+visibleClass+"circles");
+      									//console.log("."+visibleClass+"circles");
       									
 									}
 									else d3.select("#"+visibleClass).style('fill','orange');

### 2016-04-19 scraper/combine.py

@@ -8,8 +8,11 @@ with open('data.json', 'r') as jsonfile:
 
 for yeardata in data:
     year = yeardata['year']
+    print ""
+    print "NEW YEAR " + str(year)
     for team in yeardata['teams']:
         teamname = team['team']
+        print "Doing team " + teamname
         teamfile = teamfolder + str(year) + "_" + teamname.replace(" ", "_").replace("/", "-") + ".json"
         with open(teamfile, 'r') as jsonfile:
             f = json.load(jsonfile)

### 2016-04-19 teamstats.js

@@ -2,24 +2,52 @@ function drawTs(data, team, year, div, x, y, width, height) {
 
     d3.selectAll(".smallmultiple").remove();
 
+    var nRows = 2;
+    var nColumns = 4;
+
+    var singleWidth = width / nColumns;
+    var singleHeight = height / nRows;
+
     var srsScale = d3.scale.linear()
-        .range([height, 0])
-        .domain([-10,10]);
+        .range([singleHeight, 0])
+        .domain([-15,15]);
 
     var rankScale = d3.scale.linear()
-        .range([height, 0])
+        .range([singleHeight, 0])
         .domain([16,1]);
 
     var poRankScale = d3.scale.linear()
-        .range([height, 0])
+        .range([singleHeight, 0])
         .domain([6,1]);
+
+    var ageScale = d3.scale.linear()
+        .range([singleHeight, 0])
+        .domain([20, 35]);
+
+    var audienceScale = d3.scale.linear()
+        .range([singleHeight, 0])
+        .domain([0, 1000000]);
+
+    var fgScale = d3.scale.linear()
+        .range([singleHeight, 0])
+        .domain([0, 1]);
+
+    var pointScale = d3.scale.linear()
+        .range([singleHeight, 0])
+        .domain([60, 140]);
 
     var bases = [{ name: 'SRS', dataSelector : function(d) { return d.team.srs; }, scaler: srsScale}
             ,{ name: 'League Rank', dataSelector: function(d) {return d.team.leaguerank; }, scaler: rankScale}
-            ,{ name: 'Playoff Rank', dataSelector: function(d) {if (d.team.playoffrank != undefined) {return d.team.playoffrank;} else { return 6;} }, scaler: poRankScale}];
+            ,{ name: 'Playoff Rank', dataSelector: function(d) {if (d.team.playoffrank != undefined) {return d.team.playoffrank;} else { return 6;} }, scaler: poRankScale}
+            ,{ name: 'Average Age', dataSelector: function(d) {return d.team.misc.Age;}, scaler: ageScale}
+            ,{ name: 'Audience', dataSelector: function(d) {return d.team.misc.Attendance;}, scaler: audienceScale}
+            ,{ name: 'Field goal %', dataSelector: function(d) {return d.team.info['FG%'];}, scaler: fgScale}
+            ,{ name: 'Points/Game', dataSelector: function(d) {return d.team.info['PTS/G'];}, scaler: pointScale}
[…]

### 2016-04-19 zoomview.js

@@ -540,8 +540,8 @@ function separatePlayers(previousPlayers, currentPlayers) {
 function totalPER(players) {
     var total = 0;
     for (var i = 0; i<players.length; i++) {
-        if (!(players[i].PER == "")) {
-            total += parseFloat(players[i].PER);
+        if (!(players[i].advanced == undefined)) {
+            total += parseFloat(players[i].advanced.PER);
         }
     }
     return total;
@@ -663,17 +663,17 @@ function bestTwoPlayers(players) {
         max1 = - Number.MAX_VALUE,
         max2 = - Number.MAX_VALUE;
     for (var i = 0; i<players.length; i++) {
-        if (! (players[i].PER == "")) {
-            if (parseFloat(players[i].PER) > parseFloat(max1)) {
-                max1 = players[i].PER;
+        if (! (players[i].advanced == undefined)) {
+            if (parseFloat(players[i].advanced.PER) > parseFloat(max1)) {
+                max1 = players[i].advanced.PER;
                 if (result.length > 0) {
                     max2 = max1;
                     result[1] = result[0];
                 }
                 result[0] = players[i];
             }
-            else if (parseFloat(players[i].PER) > parseFloat(max2)) {
-                max2 = players[i].PER;
+            else if (parseFloat(players[i].advanced.PER) > parseFloat(max2)) {
+                max2 = players[i].advanced.PER;
                 result[1] = players[i];
             }
         }
@@ -684,7 +684,7 @@ function bestTwoPlayers(players) {
 function playerPERDifference(players1, players2) {
     var result = [];
     for (var i = 0; i<players1.length && i<players2.length; i++) {
-        var diff = parseFloat(players2[i].PER) - parseFloat(players1[i].PER)
+        var diff = parseFloat(players2[i].advanced.PER) - parseFloat(players1[i].advanced.PER)
         result.push({
             "Player": players2[i].Player,
             "No.": players2[i]["No."],
@@ -711,12 +711,12 @@ function PERChangesPosition(players1, players2, position) {
     var currTemp = 0;
     console.log(result);
     for (var j = 0; j < previousPlayers.length; j++) {
-        prevTemp += parseFloat(previousPlayers[j].PER)
+        prevTemp += parseFloat(previousPlayers[j].advanced.PER)
     }
     prevTemp = parseFloat(prevTemp) / previousPlayers.length;
 
[…]

### 2016-04-20 field.js
Added field view

@@ -1,0 +1,185 @@
+function drawF(data, team, year, svg, x, y, width, height) {
+    players = data.find(function(d) {return d.year === year;})
+                .teams.find(function(d) {return d.team === team;})
+                .players;
+
+    players.sort(function (a, b) {
+        if (a.advanced.PER > b.advanced.PER) {
+            return -1;
+        } else if (a.advanced.PER < b.advanced.PER) {
+            return 1;
+        } else {
+            return 0;
+        }
+    });
+
+    d3.selectAll(".fieldchart").remove();
+
+
+    svg.append("g")
+        .attr("class", "chart fieldchart")
+        .append("image")
+        .attr("xlink:href", "img/court.png")
+        .attr("x", x)
+        .attr("y", y)
+        .attr("width", width)
+        .attr("height", height)
+        .attr("preserveAspectRatio", "none");
+
+    var chart = svg
+                .append("g")					//Append one div to the selected div in which we will construct the visualisation. This is done to separate mutliple visualisations..
+                .attr("class","chart fieldchart")
+                .attr("width", width)
+                .attr("height", height);
+
+    var backs = chart.append("g")
+        .attr("class", "backs")
+        .attr("width", width)
+        .attr("height", height);
+    
+    perScale = d3.scale.linear()
+                    .range([0.3, 3])
+                    .domain([0, 1600]);
+
+    var alreadyOn = {SG: {before: 0, xpos: 0.4 * width, ypos: 0.2 * height}
+                    ,PF: {before: 0, xpos: 0.05 * width, ypos: 0.9 * height - 100}
+                    ,C: {before: 0, xpos: 0.7 * width, ypos: 0.5 * height - 50}
+                    ,PG: {before: 0, xpos: 0.4 * width, ypos: 0.8 * height - 100}
+                    ,SF: {before: 0, xpos: 0.05 * width, ypos: 0.05 * height}
+                    ,'S-G': {before: 0, xpos: 10000, ypos: 600}};
+
+    for (var posName in alreadyOn) {
+        backs.append("rect")
+            .attr("x",alreadyOn[posName].xpos - 25)
+            .attr("y",alreadyOn[posName].ypos - width/50)
+            .attr("width", 20)
+            .attr("height", 0)
+            .attr("class", "posrect")
+            .attr("fill", "white")
[…]

### 2016-04-20 zoomview.js

@@ -1,4 +1,3 @@
-function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos, ypos, w, h) {
     var ShirtColors = {
         TorontoRaptors : {shirt: "#ff0000", edge:  "#000000"},
         BostonCeltics : {shirt: "#008000", edge: "#ffffff"},
@@ -46,6 +45,7 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
         NewOrleansHornets : {shirt: "009acd", edge: "551a8b"},
         CharlotteBobcats : {shirt: "000080", edge: "add8e6"}
     }
+function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos, ypos, w, h) {
 
     //console.log(teamWanted, yearWanted, arrowVariable);
     d3.select(".zoomchart").remove();

### 2016-04-20 Paper/Infovis Paper.tex

@@
 not professionally active in the field of basketball. Specifically, fans of NBA
-are the core of our target audience.
+are the core of our target audience. This means the visualization assumes most
+common basketball terms are known to the audience and as such does not provide
+additional information about them.
 
@@
 site basketball-reference\cite{basketball-reference}. A wide range of data is
-available on this site. In our visualization we only use data from 1971 onwards.
+available on this site. In our visualization we only use data from 1984 onwards.
 The data we use includes league standings and playoff rankings for each team,
@@
 
+The data was gathered by scraping the basketball-reference site, mostly using
+the provided download capabilities. Most of the data was downloaded in csv
+format, while some tables had to be manually scraped. The data was then combined
+in a preprocessing step. In this step, each team's playoff rankings were
+calculated based on the matches played during the playoffs, and the rest of the
+data was combined into json format. The final preprocessing step combines all
+data into one json file.
+
 \section{Related Work}\ref{sec:literature}
+
 \section{Visualization}\ref{sec:visualization}
+
+\subsection{Technology}
+To create the visualization, the d3 javascript framework~\cite{d3} was used in
+combination with html5~\cite{html5} and jquery~\cite{jquery}. This allows easy
+access to the visualization as most modern browsers are capable of handling
+these technologies.
+The choice not to use the d3 framework for the entire visualization was made to
+ease the layout configuration of the visualization. Instead, html was used to do
+the global layout of the visualization.
+
 \section{Discussion}\ref{sec:discussion}

### 2016-04-20 zoomview.js
Bugfixes

@@ -685,13 +685,9 @@ function playerPERDifference(players1, players2) {
     var result = [];
     for (var i = 0; i<players1.length && i<players2.length; i++) {
         var diff = parseFloat(players2[i].advanced.PER) - parseFloat(players1[i].advanced.PER)
-        result.push({
-            "Player": players2[i].Player,
-            "No.": players2[i]["No."],
-            "Pos": players2[i].Pos,
-            "PER": diff.toFixed(1)
-        });
-
+        var newPlayer = players2[i];
+        newPlayer.advanced.PER = diff.toFixed(1);
+        result.push(newPlayer);
     }
 
     return result;

### 2016-04-20 Paper/Infovis Paper.tex
First draft of the paper - YL

@@
 for phenomenons visible in the NBA data. By providing easy and intuitive access
-to the data, users can draw their own conclusions. 
-Secondly, the visualization's target audience are lay persons. This means people
-not professionally active in the field of basketball. Specifically, fans of NBA
-are the core of our target audience. This means the visualization assumes most
-common basketball terms are known to the audience and as such does not provide
-additional information about them.
+to the data, users can draw their own conclusions. It thus enables users of all 
+sorts to explore the data for patterns.
+Secondly, the visualization's target audience are lay persons. This means it 
+satifies needs of people regardless if they are professionally active in the 
+field of basketball or not. Specifically, fans of NBA are the core of our target
+audience. This means the visualization assumes most common basketball terms are
+known to the audience and as such does not provide additional information about 
+them.
+The visualization should enable people to look for changes in team statistics 
+and evaluate the impact on the team. The visualization gives also a short insight
+in what caused the change in statistics. Eg. a user could notice a spike/drop in 
+the SRS score () of a team. The user will then be able to see how the spike was 
+caused (team performance, players that joined the team, players that left the 
+team). Next to that, the user will can also explore the impact of the change on
+the team. She/he will be able to scroll through time and notice how the results of
+the team change over time since the spike/drop.
 
@@
 The data we use includes league standings and playoff rankings for each team,
-the team's roster and individual player statistics for each year, including the
-PER~(Player Efficiency Rating).
+team overall statistics, the team's roster and individual player statistics 
+for each year, including the PER~(Player Efficiency Rating).
 
@@
 \section{Visualization}\ref{sec:visualization}
+The visualization consists of three parts:
+- The bubble view: a compact view on the play-offs per season
+- The statistics (zoom) view: a more detailed view on a statistic of a selected 
+team
+- The team view: a detailed view on how good or bad a team scores on a specific 
+field position
+All three of them are discussed in more detail with the rationale behind them.
+
+\subsection{The bubble view}
+This view supposes to inform the user of the results of the NBA play-offs per 
+season. A circle is used to represent a team, the size its SRS score using a ''
+distribution, a stroke to illustrate it's region or medal and curved lines to
+inform the user which teams played against each other. The user has a timeline
+at its disposal to scroll through time.
+<picture of the bubble view compared with double elimination bracket>
+A more popular visualization of a play-off is the team double elimination bracket.
+Although this is a more common representation, we wanted to create a more compact
+version by eliminating the recurring representation of a team.
+Our initial bubble view was too compact. It connected teams that played against
+each other directly. When searching for alternatives to reflect on our choice we
+discovered a similar visualization in a totally different context. 
+<https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/>
+<comparison of original buble view with white house paths>
+This visualization gives a cleaner view of the competitors for a specific team
+compared to our original sketch. Hence we decided to adapt our visualization. We
+added an intermediate step between two teams to clearer indicate how teams competed
+to become NBA champion.
+Next to the static view, the user can interact with the visualization. When he 
+hoovers over a circle of a team, only the games played are shown by highliting the 
+competitors (their circles) and the curved lines that connect them. This view gets 
+fixated when the users clicks on a team.
+He/she then automatically gets forwarded to the statistics view where the bubble 
+view is still represented in the right upper corner to make sure the user keeps
+an overview. This view will be explained in more detail in the next section.
+
+\subsection{The statistics view}
+The statistics view gives a user a clear overview of how a statistic has been 
+influenced compared to the year before. The change is represented by three arrows:
+- arrow on the left pointing towards the middle: indicating the influence of players
+who joined the team
+- arrow in the middle: indicating how the team inherent changed
+- arrow on the right pointing towards the middle: indicating the influence of players
+who left the team
+<figure of the statistics view>
+The box on the right enables the user to change between statistics. Next to that it 
+informs the user of which statistic has been choosen. Below the arrows small multiples
+are shown of some statistics. The line charts show how a specific team scored over
+time. This enables a user to identify peaks or drops that possibly influenced the 
+outcome of a team is this season/following seasons or previous seasons.
+When the user hits the circle representing the teams SRS scrore, he is guided to 
+the team view. This view will be explained in more detail in the next section.
+
+\subsection{The team view}
+
+This view as well as the statistics view should give the user the ability to search
+for explanations why a team is is performing better or worse over seasons. On the 
+other hand, a user can see what impact a change in team characteristics has on its
+overall performance.
+Initially we designed this view more elaborate. When the user would click on a shirt,
+specific player information would be shown. Next to some general text information, 
+we would have visualised a number of player statistics as small multiples. A histogram
+should have informed how a player scores compared to his team, other players in the 
+league on this position or all other players in the league.
+<Sketch of the small multiples>
 

### 2016-04-20 Paper/Infovis Paper.tex
1ste draft na accidentele commit :-D

@@
 All three of them are discussed in more detail with the rationale behind them.
+The webpage is built in such a way that it gives a hint to the user at the bottom
+or the top of a page when another view is available. By showing a different 
+background color and an arrow, we invite the visitor to explore other views. The 
+user is also at liberty to scroll through these views. However, a fixed scroll is
+implemeted so users will only end in complete views when moving around. Note that 
+the user also will be guided through all 3 views when performing actions described 
+in the section below.
 
@@
 This view supposes to inform the user of the results of the NBA play-offs per 
-season. A circle is used to represent a team, the size its SRS score using a ''
-distribution, a stroke to illustrate it's region or medal and curved lines to
-inform the user which teams played against each other. The user has a timeline
-at its disposal to scroll through time.
+season. A circle is used to represent a team, the size its SRS score using a '?'
+distribution, a stroke to illustrate it's region or medal won during regional 
+competition and curved lines to inform the user which teams played against each 
+other. The user has a timeline at its disposal to scroll through time. Note that
+the team can be recognized by the logo inside the circle. Our initial desing filled
+the circle with the red or blue depending on the region where the teams compete. 
+However, the shade of colour was darker if their final ranking was higher in their
+region. We've replaced this with the logo's of the teams as we thought this is more
+user friendly and the extra information of the ranking doesn't outweight.
 <picture of the bubble view compared with double elimination bracket>
@@
 Although this is a more common representation, we wanted to create a more compact
-version by eliminating the recurring representation of a team.
+version by eliminating the recurring representation of a team. We also considered 
+enclosed circles and a sunburst.
+<figure of 2 alternatives https://thetuftersblog.wordpress.com/2016/03/07/team-ranking-play-off-infovis/>
+Although the enclosed circles is a compact representation as well, we have continued 
+with our own design as we believe this view is more original and innovative. We 
+didn't encouter a similar view when searching for alternatives to visualize a 
+play-off or championship. 
 Our initial bubble view was too compact. It connected teams that played against
@@
 The statistics view gives a user a clear overview of how a statistic has been 
-influenced compared to the year before. The change is represented by three arrows:
+influenced compared to the year before. The view consists of 3 parts:
+- Context section: the top part should inform the user of his/her context.
+- An arrow section illustrating why a statistic changed
+- Small multiples of team statistics and how they evolve over time
+\subsubsection{Context section}
+The context section should inform the user at all times of the current selections 
+that have been made. The statistics view will always be based on a specific team
+and season that have been selected or that are initated. This should be clear based
+on the textual information in the top left. Next to that, in the top right a small 
+bubble view is shown. It indicates the team selected and how well it performed in 
+the play-offs during that season. When changing teams and/or season, this view gets 
+updated as well as the textual information.
+
+\subsubsection{Arrow section of selected statistic}
+This section illustrates a change for a choosen statistic compared with the season
+the year before. This should give a visitor more insight into why a statistic changed
+over time. The change is represented by three arrows:
 - arrow on the left pointing towards the middle: indicating the influence of players
@@
 <figure of the statistics view>
-The box on the right enables the user to change between statistics. Next to that it 
-informs the user of which statistic has been choosen. Below the arrows small multiples
-are shown of some statistics. The line charts show how a specific team scored over
-time. This enables a user to identify peaks or drops that possibly influenced the 
-outcome of a team is this season/following seasons or previous seasons.
-When the user hits the circle representing the teams SRS scrore, he is guided to 
-the team view. This view will be explained in more detail in the next section.
+The shirts next to the arrows are the shirts of players that had the biggest 
+contribution to that arrow. Note that these shirts don't change in size. The box on 
+the right enables the user to change between 
+statistics. Next to that it informs the user of which statistic has been choosen. When 
+the user hits the circle in the middle representing the teams SRS scrore, he is guided 
+to the team view. This view will be explained in more detail in the next section.
+
+\subsubsection{Small multiples of different statistics}
+Below the arrow section small multiples are shown of multiple statistics over time. 
+The line charts enables a user to identify peaks or drops that possibly influenced the 
+outcome of a team is this season/following seasons or previous seasons. A user interact
+with the visualization by shifting the bar indicating the current year displayed. By doing
+this the view gets updated and the user can see the impact on the team in the context 
+section and why the statistic changes in the arrow section. An alternative to scroll
+through time is to use the timeline at the bottom of the page.
 
 \subsection{The team view}
-
 This view as well as the statistics view should give the user the ability to search
@@
 overall performance.
+This view shows each player of the team on its field postions. Shirts aligned next to 
+each other will thus share the same position. The size of a shirt would illustrate a
+players performance on this spot. The PES score () is initiated. This view should give
+a user insight on how strong or weak a particular team is for a specific field position.
+A user can then scroll through time, see how field positions evolve and what the influence
+is over time.
 Initially we designed this view more elaborate. When the user would click on a shirt,
 specific player information would be shown. Next to some general text information, 
-we would have visualised a number of player statistics as small multiples. A histogram
+we would have visualised a number of player statistics as small multiples. A bell curve
 should have informed how a player scores compared to his team, other players in the 
-league on this position or all other players in the league.
+league on this position or all other players in the league. We considered other charts for
+the small multiples as well. A box plot or bullet was less clear for us to indicate where
+a player stands compared to the group he was compared with. Alternatives for the small 
+multiples were evaluated as well. One good alternative was the shooting signature.
+<http://peterbeshai.com/buckets/app/#/playerView/201935_2015>
+However, creating such a visualization is again a project on itself and due to time 
+constraints we left it aside. Other possibilities such as a heatmap or spider web didn't
+fulfill our need to give the visitor a clear overview of how a player performed compared
+to its peergroup.
 <Sketch of the small multiples>
@@
 ease the layout configuration of the visualization. Instead, html was used to do
-the global layout of the visualization.
+the global layout of the visualization. We opted to structure the site with multiple
+divs as containers. To each container we allocate a svg with a specific visualization.
+Eg. The statistic view is build out of x divs filled with different svg's created
+with the D3 framework.
+<sketch of how the page is structured>
+When working with D3js we encountered some obstacles:
+- A synchrounous call need to be combined with synchronous calls. More specific, when a
+visitor uses the timeline to scroll through time, the whiping of previous visualizations 
+was not synced with the creation of the visualisation. The resulted in whiping sections 
+when a visualization was not created yet. Hence the view resulted in multiple figures 
+overlapping.
 
-\section{Discussion}\ref{sec:discussion}
+\section{Lessons learned}\ref{sec:discussion}
+We've learned a couple of lessons during the creation of this visualization:
+- Find a use case as soon as possible to built your visualization around
+- Each separate component should be a sufficient informative visualization on itself
+- Creating a custom visualization costs time and opportunities
+In what follows we will discuss our experience in more detail.
+
+\subsection{A good use case}
+At the beginning of this project we were able to design the 3 separate views 
+quite quickly. Our main problem was the integrate these views as a whole. When we created drafts
+of two views being the bubble view and the statitics view, we suddenly noted patterns. The golden
+globe warriors became NBA champion a couple of years after a great uplift in SRS score. This 
+due to a increase inside the team and because of new player joining the team in 2009. In 
+the years following, the team kept on attracting talented players and increased the inherent
+team score (with ups and downs) to become NBA champion in 2015. With the team view one can even
+notice that the team gets stronger in centre team positions during this period.
+The visualization however is missing at the moment if this reinforcement of the field positions
+is due to the attraction of talented people or utilizing available talent within the team.
+
+\subsection{Each component should be informative}
+In order to have a good visualization as a whole, each component should be a clear and 
+informative visualization on itself. This is a lesson learned from our work, but also when
+evaluating the work of peers. We started noticing that most components in a visualization 
+lack references and hence are not informative enough. Eg. each small multiple of a statisctic
+in the statistics view on itself gives sufficient information to stand on itselfs. The bar in
+the small multiple indicates the score in the selected season and enables the user to evaluate
+that score over time. Have there been better scores, or worse scores. Is this score part of 
+an upward movement during years? 
+For this reason we can argue that the circle representing the teams SRS score in the statisctics
+view is not the best choice. We could have chosen a bar chart with an indication of the best and
+worst score for that specific team over time or for the best & worst team in that season.
+
+\subsection{Custom visualizations]
+Before exploring what is out there in the D3js world, we made our own sketches. This is how
+we came up with the 3 custom views that makes our visualization. When comparing our design 
+with alternatives we could not find something that satisfied our needs. Nonetheless, they gave
+us inspiration to fine tune our designs. Although we are happy very happy with our result and
+we are convinced it was the right choice reach our goal, it had some draw backs.
+We noted that other teams had more freedom in exploring multiple different visualizations which
+enabled them to evaluate different options with trial and error. A luxury we did not have. We 
+could have prepared our data in a format that was more compatible with out of the box examples,
+but this would have had too much of an impact on our result.
+Next to that, although we've tried to write our code as general as possible, we're not convinced
+this could be shared in such a way that other information visualizer could easily re-use our work
+to vizualise other sports.
+
 \section{Conclusion}\ref{sec:conclusion}

### 2016-04-21 Paper/paper.tex
Bugfix and minor updates

--- 2016-04-20 Paper/Infovis Paper.tex
@@
 \usepackage{textcomp}
+
+\usepackage[
+    backend=biber,
+    style=draft
+]{biblatex}
+
+\addbibresource{paper.bib}
 
@@
 a player stands compared to the group he was compared with. Alternatives for the small 
-multiples were evaluated as well. One good alternative was the shooting signature.
-<http://peterbeshai.com/buckets/app/#/playerView/201935_2015>
+multiples were evaluated as well. One good alternative was the shooting signature.\cite{peterbeshai}
 However, creating such a visualization is again a project on itself and due to time 
@@
 view is not the best choice. We could have chosen a bar chart with an indication of the best and
-worst score for that specific team over time or for the best & worst team in that season.
+worst score for that specific team over time or for the best and worst team in that season.
 
-\subsection{Custom visualizations]
+\subsection{Custom visualizations}
 Before exploring what is out there in the D3js world, we made our own sketches. This is how
@@
 % References must be the same font size as other body text.
-\bibliographystyle{SIGCHI-Reference-Format}
-\bibliography{sample}
+%\bibliographystyle{SIGCHI-Reference-Format}
+\printbibliography
 

### 2016-04-21 field.js

@@ -1,3 +1,13 @@
+var perScale = d3.scale.linear()
+                    .range([0.3, 3])
+                    .domain([0, 1600]);
+
+function exagerratedPerScale(per) {
+    var exagerratedPer = Math.pow(per, 2);
+    if (per < 0) { exagerratedPer = 0; }
+    return perScale(exagerratedPer);
+}
+
 function drawF(data, team, year, svg, x, y, width, height) {
     players = data.find(function(d) {return d.year === year;})
                 .teams.find(function(d) {return d.team === team;})
@@ -32,21 +42,32 @@ function drawF(data, team, year, svg, x, y, width, height) {
                 .attr("width", width)
                 .attr("height", height);
 
+    var tip = d3.tip()
+                .attr('class', 'd3-tip')
+                .offset([-10, 0])
+                .html(function(d) {
+                    console.log(d);
+                    return  "<p><span style='color:orange'>" + d["Player"] + "</span> <\p>" +
+                        "PER: <span style='color:red'>" + d.advanced["PER"] + "</span> </br>" +
+                        "Height: <span style='color:red'>" + d["Ht"] + "</span> </br>" +
+                        "Weight: <span style='color:red'>" + d["Wt"] + "</span> </br>" +
+                        "Birthday: <span style='color:red'>" + d["Birth Date"] + "</span> </br>";
+                });
+    chart.call(tip);
+
+
+
     var backs = chart.append("g")
         .attr("class", "backs")
         .attr("width", width)
         .attr("height", height);
     
-    perScale = d3.scale.linear()
-                    .range([0.3, 3])
-                    .domain([0, 1600]);
 
-    var alreadyOn = {SG: {before: 0, xpos: 0.4 * width, ypos: 0.2 * height}
-                    ,PF: {before: 0, xpos: 0.05 * width, ypos: 0.9 * height - 100}
-                    ,C: {before: 0, xpos: 0.7 * width, ypos: 0.5 * height - 50}
-                    ,PG: {before: 0, xpos: 0.4 * width, ypos: 0.8 * height - 100}
-                    ,SF: {before: 0, xpos: 0.05 * width, ypos: 0.05 * height}
-                    ,'S-G': {before: 0, xpos: 10000, ypos: 600}};
+    var alreadyOn = {SG: {before: 0, xpos: 0.4 * width, ypos: 0.2 * height, fullname: "Shooting Guard"}
[…]

### 2016-04-21 scraper/remove.py

@@ -15,13 +15,18 @@ for yeardata in data:
         print "Doing team " + teamname
         team['info'] = { "PTS/G": team['info']['PTS/G'], "FG%": team['info']['FG%']}
         team['opponent'] = { "PTS/G": team['info']['PTS/G']}
-        team['misc'] = {"Attendance": team['misc']['Attendance']}
+        team['misc'] = {"Attendance": team['misc']['Attendance'], "Age": team['misc']['Age']}
+        newplayers = []
         for player in team['players']:
-            del player['pergame']
-            del player['perminute']
-            del player['perposs']
-            del player['totals']
-            player['advanced'] = { "PER": player['advanced']['PER'] }
+            if player['advanced']['G'] > 5:
+                del player['pergame']
+                del player['perminute']
+                del player['perposs']
+                del player['totals']
+                player['advanced'] = { "PER": player['advanced']['PER'] }
+                newplayers.append(player)
+
+        team['players'] = newplayers
 
 with open('removed.json', 'w') as outjsonfile:
     json.dump(data, outjsonfile, indent=4)

### 2016-04-21 zoomview.js

@@ -92,6 +92,18 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
                 .attr("width", w)
                 .attr("height", h);
 
+    var tip = d3.tip()
+                .attr('class', 'd3-tip')
+                .offset([-10, 0])
+                .html(function(d) {
+                    console.log(d);
+                    return  "<p><span style='color:orange'>" + d["Player"] + "</span> <\p>" +
+                        "PER: <span style='color:red'>" + d.advanced["PER"] + "</span> </br>" +
+                        "Height: <span style='color:red'>" + d["Ht"] + "</span> </br>" +
+                        "Weight: <span style='color:red'>" + d["Wt"] + "</span> </br>" +
+                        "Birthday: <span style='color:red'>" + d["Birth Date"] + "</span> </br>";
+                });
+    chart.call(tip);
     var defs = chart.append("defs");
 
     defs.append("path")
@@ -380,9 +392,23 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
         var height = (2*maxRad + 2*margin.top + 30);
         var w1 = (w/2 - maxRad/4 + 40);
         if (bestPlayers.length >= 1) {
-            drawShirt(w1, height, bestPlayers[0].Player, bestPlayers[0]["No."], teamWanted);
+            //drawShirt(w1, height, bestPlayers[0].Player, bestPlayers[0]["No."], teamWanted);
+            var wh = drawScaledShirt(w1,
+                            height, 
+                            bestPlayers[0], 
+                            teamWanted, 
+                            chart, 
+                            exagerratedPerScale(bestPlayers[0].advanced.PER),
+                            tip);
             if (bestPlayers.length >= 2) {
-                drawShirt(w1 + 50 + margin.left, height, bestPlayers[1].Player, bestPlayers[1]["No."], teamWanted);
+                //drawShirt(w1 + 50 + margin.left, height, bestPlayers[1].Player, bestPlayers[1]["No."], teamWanted);
+                drawScaledShirt(w1 + wh.width + margin.left,
+                                height, 
+                                bestPlayers[1], 
+                                teamWanted, 
+                                chart, 
+                                exagerratedPerScale(bestPlayers[1].advanced.PER),
+                                tip);
             }
         }
     }
@@ -390,9 +416,23 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
[…]

### 2016-04-22 paper/paper.tex
A bit of work on the paper

--- 2016-04-21 Paper/paper.tex
@@
 % so you remain anonymous.
-\def\plaintitle{Visualizing NBA teams across multiple seasons}
+\def\plaintitle{Why did they win?\\Visualizing NBA teams across multiple seasons}
 \def\plainauthor[#1]{#1}
@@
 \def\emptyauthor{}
-\def\plainkeywords{Authors' choice; of terms; separated; by
-  semicolons; include commas, within terms only; required.}
-\def\plaingeneralterms{Documentation, Standardization}
+\def\plainkeywords{Information Visualization; CHI; NBA}
+\def\plaingeneralterms{}
 
@@
 This paper describes the visualization of NBA data made by the ``The Tufters''
-team for the course ``Information Visualization''. In section~\ref{sec:goal}
-we describe the goal of the visualization and the target audience. In
+team for the course ``Information Visualization''. In section~\ref{sec:goal} we
+describe the goal of the visualization and the target audience. In
 section~\ref{sec:data} we describe the data used, including its origins,
 advantages and limitations. In section~\ref{sec:literature} we give an overview
-of related literature and web resources, including related visualizations.
-Related visualizations include both visualizations of NBA or other sports data,
-as well as visualizations tackling a problem we encountered during the
-development of our visualization. In section~\ref{sec:visualization} we describe
-the visualization itself. We give an overview of the different stages of
-development of the visualization, as well as the major design decisions made
-during the development process. Section~\ref{sec:discussion} discusses potential
-improvements of the final visualization and lessons learned from the project. We
-conclude in section~\ref{sec:conclusion} by giving a short overview.
+of related literature and web resources, including related visualizations.  This
+includes both visualizations of NBA or other sports data, as well as
+visualizations tackling an isue encountered during the development of our
+visualization. In section~\ref{sec:visualization} we describe the visualization
+itself. We give an overview of the different stages of development of the
+visualization, as well as the major design decisions.
+Section~\ref{sec:discussion} discusses potential improvements of the final
+visualization and lessons learned from the project. We conclude in
+section~\ref{sec:conclusion}. Appendix~\ref{sec:terminology} contains
+definitions for common basketball terms used in this paper.
 
-\section{Goal and Target Audience}\ref{sec:goal}
-The visualization's goal is allowing exploration of NBA data by lay persons.
-More specifically, the visualization does not try to offer premade explanations
-for phenomenons visible in the NBA data. By providing easy and intuitive access
-to the data, users can draw their own conclusions. It thus enables users of all 
-sorts to explore the data for patterns.
-Secondly, the visualization's target audience are lay persons. This means it 
-satifies needs of people regardless if they are professionally active in the 
-field of basketball or not. Specifically, fans of NBA are the core of our target
-audience. This means the visualization assumes most common basketball terms are
-known to the audience and as such does not provide additional information about 
-them.
-The visualization should enable people to look for changes in team statistics 
-and evaluate the impact on the team. The visualization gives also a short insight
-in what caused the change in statistics. Eg. a user could notice a spike/drop in 
-the SRS score () of a team. The user will then be able to see how the spike was 
-caused (team performance, players that joined the team, players that left the 
-team). Next to that, the user will can also explore the impact of the change on
-the team. She/he will be able to scroll through time and notice how the results of
-the team change over time since the spike/drop.
+\section{Goal and Target Audience}\label{sec:goal} 
+The visualization's goal can be best summarized by the following sentence: ``Why
+did they win?'' The visualization focuses on exposing relations between team
+performance and their player roster over several years. It allows users to find
+explanations for major improvements and declines in team performance.  The
+visualization allows exploration of NBA data by lay persons.  More specifically,
+the visualization does not try to offer premade explanations for phenomenons
+visible in the NBA data. By providing easy and intuitive access to the data, the
+visualization allows users to draw their own conclusions. The visualization's
+target audience are lay persons. Specifically, fans of NBA are the core of our
+target audience. This means the visualization assumes most common basketball
+terms are known to the audience and does not provide basic terminology
+explanations.
 
-\section{Data}\ref{sec:data}
+\section{Data}\label{sec:data}
 The data visualized is a subset of the data available on basketball statistics
 site basketball-reference\cite{basketball-reference}. A wide range of data is
-available on this site. In our visualization we only use data from 1984 onwards.
-The data we use includes league standings and playoff rankings for each team,
-team overall statistics, the team's roster and individual player statistics 
-for each year, including the PER~(Player Efficiency Rating).
+available on this site. This includes common basketball performance statistics,
+such as field goals, percentage of shots scored, shooting distance, minutes
+played per season, number of personal fouls made, salaries and much more. These
+statistics are avalable for each player individually for each season. Aggregated
+data for entire teams and for entire seasons is also available. Additionally,
+idividual game statistics are also available, including playoff games.  In our
+visualization we only use data from 1984 onwards. This is done because of rule
+changes in the NBA which changed the number of teams competing and the
+competition's structure. To simplify implementation of our visualization, only
+data after the last major rule change was used.  The data we use includes league
+standings and playoff rankings for each team, team overall statistics, the
+team's roster and individual player statistics for each year, including the
+PER~(Player Efficiency Rating)\cite{per}.
 
-The data was gathered by scraping the basketball-reference site, mostly using
-the provided download capabilities. Most of the data was downloaded in csv
-format, while some tables had to be manually scraped. The data was then combined
-in a preprocessing step. In this step, each team's playoff rankings were
-calculated based on the matches played during the playoffs, and the rest of the
-data was combined into json format. The final preprocessing step combines all
-data into one json file.
+The data was gathered by scraping the basketball-reference site using the
+provided download capabilities. Most of the data was downloaded in csv format,
+while some tables had to be manually scraped. This process was automated using
+Python scripts. The data was then combined in a preprocessing step. In this
+step, each team's playoff rankings were calculated based on the matches played
+during the playoffs, and the rest of the data was combined into json format. The
+final preprocessing step combined all data into one json file.
 
-\section{Related Work}\ref{sec:literature}
+\section{Related Work}\label{sec:literature}
 
-\section{Visualization}\ref{sec:visualization}
+\section{Visualization}\label{sec:visualization}
 The visualization consists of three parts:
-- The bubble view: a compact view on the play-offs per season
-- The statistics (zoom) view: a more detailed view on a statistic of a selected 
+\paragraph{The bubble view} a compact view on the play-offs per season
+\paragraph{The statistics (zoom) view} a more detailed view on a statistic of a selected 
 team
-- The team view: a detailed view on how good or bad a team scores on a specific 
+\paragraph{The team view} a detailed view on how good or bad a team scores on a specific 
 field position
-All three of them are discussed in more detail with the rationale behind them.
-The webpage is built in such a way that it gives a hint to the user at the bottom
-or the top of a page when another view is available. By showing a different 
-background color and an arrow, we invite the visitor to explore other views. The 
-user is also at liberty to scroll through these views. However, a fixed scroll is
-implemeted so users will only end in complete views when moving around. Note that 
-the user also will be guided through all 3 views when performing actions described 
-in the section below.
+All three are discussed in more detail below. The three views are available to
+the user on a single webpage. The user start on the bubble view and is able to
+navigate to the statistics and team views by scrolling down or clicking on
+teams.
+%The webpage is built in such a way that it gives a hint to the user at the bottom
+%or the top of a page when another view is available. By showing a different 
+%background color and an arrow, we invite the visitor to explore other views. The 
+%user is also at liberty to scroll through these views. However, a fixed scroll is
+%implemeted so users will only end in complete views when moving around. Note that 
+%the user also will be guided through all 3 views when performing actions described 
+%in the section below.
 
@@
 influenced compared to the year before. The view consists of 3 parts:
-- Context section: the top part should inform the user of his/her context.
-- An arrow section illustrating why a statistic changed
-- Small multiples of team statistics and how they evolve over time
+\paragraph{Context section} the top part should inform the user of his/her context.
+\paragraph{Arrow section} This section illustrates why a statistic changed
+\paragraph{Small multiples section} Small multiples of team statistics and how they evolve over time
 \subsubsection{Context section}
@@
 over time. The change is represented by three arrows:
-- arrow on the left pointing towards the middle: indicating the influence of players
+\begin{itemize}
+    \item arrow on the left pointing towards the middle: indicating the influence of players
     who joined the team
-- arrow in the middle: indicating how the team inherent changed
-- arrow on the right pointing towards the middle: indicating the influence of players
+    \item arrow in the middle: indicating how the team inherent changed
+    \item arrow on the right pointing towards the middle: indicating the influence of players
 who left the team
+\end{itemize}
 <figure of the statistics view>
@@
 
-\section{Lessons learned}\ref{sec:discussion}
+\section{Lessons learned}\label{sec:discussion}
 We've learned a couple of lessons during the creation of this visualization:
@@
 
-\section{Conclusion}\ref{sec:conclusion}
+\section{Conclusion}\label{sec:conclusion}
 
@@
 
+\begin{appendices}
+
+    \section{Terminology}\label{sec:terminology}
+
+    \paragraph{PER} Player Efficiency Rating - An all-in-one basketball rating,
+    boiling down all of a player's contributions into one number\cite{per}.
+\end{appendices}
+
 \end{document}

### 2016-04-22 paper/paper.tex
Small changes to paper, compiled

@@
 The visualization consists of three parts:
-\paragraph{The bubble view} a compact view on the play-offs per season
-\paragraph{The statistics (zoom) view} a more detailed view on a statistic of a selected 
-team
-\paragraph{The team view} a detailed view on how good or bad a team scores on a specific 
-field position
+\begin{itemize}
+    \item The bubble view: a compact view on the play-offs per season
+    \item The statistics (zoom) view: a more detailed view on a statistic of a
+        selected team
+    \item The team view: a detailed view on how good or bad a team scores on a
+        specific field position
+\end{itemize}
 All three are discussed in more detail below. The three views are available to
@@
 influenced compared to the year before. The view consists of 3 parts:
-\paragraph{Context section} the top part should inform the user of his/her context.
-\paragraph{Arrow section} This section illustrates why a statistic changed
-\paragraph{Small multiples section} Small multiples of team statistics and how they evolve over time
+\begin{itemize}
+    \item Context section: the top part should inform the user of his/her
+        context.
+    \item Arrow section: This section illustrates why a statistic changed
+    \item Small multiples section: Small multiples of team statistics and how
+        they evolve over time
+\end{itemize}
 \subsubsection{Context section}
@@
 We've learned a couple of lessons during the creation of this visualization:
-- Find a use case as soon as possible to built your visualization around
-- Each separate component should be a sufficient informative visualization on itself
-- Creating a custom visualization costs time and opportunities
+\begin{itemize}
+    \item Find a use case as soon as possible to build your visualization around
+    \item Each separate component should be a sufficient informative visualization on itself
+    \item Creating a custom visualization costs time and opportunities
+\end{itemize}
 In what follows we will discuss our experience in more detail.
@@
 
-\begin{appendices}
-
+\appendix
 \section{Terminology}\label{sec:terminology}
@@
 boiling down all of a player's contributions into one number\cite{per}.
-\end{appendices}
 

### 2016-04-22 paper/paper.tex
Kleine aanpassingen

--- 2016-04-22 paper/paper.tex
@@
 data for entire teams and for entire seasons is also available. Additionally,
-idividual game statistics are also available, including playoff games.  In our
+individual game statistics are also available, including playoff games.  In our
 visualization we only use data from 1984 onwards. This is done because of rule
@@
 The visualization consists of three parts:
-\paragraph{The bubble view} a compact view on the play-offs per season
-\paragraph{The statistics (zoom) view} a more detailed view on a statistic of a selected 
-team
-\paragraph{The team view} a detailed view on how good or bad a team scores on a specific 
-field position
+\paragraph{The bubble view} 
+A compact view on the play-offs per season.
+\paragraph{The statistics (zoom) view} 
+A more detailed view on a statistic of a selected 
+team.
+\paragraph{The team view} 
+A detailed view on how good or bad a team scores on a specific 
+field position. \\
+\newline
 All three are discussed in more detail below. The three views are available to
@@
 <https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/>
-<comparison of original buble view with white house paths>
+<comparison of original bubble view with white house paths>
 This visualization gives a cleaner view of the competitors for a specific team
@@
 Next to the static view, the user can interact with the visualization. When he 
-hoovers over a circle of a team, only the games played are shown by highliting the 
+hoovers over a circle of a team, only the games played by that team are shown by highliting the 
 competitors (their circles) and the curved lines that connect them. This view gets 
@@
 view is still represented in the right upper corner to make sure the user keeps
-an overview. This view will be explained in more detail in the next section.
+an overview. The statistics view will be explained in more detail in the next section.
 
@@
 influenced compared to the year before. The view consists of 3 parts:
-\paragraph{Context section} the top part should inform the user of his/her context.
-\paragraph{Arrow section} This section illustrates why a statistic changed
-\paragraph{Small multiples section} Small multiples of team statistics and how they evolve over time
+\paragraph{Context section} 
+The top part informs the user about the shown context.
+\paragraph{Arrow section} 
+This section illustrates why a statistic changed.
+\paragraph{Small multiples section} 
+Small multiples of team statistics and how they evolve over time.
 \subsubsection{Context section}
@@
 bubble view is shown. It indicates the team selected and how well it performed in 
-the play-offs during that season. When changing teams and/or season, this view gets 
+the play-offs during that season, as ecplained in the bubble view section. When changing teams and/or season, this view gets 
 updated as well as the textual information.
@@
 \subsubsection{Arrow section of selected statistic}
-This section illustrates a change for a choosen statistic compared with the season
-the year before. This should give a visitor more insight into why a statistic changed
-over time. The change is represented by three arrows:
+This section illustrates a change for a chosen statistic for the selected team compared with the season before.
+This should give a visitor more insight into why a statistic changed over time.
+The selected team is represented by a circle in the same way as explained in the bubble view.
+The change for a statistic is represented by three arrows:
 \begin{itemize}
-    \item arrow on the left pointing towards the middle: indicating the influence of players
-    who joined the team
-    \item arrow in the middle: indicating how the team inherent changed
-    \item arrow on the right pointing towards the middle: indicating the influence of players
-who left the team
+    \item arrow on the left pointing towards the circle: indicating the influence of players who joined the team (team member current season, not previous season)
+    \item arrow on the right of the circle: indicating how the team inherent changed (team member this season and previous season)
+    \item arrow on the left pointing away from the circle: indicating the influence of players who left the team (team member previous season, not current season)
 \end{itemize}
@@
 the right enables the user to change between 
-statistics. Next to that it informs the user of which statistic has been choosen. When 
-the user hits the circle in the middle representing the teams SRS scrore, he is guided 
+statistics. Next to that it informs the user of which statistic has been chosen. When 
+the user hits the circle in the middle representing the teams, he is guided 
 to the team view. This view will be explained in more detail in the next section.
@@
 The line charts enables a user to identify peaks or drops that possibly influenced the 
-outcome of a team is this season/following seasons or previous seasons. A user interact
-with the visualization by shifting the bar indicating the current year displayed. By doing
-this the view gets updated and the user can see the impact on the team in the context 
-section and why the statistic changes in the arrow section. An alternative to scroll
-through time is to use the timeline at the bottom of the page.
+outcome of a team is this season, following seasons or previous seasons.
+A user interacts with the visualization by shifting the bar indicating the current year displayed.
+By doing this the view gets updated and the user can see the impact on the team in the context section and why the statistic changes in the arrow section.
+An alternative to scroll through time is to use the timeline at the bottom of the page.
 
@@
 overall performance.
-This view shows each player of the team on its field postions. Shirts aligned next to 
-each other will thus share the same position. The size of a shirt would illustrate a
-players performance on this spot. The PES score () is initiated. This view should give
+This view shows each player of the team on its field positions. Shirts aligned next to 
+each other will thus share the same position. The size of a shirt would illustrate a players performance on this spot. The PER score is initiated. This view should give
 a user insight on how strong or weak a particular team is for a specific field position.
@@
 \subsection{Technology}
-To create the visualization, the d3 javascript framework~\cite{d3} was used in
-combination with html5~\cite{html5} and jquery~\cite{jquery}. This allows easy
+To create the visualization, the d3 javascript framework~\cite{d3.js} was used in combination with html5~\cite{html5} and jquery~\cite{jquery}. This allows easy
 access to the visualization as most modern browsers are capable of handling
@@
 Eg. The statistic view is build out of x divs filled with different svg's created
-with the D3 framework.
+with the d3 framework.
 <sketch of how the page is structured>
-When working with D3js we encountered some obstacles:
-- A synchrounous call need to be combined with synchronous calls. More specific, when a
-visitor uses the timeline to scroll through time, the whiping of previous visualizations 
-was not synced with the creation of the visualisation. The resulted in whiping sections 
-when a visualization was not created yet. Hence the view resulted in multiple figures 
-overlapping.
+When working with d3.js we encountered some obstacles:
+- A synchrounous call need to be combined with synchronous calls.
+More specific, when a visitor uses the timeline to scroll through time, the whiping of previous visualizations was not synced with the creation of the visualisation.
+This resulted in whiping sections when a visualization was not created yet.
+Hence the view resulted in multiple figures overlapping.
 
 \section{Lessons learned}\label{sec:discussion}
+
 We've learned a couple of lessons during the creation of this visualization: 
-- Find a use case as soon as possible to built your visualization around
-- Each separate component should be a sufficient informative visualization on itself
-- Creating a custom visualization costs time and opportunities
+\begin{itemize}
+    \item find a use case as soon as possible to built your visualization around
+    \item each separate component should be a sufficient informative visualization on itself
+    \item creating a custom visualization costs time and opportunities 
+\end{itemize}
+
 In what follows we will discuss our experience in more detail.
@@
 At the beginning of this project we were able to design the 3 separate views 
-quite quickly. Our main problem was the integrate these views as a whole. When we created drafts
-of two views being the bubble view and the statitics view, we suddenly noted patterns. The golden
-globe warriors became NBA champion a couple of years after a great uplift in SRS score. This 
-due to a increase inside the team and because of new player joining the team in 2009. In 
-the years following, the team kept on attracting talented players and increased the inherent
-team score (with ups and downs) to become NBA champion in 2015. With the team view one can even
-notice that the team gets stronger in centre team positions during this period.
+quite quickly.
+Our main problem was to integrate these views as a whole.
+When we created drafts of two views being the bubble view and the statitics view, we suddenly noted patterns.
+The Golden State Warriors became NBA champion a couple of years after a great uplift in SRS score.
+This due to a increase inside the team and because of new players joining the team in 2009.
+In
+the years following, the team kept on attracting talented players and increased the inherent team score (with ups and downs) to become NBA champion in 2015.
+With the team view, one can even notice that the team gets stronger in center team positions during this period.
 The visualization however is missing at the moment if this reinforcement of the field positions
-is due to the attraction of talented people or utilizing available talent within the team.
+is due to the attraction of talented people or by utilizing available talent within the team.
 
@@
 \subsection{Custom visualizations}
-Before exploring what is out there in the D3js world, we made our own sketches. This is how
-we came up with the 3 custom views that makes our visualization. When comparing our design 
-with alternatives we could not find something that satisfied our needs. Nonetheless, they gave
-us inspiration to fine tune our designs. Although we are happy very happy with our result and
-we are convinced it was the right choice reach our goal, it had some draw backs.
-We noted that other teams had more freedom in exploring multiple different visualizations which
-enabled them to evaluate different options with trial and error. A luxury we did not have. We 
-could have prepared our data in a format that was more compatible with out of the box examples,
-but this would have had too much of an impact on our result.
-Next to that, although we've tried to write our code as general as possible, we're not convinced
-this could be shared in such a way that other information visualizer could easily re-use our work
-to vizualise other sports.
+Before exploring what is out there in the d3.js world, we made our own sketches.
+This is how we came up with the 3 custom views that makes our visualization.
+When comparing our design with alternatives we could not find something that satisfied our needs.
+Nonetheless, they gave us inspiration to finetune our designs.
+Although we are satisfied with our result and we are convinced it was the right choice to reach our goal, it had some draw backs.
+We noted that other teams had more freedom in exploring multiple different visualizations which enabled them to evaluate different options with trial and error.
+We could have prepared our data in a format that was more compatible with out of the box examples, but this would have had too much of an impact on our result.
+Next to that, although we've tried to write our code as general as possible, we're not convinced this could be shared in such a way that other people building a visualization could easily re-use our work to visualize other sports.
 

### 2016-04-22 paper/paper.tex

@@
 
+\pagenumbering{arabic}
+
 % create a shortcut to typeset table headings
@@
 The visualization consists of three parts:
-\paragraph{The bubble view} 
-A compact view on the play-offs per season.
-\paragraph{The statistics (zoom) view} 
-A more detailed view on a statistic of a selected 
-team.
-\paragraph{The team view} 
-A detailed view on how good or bad a team scores on a specific 
-field position. \\
-\newline
+\begin{itemize}
+    \item The bubble view: a compact view on the play-offs per season
+    \item The statistics (zoom) view: a more detailed view on a statistic of a selected team
+    \item The team view: a detailed view on how good or bad a team scores on a specific field position
+\end{itemize}
 All three are discussed in more detail below. The three views are available to
@@
 \subsection{The statistics view}
-The statistics view gives a user a clear overview of how a statistic has been 
-influenced compared to the year before. The view consists of 3 parts:
-\paragraph{Context section} 
-The top part informs the user about the shown context.
-\paragraph{Arrow section} 
-This section illustrates why a statistic changed.
-\paragraph{Small multiples section} 
-Small multiples of team statistics and how they evolve over time.
+The statistics view gives a user a clear overview of how a statistic has been influenced compared to the year before. The view consists of 3 parts:
+\begin{itemize}
+    \item Context section: the top part should inform the user of his/her
+        context
+    \item Arrow section: hhis section illustrates why a statistic changed
+    \item Small multiples section: hmall multiples of team statistics and how they evolve over time
+\end{itemize}
+
 \subsubsection{Context section}
@@
 \begin{itemize}
-    \item arrow on the left pointing towards the circle: indicating the influence of players who joined the team (team member current season, not previous season)
-    \item arrow on the right of the circle: indicating how the team inherent changed (team member this season and previous season)
-    \item arrow on the left pointing away from the circle: indicating the influence of players who left the team (team member previous season, not current season)
+    \item Arrow on the left pointing towards the circle: indicating the influence of players who joined the team (team member current season, not previous season)
+    \item Arrow on the right of the circle: indicating how the team inherent changed (team member this season and previous season)
+    \item Arrow on the left pointing away from the circle: indicating the influence of players who left the team (team member previous season, not current season)
 \end{itemize}
@@
 \section{Lessons learned}\label{sec:discussion}
-
 We've learned a couple of lessons during the creation of this visualization:
 \begin{itemize}
-    \item find a use case as soon as possible to built your visualization around
-    \item each separate component should be a sufficient informative visualization on itself
-    \item creating a custom visualization costs time and opportunities 
+    \item Find a use case as soon as possible to build your visualization around
+    \item Each separate component should be a sufficient informative visualization on itself
+    \item Creating a custom visualization costs time and opportunities
 \end{itemize}
-
 In what follows we will discuss our experience in more detail.

### 2016-04-23 zoomview.js
Better use of space, bug in stayedArrow, buttons not yet done

@@ -1,118 +1,184 @@
-    var ShirtColors = {
-        TorontoRaptors : {shirt: "#ff0000", edge:  "#000000"},
-        BostonCeltics : {shirt: "#008000", edge: "#ffffff"},
-        BrooklynNets : {shirt: "#000000", edge: "#ffffff"},
-        Philadelphia76ers : {shirt: "#ff0000", edge: "#0000ff"},
-        NewYorkKnicks  : {shirt: "#0000ff;", edge: "#ff4500"},
-        ClevelandCavaliers : {shirt: "#8b0000", edge: "#ffd700"},
-        ChicagoBulls : {shirt: "#ff0000", edge: "#000000"},
-        MilwaukeeBucks : {shirt: "#006400", edge: "#ff0000"},
-        IndianaPacers : {shirt: "#000080", edge: "#ffd700"},
-        DetroitPistons : {shirt: "#0000ff", edge: "##ff0000"},
-        AtlantaHawks : {shirt: "#000080", edge: "#ff0000"},
-        WashingtonWizards : {shirt: "#ff0000", edge: "#000080"},
-        MiamiHeat : {shirt: "#000000", edge: "#8b0000"},
-        CharlotteHornets : {shirt: "#800080", edge: "#008080"},
-        OrlandoMagic : {shirt: "#0000ff", edge: "#000000"},
-        PortlandTrailBlazers : {shirt: "#000000", edge: "#ff0000"},
-        OklahomaCityThunder : {shirt: "#009acd", edge: "#00008b"},
-        UtahJazz : {shirt: "#000080", edge: "#006400"},
-        DenverNuggets : {shirt: "#6ca6cd", edge: "#000080"},
-        MinnesotaTimberwolves : {shirt: "#00688b", edge: "#000000"},
-        GoldenStateWarriors : {shirt: "#0000ff", edge: "#ffd700"},
-        LosAngelesClippers : {shirt: "#ff0000", edge: "#0000ff"},
-        PhoenixSuns : {shirt: "#800080", edge: "#ff8c00"},
-        SacramentoKings : {shirt: "#800080", edge: "#000000"},
-        LosAngelesLakers : {shirt: "#800080", edge: "#ffd700"},
-        HoustonRockets : {shirt: "#ff0000", edge: "#ffffff"},
-        SanAntonioSpurs : {shirt: "#000000", edge: "#c0c0c0"},
[…]

### 2016-04-23 paper/paper.tex
update niet compleet :-(

@@
 
+<Tell something more about the data in terms of tall/wide data
+Tell something more about pre-processing steps (why?, did we change it over time?)
+rfr to picture in 1st ppt sl 60>
 \section{Related Work}\label{sec:literature}
@@
 \end{itemize}
+
+<Explain that this complies with the infovis seeking mantra: Overview, zoom&filter, details on demand
+Bubble = overview
+statistics= details on demand
+>
+
 All three are discussed in more detail below. The three views are available to
@@
 an overview. The statistics view will be explained in more detail in the next section.
+< Hierarchy = sorting = facilitate trends of winners
+position have a meaning in the bubble view
+when selecting circle => zoom/more detail
+Bubble principle of 
+- Symmetry (whole = play-off)
+- conectedness arc = games played
+Pre attentive characteristics
+- size of circles
+- hue to highligt selections or indicate increase decline arrows next view
+
+We considered a map view, but didn't do it because geographical position has no added value
+This visualization helped us discover data quality issues => cut off year 1984
+>
 
@@
 bubble view is shown. It indicates the team selected and how well it performed in 
-the play-offs during that season, as ecplained in the bubble view section. When changing teams and/or season, this view gets 
+the play-offs during that season, as explained in the bubble view section. When changing teams and/or season, this view gets 
 updated as well as the textual information.
+<Explain that the context supports the short term memory of human beings rfr. Monkey Illusion>
 
@@
 to the team view. This view will be explained in more detail in the next section.
+<filtering: 
+- time & team
+- based on exploration (small multiples)
+- dynamically: changes impact multiple parts>
 
@@
 An alternative to scroll through time is to use the timeline at the bottom of the page.
+<this vis helps us comparre statistics over time and relations between them
+eg. srs score & play-off ranking>
 
@@
 
+\section{Possible improvements}
+As with most projects, the result is never 'finished', meaning that there are always improvements
+possible but due to timing or resource constraints the haven't all been implemented. In what 
+follows, we suggest a number of features that might improve the visualization:
+\begin{itemize}
+    \item Smooth transitions between states could improve the user experience. For example when changing
+    year, a transformation of the arrow in the statistics view would help the user see a difference between years.
+    At the moment the arrows change at the blink of an eye, making it difficult for us humans with a short term memory
+    to make the comparison.
+    \item To further support our short term memory we could also 'save' the previous state. Eg. in the field view we could
+    visualize ghost shirts to indicate which players left the team or position and highlight players that joined the team. 
+    This would help the user see how the team changes over time.
+    \item At the moment no attention has been payed to the process & provenance of the exploration. A user thus needs to 
+    memorize his action to be able to share or reproduce his actions. This could be improved by adding URL parameters. This
+    way a user at least can share or save a state of an analysis. One step further is to provide a play-feature. The visitor could 
+    them share his exploration and replay it. A simple implemention would ask the visitor for a start date and an end date and
+    would then 'play' the changes of a particular team and characteristic over time.
+\end{itemize}
+
 \subsection{Technology}
-To create the visualization, the d3 javascript framework~\cite{d3.js} was used in combination with html5~\cite{html5} and jquery~\cite{jquery}. This allows easy
+To create the visualization, the d3 javascript framework~\cite{d3.js} was used in combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3 is based on HTML, CSS, Javascript and SVG.
+Next to that,the use of D3 was a requirement for the project. This allows easy
 access to the visualization as most modern browsers are capable of handling
-these technologies.
+these webstandards. 
 The choice not to use the d3 framework for the entire visualization was made to
@@
 \begin{itemize}
-    \item Find a use case as soon as possible to build your visualization around
+    \item Find a story to tell with your visualization
     \item Each separate component should be a sufficient informative visualization on itself
@@
 
-\subsection{A good use case}
+\subsection{Story telling}
 At the beginning of this project we were able to design the 3 separate views 
@@
 is due to the attraction of talented people or by utilizing available talent within the team.
+Rfr to possible improvements.
+Being able to visualize this story, enabled us to support visistors in searching their own stories.
+They are able to explore the data for patters.
+On top of that, we provide the insight on how the changes were impacted by changes in teamplayers.
 

### 2016-04-23 teamstats.js
Some visual tweaks

@@ -36,15 +36,17 @@ function drawTs(data, team, year, div, x, y, width, height) {
         .range([singleHeight, 0])
         .domain([60, 140]);
 
-    var bases = [{ name: 'SRS', dataSelector : function(d) { return d.team.srs; }, scaler: srsScale}
-            ,{ name: 'League Rank', dataSelector: function(d) {return d.team.leaguerank; }, scaler: rankScale}
-            ,{ name: 'Playoff Rank', dataSelector: function(d) {if (d.team.playoffrank != undefined) {return d.team.playoffrank;} else { return 6;} }, scaler: poRankScale}
-            ,{ name: 'Average Age', dataSelector: function(d) {return d.team.misc.Age;}, scaler: ageScale}
-            ,{ name: 'Audience', dataSelector: function(d) {return d.team.misc.Attendance;}, scaler: audienceScale}
-            ,{ name: 'Field goal %', dataSelector: function(d) {return d.team.info['FG%'];}, scaler: fgScale}
-            ,{ name: 'Points/Game', dataSelector: function(d) {return d.team.info['PTS/G'];}, scaler: pointScale}
-            ,{ name: 'Opponent Points/Game', dataSelector: function(d) {return d.team.opponent['PTS/G'];}, scaler: pointScale}
+    var bases = [{ name: 'Field goal %', dataSelector: function(d) {return d.team.info['FG%'];}, scaler: fgScale, type: "line"}
+            ,{ name: 'League Rank', dataSelector: function(d) {return d.team.leaguerank; }, scaler: rankScale, type: "line"}
+            ,{ name: 'Playoff Rank', dataSelector: function(d) {if (d.team.playoffrank != undefined) {return d.team.playoffrank;} else { return 6;} }, scaler: poRankScale, type: "line"}
+            ,{ name: 'Average Age', dataSelector: function(d) {return d.team.misc.Age;}, scaler: ageScale, type: "line"}
+            ,{ name: 'Audience', dataSelector: function(d) {return d.team.misc.Attendance;}, scaler: audienceScale, type: "area"}
+            ,{ name: 'SRS', dataSelector : function(d) { return d.team.srs; }, scaler: srsScale, type: "area"}
+            ,{ name: 'Points/Game', dataSelector: function(d) {return d.team.info['PTS/G'];}, scaler: pointScale, type: "area"}
[…]

### 2016-04-23 zoomview.js
logo's

@@ -202,6 +202,20 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
         .attr("id", "arrowLeft2")
         .attr("class", "arrowOut");
 
+    var teamName = teamWanted.replace(/\s+/g, '');
+    defs.append('pattern')
+        .attr('id', function(d) { return (teamName+"logo");}) // just create a unique id (id comes from the json)
+        .attr('patternContentUnits', 'objectBoundingBox')
+        .attr('width', 1)
+        .attr('height', 1)
+        .append("svg:image")
+        .attr("xlink:xlink:href", function(d) { return ("./teamlogos/" + teamName + ".png");})
+        .attr("height", 0.8)
+        .attr("width", 0.8)
+        .attr("x", 0.1)
+        .attr("y", 0.1)
+        .attr("preserveAspectRatio", "xMidYMid meet");
+
 
     function drawShirt(xPos, yPos, name, number, team) {
         team = team.replace(/\s+/g, '');
@@ -305,7 +319,10 @@ function drawTransfers(inData, teamWanted, yearWanted, arrowVariable, svg, xpos,
         .attr("class", "teambubblezoom")
         .attr("cx", (chartWidth/2 + margin.left/2 + rad)) //arrow heigth is half of the width
         .attr("cy", (h/2))
-        .attr("r", rad);
+        .attr("r", rad)
+        //.attr("overflow", "hidden")
+        .style("fill", function(d) { return ("url(#" + teamName + "logo)");});
+    //TODO als logo er ni instaat een placeholder?
 
 
     var arrowIncoming = chart.append('g')

### 2016-04-24 bubbleview.js
Update from remote

@@ -1,5 +1,4 @@
 function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
-		console.log(dataInput);
 		
 		var playOffs = dataInput[0].playoffs;
 		var dataInput = dataInput[0].teams;
@@ -115,16 +114,29 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			else
 				return "black";
 		}
+
+		svg.append("defs")
+   			.append("pattern")
+   			.attr("id", "atlantaHawks")
+   			.attr('width', 1)
+     		.attr('height', 1)
+   			.append("svg:image")
+   				.attr("xlink:href", "logos/atlanta hawks.gif")
+   				.attr('width', 1)
+     			.attr('height', 1)
+   			;
    		
    		var tip = d3.tip()
   					.attr('class', 'd3-tip')
-  					.offset([-10, 0])
   					.html(function(d) {
     					return  "<p><span style='color:orange'>" + d["team"] + "</span> <\p>" +
     						"SRS: <span style='color:red'>" + d["srs"] + "</span> </br>" + 
     						"League standings: <span style='color:red'>" + d["leaguerank"] + "</span> </br>" + 
     						"Play off result: <span style='color:red'>" + d["playoffrank"] + "</span>";
   						})
+  					.style('top', 0)
+  					.style('right', 0)
+
   		
 		
         var chart = svg.append("g")			//Append one div to the selected div in which we will construct the visualisation. This is done to separate mutliple visualisations..
@@ -185,9 +197,10 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			    			.attr("cx", function(data){					// Separate position per region. Create more dynamic!
 								return xPosition(data[outlineVariable], data[id]);
 			    					})
-			    			.style("fill", function(data) { 
-			    					return rgbcolor(data[strokeVariable], data["region"]);
-			    					})	//Separate fill color per region. Create more dynamic!			    			
+			    			.attr("fill", "url(#atlantaHawks)")
+			    				//function(data) { 
+			    				//	return rgbcolor(data[strokeVariable], data["region"]);
+			    				//	})	//Separate fill color per region. Create more dynamic!			    			
 			    			.style("stroke-width", 5)
 			    			.attr("stroke", function(data) { 					// Color stroke based on strokeVariable.
 			    				return strokeColor(data[strokeVariable], data.region)
[…]

### 2016-04-24 bubbleview.js
update logo's

@@ -5,7 +5,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 		var dataInput = dataInput[0].teams;
 		var east = dataInput.filter(function(data) { return data.region == "east"});
 		var west = dataInput.filter(function(data) { return data.region == "west"});
-		
+
 		//var height = 600,
 			//width = document.body.clientWidth*0.8;
 		
@@ -115,6 +115,24 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			else
 				return "black";
 		}
+
+		//var teamName = (teamWanted.replace(/\s+/g, '')).toLowerCase();
+		var node = svg.selectAll('node')
+				.data(dataInput);
+		var defs = node.enter().append('defs');
+		defs.append('pattern')
+				.attr('id', function(data) { return (data[id].split(" ").join("_")+"logo");}) // just create a unique id (id comes from the json)
+				.attr('patternContentUnits', 'objectBoundingBox')
+				.attr('width', 1)
+				.attr('height', 1)
+				.append("svg:image")
+				.attr("xlink:xlink:href", function(data) { 
+					return ("./teamlogos/" + ((data[id].replace(/\s+/g, '')).toLowerCase()) + ".png");})
+				.attr("height", 0.8)
+				.attr("width", 0.8)
+				.attr("x", 0.1)
+				.attr("y", 0.1)
+				.attr("preserveAspectRatio", "xMidYMid meet");
    		
    		var tip = d3.tip()
   					.attr('class', 'd3-tip')
@@ -185,9 +203,11 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			    			.attr("cx", function(data){					// Separate position per region. Create more dynamic!
 								return xPosition(data[outlineVariable], data[id]);
 			    					})
-			    			.style("fill", function(data) { 
-			    					return rgbcolor(data[strokeVariable], data["region"]);
-			    					})	//Separate fill color per region. Create more dynamic!			    			
+			    			//.style("fill", function(data) { 
+			    			//		return rgbcolor(data[strokeVariable], data["region"]);
+			    			//		})	//Separate fill color per region. Create more dynamic!	
+			    			.style("fill", function(data) { console.log("url(#" + data[id].split(" ").join("_") + "logo)")
+			    				return ("url(#" + data[id] + "logo)");})		    			
 			    			.style("stroke-width", 5)
[…]

### 2016-04-24 paper/paper.tex

@@
 The visualization's goal can be best summarized by the following sentence: ``Why
-did they win?'' The visualization focuses on exposing relations between team
+did a certain NBA-team win the Championship?'' The visualization focuses on exposing relations between team
 performance and their player roster over several years. It allows users to find
@@
 visualization allows exploration of NBA data by lay persons.  More specifically,
-the visualization does not try to offer premade explanations for phenomenons
+the visualization does not focus on premade explanations for phenomenons
 visible in the NBA data. By providing easy and intuitive access to the data, the
@@
 explanations. 
+< We also give basic insights in why a statistic changes (arrow views)
+we try to help them explore patters and trigger them for questions
+Do we create awareness?
+>
 
@@
 <Tell something more about the data in terms of tall/wide data
+Mention that it's a time series
 Tell something more about pre-processing steps (why?, did we change it over time?)
@@
 statistics= details on demand
+Talk about th organization of the 3 views + the choice for a tiled approach
+- Each page is a 'tile' that we force with fixed scrolling
+- Use of the keyboard is possible (space bar to scroll + arrows, arrows for timeline)
 >
@@
 didn't encouter a similar view when searching for alternatives to visualize a 
-play-off or championship. 
+play-off or championship. A new visualization has the advantage that it will 
+trigger persons as well.
+<Talk about the importance of estethics seen in the lesson>
 Our initial bubble view was too compact. It connected teams that played against
@@
 \end{itemize}
+
+\subsection{Improving the result}
+Talk about how we improved the data-ink ration in the end..
 
@@
     would then 'play' the changes of a particular team and characteristic over time.
+    \item We could support users further in exploring for patterns by focussing more on comparisons. We could integrate a view 
+    where a visitor is able to compare two teams in the same year, or in two different years.
 \end{itemize}
@@
 We could have prepared our data in a format that was more compatible with out of the box examples, but this would have had too much of an impact on our result.
+<Talk about redesign>
 Next to that, although we've tried to write our code as general as possible, we're not convinced this could be shared in such a way that other people building a visualization could easily re-use our work to visualize other sports.

### 2016-04-24 bubbleview.js
update legend

@@ -1,6 +1,78 @@
+function drawLegend(){
+		var legende = d3.select("#legende").append('svg')
+							.attr("width", '500px')
+					
+		legende.append('circle')
+					.attr("r", 10)
+					.attr("cx", 30)
+					.attr("cy", 30)
+					.style("fill", d3.rgb(255,255,255))		    			
+			    	.style("stroke-width", 5)
+			    	.attr("stroke", 'gold');
+
+		legende.append('text')
+					.attr('dx', 50)
+					.attr('dy', 33)
+					.text('Team won gold medal')
+
+		legende.append('circle')
+					.attr("r", 10)
+					.attr("cx", 30)
+					.attr("cy", 60)
+					.style("fill", d3.rgb(255,255,255))		    			
+			    	.style("stroke-width", 5)
+			    	.attr("stroke", 'silver');
+
+		legende.append('text')
+					.attr('dx', 50)
+					.attr('dy', 63)
+					.text('Team won silver medal')
+
+		legende.append('circle')
+					.attr("r", 10)
+					.attr("cx", 30)
+					.attr("cy", 90)
+					.style("fill", d3.rgb(255,255,255))		    			
+			    	.style("stroke-width", 5)
+			    	.attr("stroke", '#D3782F');
+
+		legende.append('text')
+					.attr('dx', 50)
+					.attr('dy', 93)
+					.text('Team won bronze medal in conference')
+
+		legende.append('circle')
+					.attr("r", 10)
+					.attr("cx", 250)
+					.attr("cy", 30)
+					.style("fill", d3.rgb(255,255,255))		    			
+			    	.style("stroke-width", 5)
+			    	.attr("stroke", 'steelblue');
+
+		legende.append('text')
+					.attr('dx', 270)
+					.attr('dy', 33)
+					.text('Team part of east conference')
+
+		legende.append('circle')
+					.attr("r", 10)
+					.attr("cx", 250)
+					.attr("cy", 60)
+					.style("fill", d3.rgb(255,255,255))		    			
+			    	.style("stroke-width", 5)
+			    	.attr("stroke", 'red');
+
+		legende.append('text')
+					.attr('dx', 270)
+					.attr('dy', 63)
+					.text('Team part of west conference')
+}
+
 function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
-		//console.log(dataInput);
-		
+		//Legende visualiseren
+
+
+
[…]

### 2016-04-24 bubbleview.js
update bubbleView layout

@@ -69,17 +69,13 @@ function drawLegend(){
 }
 
 function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
-		//Legende visualiseren
 
-
+		height = height - 100;
 
 		var playOffs = dataInput[0].playoffs;
 		var dataInput = dataInput[0].teams;
 		var east = dataInput.filter(function(data) { return data.region == "east"});
 		var west = dataInput.filter(function(data) { return data.region == "west"});
-
-		//var height = 600,
-			//width = document.body.clientWidth*0.8;
 		
 		var yMaxData = d3.max(dataInput, function(data) {return parseInt(data[outlineVariable]);});
 		
@@ -281,7 +277,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			    			//.style("fill", function(data) { 
 			    			//		return rgbcolor(data[strokeVariable], data["region"]);
 			    			//		})	//Separate fill color per region. Create more dynamic!	
-			    			.style("fill", function(data) { console.log("url(#" + data[id].split(" ").join("_") + "logo)")
+			    			.style("fill", function(data) { //console.log("url(#" + data[id].split(" ").join("_") + "logo)")
 			    				return ("url(#" + data[id] + "logo)");})		    			
 			    			.style("stroke-width", 3)
 			    			.attr("stroke", function(data) { 					// Color stroke based on strokeVariable.
@@ -368,46 +364,124 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 									else d3.select("#"+visibleClass).style('fill','orange');
                             });
 
-      var links = [];
+       var layoutDict = {
+		      Champion: {
+		            first:      [ xPos + width/3,   yPos + 20+(2)*(height/(yMaxData*2+2))],
+		            second:     [ xPos + width/5,   yPos + 20+(4)*(height/(yMaxData*2+2))],
+		            third:      [ xPos + width/9,   yPos + 20+(6)*(height/(yMaxData*2+2))]
+		      },
+		      Second: {
+		            first:      [ xPos + width*4/5,   yPos + 20+(4)*(height/(yMaxData*2+2))],
+		            second:     [ xPos + width*8/9,   yPos + 20+(6)*(height/(yMaxData*2+2))]
+		      },
+		      ThirdEast: {
[…]

### 2016-04-24 bubbleview.js
update bubbleView layout

@@ -472,7 +472,9 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
                      				else if (winner.playoffrank == 2 && loser.playoffrank == 4) 
                      					return {	"x": layoutDict.Second.second[0],
                      								"y": layoutDict.Second.second[1]	}
-
+                     				else if (winner.playoffrank == 2 && loser.playoffrank == 5) 
+                     					return {	"x": layoutDict.Second.second[0],
+                     								"y": layoutDict.Second.second[1]	}
             
       								else 
       									return  {	"x": document.getElementById(winner.team.split(" ").join("_")).cx.animVal.value, 

### 2016-04-25 bubbleview.js
update bubbleView layout

@@ -52,7 +52,7 @@ function drawLegend(){
 		legende.append('text')
 					.attr('dx', 270)
 					.attr('dy', 33)
-					.text('Team part of east conference')
+					.text('Team part of west conference')
 
 		legende.append('circle')
 					.attr("r", 10)
@@ -65,8 +65,23 @@ function drawLegend(){
 		legende.append('text')
 					.attr('dx', 270)
 					.attr('dy', 63)
-					.text('Team part of west conference')
+					.text('Team part of east conference')
 }
+
+function updateTeamInfo(team){
+		var teamInfo = d3.select("#teamInfo").select("text").remove();
+		teamInfo = d3.select("#teamInfo")
+							.attr("width", '500px')
+							.attr('class', 'd3-tip')
+
+		teamInfo.append('text')
+					.attr("dx", '20px').attr("dy", '20px')
+  					.html("<p><span style='color:orange'>" + team["team"] + "</span> <\p>" +
+    						"SRS: <span style='color:red'>" + team["srs"] + "</span> </br>" + 
+    						"League standings: <span style='color:red'>" + team["leaguerank"] + "</span> </br>" + 
+    						"Play off result: <span style='color:red'>" + team["playoffrank"] + "</span>")
+}
+   		
 
 function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
 
@@ -165,9 +180,11 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 		  };
 		}
 		
-		function radius(radiusVariable) {
-			return (11 + radiusVariable)*1.8 * width/1000;				// veranderen obv D3.scale
-		}
+		var radius = d3.scale.sqrt()
+				.domain([d3.min(dataInput, function(data) {return parseInt(data[radiusVariable]);}),
+					d3.max(dataInput, function(data) {return parseInt(data[radiusVariable]);})	])
+				.range([width/60,width/20])
+
 		
 		function strokeColor(strokeVariable, groupVariable){
 			if (strokeVariable == 1)
@@ -200,28 +217,12 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 				.attr("width", 0.8)
 				.attr("x", 0.1)
 				.attr("y", 0.1)
-				.attr("preserveAspectRatio", "xMidYMid meet");
-   		
-   		var tip = d3.tip()
[…]

### 2016-04-25 bubbleview.js
selection highligted

@@ -260,7 +260,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 							.enter()
 							.append("g");
 
-		
+		var visibleBoolean = false;
 
 		circleWest.append("circle")
 							.attr("class","circle teambubble")
@@ -286,7 +286,17 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
    									d3.selectAll("."+name).style("visibility", "visible");
    									visibleClass = name;
    									visibleBoolean = !visibleBoolean;
-   									d3.select(this).style('fill','orange');
+   									d3.select(this)
+                                        .style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
+   									if (!visibleBoolean)	
+   										d3.select(this).style('fill','orange'); 
+   									team = data.team;
+				                    $("#bubbleheader h1").text(data.team);
+				                    if (view === "bubble") {
+				                        scrollMe("zoom");
+				                    }
+				                    drawFullTeamChange(window.year);
+				                    console.log(visibleBoolean)
       						})
    							.on('mouseover', function(data) {
    									d3.select(this).style('fill','orange');
@@ -299,14 +309,14 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 								})
       						.on('mouseout', function(data) {
       								d3.select("#teamInfo").select("text").remove();
+      								updateTeamInfo(team)
                                     d3.select(this)
                                         .style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
       								if (visibleBoolean){
       									d3.selectAll(".arc").style("visibility", "visible");
-      									d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
-									}
-
+									} else d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
                             });
+      	
 
 		circleEast.append("circle")
 							.attr("class","circle teambubble")
@@ -331,10 +341,20 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
       								d3.selectAll(".arc").style("visibility", "hidden");
    									var name = data.team.split(" ").join(".");
    									d3.selectAll("."+name).style("visibility", "visible");
+   									d3.select(this)
+										.style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
    									visibleClass = name;
    									visibleBoolean = !visibleBoolean;
[…]

### 2016-04-25 bubbleview.js
selection highligted

@@ -73,8 +73,10 @@ function updateTeamInfo(team){
 		teamInfo = d3.select("#teamInfo")
 							.attr("width", '500px')
 							.attr('class', 'd3-tip')
+		console.log(team)
+		
 
-		teamInfo.append('text')
+		d3.select("#teamInfo").append('text')
 					.attr("dx", '20px').attr("dy", '20px')
   					.html("<p><span style='color:orange'>" + team["team"] + "</span> <\p>" +
     						"SRS: <span style='color:red'>" + team["srs"] + "</span> </br>" + 
@@ -367,7 +369,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 								})
       						.on('mouseout', function(data) {
       								d3.select("#teamInfo").select("text").remove();
-      								updateTeamInfo(team)
+      								updateTeamInfo(data)
       								d3.select(this)
 										.style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
       								if (visibleBoolean){

### 2016-04-25 bubbleview.js
update global status

@@ -283,6 +283,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			    				return strokeColor(data[strokeVariable], data.region)
 			    					})
       						.on('click', function(data){
+      								team = data;
       								d3.selectAll(".arc").style("visibility", "hidden");
    									var name = data.team.split(" ").join(".");
    									d3.selectAll("."+name).style("visibility", "visible");
@@ -292,7 +293,6 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
                                         .style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
    									if (!visibleBoolean)	
    										d3.select(this).style('fill','orange'); 
-   									team = data.team;
 				                    $("#bubbleheader h1").text(data.team);
 				                    if (view === "bubble") {
 				                        scrollMe("zoom");
@@ -305,7 +305,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
    									updateTeamInfo(data)
    									if (visibleBoolean){
    										d3.selectAll(".arc").style("visibility", "hidden");
-   										var name = data.team.split(" ").join(".");
+   										name = data.team.split(" ").join(".");
    										d3.selectAll("."+name).style("visibility", "visible");	
    									}
 								})
@@ -340,6 +340,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			    				return strokeColor(data[strokeVariable], data.region)
 			    					})
       						.on('click', function(data){
+      								team = data;
       								d3.selectAll(".arc").style("visibility", "hidden");
    									var name = data.team.split(" ").join(".");
    									d3.selectAll("."+name).style("visibility", "visible");
@@ -349,7 +350,6 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
    									visibleBoolean = !visibleBoolean;
 									if (!visibleBoolean)	
    										d3.select(this).style('fill','orange'); 
-   									team = data.team;
 				                    $("#bubbleheader h1").text(data.team);
 				                    if (view === "bubble") {
[…]

### 2016-04-25 bubbleview.js
update global status

@@ -72,9 +72,7 @@ function updateTeamInfo(team){
 		var teamInfo = d3.select("#teamInfo").select("text").remove();
 		teamInfo = d3.select("#teamInfo")
 							.attr("width", '500px')
-							.attr('class', 'd3-tip')
-		console.log(team)
-		
+							.attr('class', 'd3-tip')		
 
 		d3.select("#teamInfo").append('text')
 					.attr("dx", '20px').attr("dy", '20px')
@@ -298,7 +296,6 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 				                        scrollMe("zoom");
 				                    }
 				                    drawFullTeamChange(window.year);
-				                    console.log(visibleBoolean)
       						})
    							.on('mouseover', function(data) {
    									d3.select(this).style('fill','orange');
@@ -311,12 +308,15 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 								})
       						.on('mouseout', function(data) {
       								d3.select("#teamInfo").select("text").remove();
-      								updateTeamInfo(team)
                                     d3.select(this)
                                         .style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
       								if (visibleBoolean){
       									d3.selectAll(".arc").style("visibility", "visible");
-									} else d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
+									} else {
+										d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
+										updateTeamInfo(team)	
+									}
+									
                             });
       	
 
@@ -369,13 +369,14 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 								})
       						.on('mouseout', function(data) {
       								d3.select("#teamInfo").select("text").remove();
-      								updateTeamInfo(data)
       								d3.select(this)
 										.style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
       								if (visibleBoolean){
-      									d3.selectAll(".arc").style("visibility", "visible");	
-									}	else d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
-
+      									d3.selectAll(".arc").style("visibility", "visible");
+      								} else {
+										d3.select("#"+visibleClass.split(".").join("_")).style('fill','orange');
+										updateTeamInfo(team)	
+									}
                             });
 				
       	var finale = playOffs.filter(function(d) { return d.game == 'Finals'})[0]

### 2016-04-25 bubbleview.js
update global status

@@ -3,11 +3,11 @@ function drawLegend(){
 							.attr("width", '500px')
 					
 		legende.append('circle')
-					.attr("r", 10)
+					.attr("r", 5)
 					.attr("cx", 30)
 					.attr("cy", 30)
 					.style("fill", d3.rgb(255,255,255))		    			
-			    	.style("stroke-width", 5)
+			    	.style("stroke-width", 3)
 			    	.attr("stroke", 'gold');
 
 		legende.append('text')
@@ -16,11 +16,11 @@ function drawLegend(){
 					.text('Team won gold medal')
 
 		legende.append('circle')
-					.attr("r", 10)
+					.attr("r", 5)
 					.attr("cx", 30)
 					.attr("cy", 60)
 					.style("fill", d3.rgb(255,255,255))		    			
-			    	.style("stroke-width", 5)
+			    	.style("stroke-width", 3)
 			    	.attr("stroke", 'silver');
 
 		legende.append('text')
@@ -29,11 +29,11 @@ function drawLegend(){
 					.text('Team won silver medal')
 
 		legende.append('circle')
-					.attr("r", 10)
+					.attr("r", 5)
 					.attr("cx", 30)
 					.attr("cy", 90)
 					.style("fill", d3.rgb(255,255,255))		    			
-			    	.style("stroke-width", 5)
+			    	.style("stroke-width", 3)
 			    	.attr("stroke", '#D3782F');
 
 		legende.append('text')
@@ -42,11 +42,11 @@ function drawLegend(){
 					.text('Team won bronze medal in conference')
 
 		legende.append('circle')
-					.attr("r", 10)
+					.attr("r", 5)
 					.attr("cx", 250)
 					.attr("cy", 30)
 					.style("fill", d3.rgb(255,255,255))		    			
-			    	.style("stroke-width", 5)
+			    	.style("stroke-width", 3)
 			    	.attr("stroke", 'steelblue');
 
 		legende.append('text')
@@ -55,11 +55,11 @@ function drawLegend(){
 					.text('Team part of west conference')
 
 		legende.append('circle')
-					.attr("r", 10)
+					.attr("r", 5)
 					.attr("cx", 250)
 					.attr("cy", 60)
 					.style("fill", d3.rgb(255,255,255))		    			
-			    	.style("stroke-width", 5)
+			    	.style("stroke-width", 3)
 			    	.attr("stroke", 'red');
 
 		legende.append('text')
[…]

### 2016-04-25 bubbleview.js
update bubble layout

@@ -85,7 +85,7 @@ function updateTeamInfo(team){
 
 function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
 
-		height = height - 100;
+		height = height;
 
 		var playOffs = dataInput[0].playoffs;
 		var dataInput = dataInput[0].teams;

### 2016-04-25 bubbleview.js
logo tooltip

@@ -68,18 +68,37 @@ function drawLegend(){
 					.text('Team part of east conference')
 }
 
-function updateTeamInfo(team){
+function removeTeamInfo(){
 		var teamInfo = d3.select("#teamInfo").select("text").remove();
+		var teamInfo = d3.select("#teamInfo").select("svg").remove();
+}
+
+function updateTeamInfo(team){
+		removeTeamInfo()
 		teamInfo = d3.select("#teamInfo")
 							.attr("width", '500px')
 							.attr('class', 'd3-tip')		
 
 		d3.select("#teamInfo").append('text')
-					.attr("dx", '20px').attr("dy", '20px')
+					.attr("dx", '20px')
+					.attr("dy", '20px')
+					.attr("width", '200px')
+					.attr('float','left')
+					.attr('class', 'textTeamInfo')
   					.html("<p><span style='color:orange'>" + team["team"] + "</span> <\p>" +
     						"SRS: <span style='color:red'>" + team["srs"] + "</span> </br>" + 
     						"League standings: <span style='color:red'>" + team["leaguerank"] + "</span> </br>" + 
     						"Play off result: <span style='color:red'>" + team["playoffrank"] + "</span>")
+
+		d3.select("#teamInfo").append('svg')
+					.attr("width", 100)
+    				.attr("height", 100)
+				.append('circle')
+					.style('fill',("url(#" + team["team"].split(" ").join("_") + "logo)"))
+					.attr("r", width/20)
+	    			.attr("cy", 50)
+	    			.attr("cx", 50)
+	    			.attr('float','left')
 }
    		
 
@@ -309,7 +328,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
    									}
 								})
       						.on('mouseout', function(data) {
-      								d3.select("#teamInfo").select("text").remove();
+      								removeTeamInfo()
                                     d3.select(this)
                                         .style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
       								if (visibleBoolean){
@@ -370,7 +389,7 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
    									}
 								})
       						.on('mouseout', function(data) {
-      								d3.select("#teamInfo").select("text").remove();
+      								removeTeamInfo()
                                     d3.select(this)
                                         .style('fill',("url(#" + data[id].split(" ").join("_") + "logo)"));
       								if (visibleBoolean){

### 2016-04-25 bubbleview.js
Some changes

@@ -83,7 +83,7 @@ function updateTeamInfo(team){
 }
    		
 
-function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height){
+function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable, id, svg, xPos, yPos, width, height, colors){
 
 		height = height - 100;
 
@@ -275,10 +275,21 @@ function drawCircles(dataInput, radiusVariable, strokeVariable, outlineVariable,
 			    			.attr("cx", function(data){					// Separate position per region. Create more dynamic!
 								return xPosition(data[outlineVariable], data[id]);
 			    					})
-			    			.style("fill", function(data) { return ("url(#" + data[id].split(" ").join("_") + "logo)");})		    			
+                            .style("fill", function(data) { 
+                                if (colors) { 
+                                    return ShirtColors[data[id].replace(/\s+/g, '')].shirt;
+                                } else {
+                                    return ("url(#" + data[id].split(" ").join("_") + "logo)");
+                                }
+                            })
 			    			.style("stroke-width", 3)
-			    			.attr("stroke", function(data) { 					// Color stroke based on strokeVariable.
-			    				return strokeColor(data[strokeVariable], data.region)
+			    			.attr("stroke", function(data) {
+                                return ShirtColors[data[id].replace(/\s+/g, '')].edge;
+                                if (colors) { 
+                                    return ShirtColors[data[id].replace(/\s+/g, '')].edge;
+                                } else {
+                                    return strokeColor(data[strokeVariable], data.region);
+                                }
 			    					})
       						.on('click', function(data){
       								d3.selectAll(".arc").style("visibility", "hidden");

### 2016-04-26 bubbleview.js
Some more changes

@@ -1,6 +1,6 @@
 function drawLegend(){
 		var legende = d3.select("#legende").append('svg')
-							.attr("width", '500px')
+							.attr("width", '300px')
 					
 		legende.append('circle')
 					.attr("r", 5)
@@ -11,22 +11,35 @@ function drawLegend(){
 			    	.attr("stroke", 'gold');
 
 		legende.append('text')
-					.attr('dx', 50)
-					.attr('dy', 33)
+					.attr('dx', 40)
+					.attr('dy', 34)
 					.text('Team won gold medal')
 
 		legende.append('circle')
 					.attr("r", 5)
 					.attr("cx", 30)
-					.attr("cy", 60)
+					.attr("cy", 50)
 					.style("fill", d3.rgb(255,255,255))		    			
 			    	.style("stroke-width", 3)
 			    	.attr("stroke", 'silver');
 
 		legende.append('text')
-					.attr('dx', 50)
-					.attr('dy', 63)
+					.attr('dx', 40)
+					.attr('dy', 54)
 					.text('Team won silver medal')
+
+		legende.append('circle')
+					.attr("r", 5)
+					.attr("cx", 30)
+					.attr("cy", 70)
+					.style("fill", d3.rgb(255,255,255))		    			
+			    	.style("stroke-width", 3)
+			    	.attr("stroke", '#D3782F');
+
+		legende.append('text')
+					.attr('dx', 40)
+					.attr('dy', 74)
+					.text('Team won bronze medal in conference')
 
 		legende.append('circle')
 					.attr("r", 5)
@@ -34,71 +47,43 @@ function drawLegend(){
 					.attr("cy", 90)
 					.style("fill", d3.rgb(255,255,255))		    			
 			    	.style("stroke-width", 3)
-			    	.attr("stroke", '#D3782F');
-
-		legende.append('text')
-					.attr('dx', 50)
-					.attr('dy', 93)
-					.text('Team won bronze medal in conference')
-
-		legende.append('circle')
-					.attr("r", 5)
-					.attr("cx", 250)
-					.attr("cy", 30)
-					.style("fill", d3.rgb(255,255,255))		    			
-			    	.style("stroke-width", 3)
 			    	.attr("stroke", 'steelblue');
 
 		legende.append('text')
-					.attr('dx', 270)
-					.attr('dy', 33)
+					.attr('dx', 40)
+					.attr('dy', 93)
 					.text('Team part of west conference')
 
[…]

### 2016-04-26 teamstats.js

@@ -134,14 +134,14 @@ function drawTs(data, team, year, div, x, y, width, height) {
         highlightYear = year;
         fixcyline();
         drawTeamChange(year);
-        d3.select("#bubbleheader").select("h2").text(year);
+        updateTeamInfo(data.find(function (d) { return d.year == highlightYear; }).teams.find(function (d) {return d.team === window.team.team;}), highlightYear);
     });
 
     svg.on("mousemove", function() {
          highlightYear = Math.round(x.invert(d3.mouse(this)[0]));
          fixcyline();
          drawTeamChange(highlightYear);
-         d3.select("#bubbleheader").select("h2").text(highlightYear);
+        updateTeamInfo(data.find(function (d) { return d.year == highlightYear; }).teams.find(function (d) {return d.team === window.team.team;}), highlightYear);
     });
 
     svg.on("click", function() {

### 2016-05-16 paper/paper.tex
Beetje gewerkt aan paper (vooral bubbleview tot hiertoe)
Kan zijn da er wa van bubbleview bij relevante literature mag

@@
 \begin{abstract}
-    This paper details the conceptualization and implementation of an
-    interactive visualization of data about NBA seasons including teams, game
-    results and player statistics. The d3.js framework was used to develop this
-    visualization.
+    In this paper we describe an interactive visualization of the NBA competition over the last 30 years.
+With three different views, we give the opportunity for the user to explore various kinds of data about NBA seasons including standings, team statistics, transfers and player statistics.
+The d3.js framework was used to develop this visualization.
     %TODO add results
+
 \end{abstract}
@@
 \section{Introduction}
+
+The National Basketball Association (NBA) is the most famous and perhaps the best professional basketball league in the world.
+Like in every other sports branch, a lot of data and statistics about players, teams, games and seasons are available, mainly in large tables.
+Despite these tables contains a lot of information, it is very hard if not impossible, to interpret, compare efficiently and to gather insights in it.
+In our work we try to present this data in a whole different way.
+Therefore we have focused ourselves on some parts and not the entire available data.
 
@@
 
+%Related work/References
+
+%http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado_MultipleViews_AVI00.pdf
+%http://lemonchiu.github.io/NBA-Visualization/
+%http://peterbeshai.com/buckets/app
+%https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/
+%http://projects.fivethirtyeight.com/complete-history-of-the-nba/#warriors
+%http://www.theguardian.com/news/datablog/interactive/2013/jun/03/premier-league-season-visualised
+%http://gizmodo.com/5927503/all-the-major-sport-competitions-since-1903-condensed-in-beautiful-circular-graphics
+%http://www.nbaplayoffsbracket.com/2016/index.php
+%http://inf.ufrgs.br/~lsguedes/skooth/Publications_&_Works_files/nbavis_final.pdf
+%https://en.wikipedia.org/wiki/National_Basketball_Association
+
 \section{Visualization}\label{sec:visualization}
@@
 \begin{itemize}
-    \item The bubble view: a compact view on the play-offs per season
+    \item The play-off view: a compact view on the play-offs per season
     \item The statistics (zoom) view: a more detailed view on a statistic of a selected team
@@
 
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/playoffviewwithcontext}
+  \caption{The play-off view without a team selected.}~\label{fig:playoffviewnoteam}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/playoffviewteamselected}
+  \caption{The play-off view with a team selected.}~\label{fig:playoffviewteam}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/statisticsview}
+  \caption{The statistics view.}~\label{fig:statisticsview}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/teamview}
+  \caption{The team view.}~\label{fig:teamview}
+\end{figure}
+
 <Explain that this complies with the infovis seeking mantra: Overview, zoom&filter, details on demand
-Bubble = overview
-statistics= details on demand
+play-off view = overview
+statistics view= details on demand
 Talk about th organization of the 3 views + the choice for a tiled approach
@@
 All three are discussed in more detail below. The three views are available to
-the user on a single webpage. The user start on the bubble view and is able to
+the user on a single webpage. The user start on the play-off view and is able to
 navigate to the statistics and team views by scrolling down or clicking on
-teams.
+teams. A different season can be selected with the left and right arrows or by clicking on the timeline. The timeline is a fixed part of the visualization, hence visible in the three different views.
 %The webpage is built in such a way that it gives a hint to the user at the bottom
@@
 
-\subsection{The bubble view}
-This view supposes to inform the user of the results of the NBA play-offs per 
-season. A circle is used to represent a team, the size its SRS score using a '?'
-distribution, a stroke to illustrate it's region or medal won during regional 
-competition and curved lines to inform the user which teams played against each 
-other. The user has a timeline at its disposal to scroll through time. Note that
-the team can be recognized by the logo inside the circle. Our initial desing filled
-the circle with the red or blue depending on the region where the teams compete. 
-However, the shade of colour was darker if their final ranking was higher in their
-region. We've replaced this with the logo's of the teams as we thought this is more
-user friendly and the extra information of the ranking doesn't outweight.
-<picture of the bubble view compared with double elimination bracket>
+\subsection{The play-off view}
+This view informs the user of the roster and end positions in the NBA play-offs for a selected season, in general or for a selected team. The view consists of two parts: 
+\begin{itemize}
+    \item Context section: the top part informs the user about the selected team and shows a legend of the play-off view.
+    \item Bubble section: this section represents the play-offs of the selected season
+\end{itemize}
+The play-off view is shown in Figure~\ref{fig:playoffviewwithcontext}.
+
+\subsubsection{Context section}
+The context section contains a legend that explains the stroke colors used in the bubble section.
+The stroke is gold, silver or bronze if the team ended respectively first, second or third in the regular competition.
+Teams that didn't end in a podium place during the regular competition have a red or blue stroke color indicating the team's region, respectively the western or eastern conference.
+The context section also contains a link to another page (``About'') where the user can get an explanation of the visualization with its different views, can get some background information about the NBA and the terminolgy used, and can check the source of our used data.
+If the user hovers over or selects a team in the bubble section, the context section shows more information about that team, as shown in Figure~\ref{fig:playoffviewteamselected}.
+This includes the Simple Rating System score (SRS), the end position in the play-offs and the end position in the regular competition.
+
+\subsubsection{Bubble section}
+This section presents the roster and end positions of the play-offs for a selected season.
+Every team is respresented by a circle with the logo of the team inside it.
+Our initial desing filled the circle with the logo of the team and colored the circle red or blue depending on the region where the team competed.
+The shade of colour was darker if their final ranking was higher in their region.
+According to the Ranking of Perceptual Tasks~\cite{perceptualranking} this is a good property for ordered values like the regular competition ranking.
+But the combination of the logo's of the teams with the colors made it to heavy and less easy to compare the end ranking of different teams.
+As an alternative we tried the same view without the logo's, but this is inconvenient because you don't know which team is represented by which circle.
+Therefore we've chosen to only use the logo's of the teams as we thought this is more user friendly and the extra information of the ranking doesn't outweight.
+The size of the circle represents the SRS score using an own interpretation of perceptual scaling because the (Flannery) Appearance Compensation is not usable with a domain of both negative and positive values.
+An exponential function with exponent 1.5 was used.
+This way the proportion between the area of a bigger and the area of a smaller circle will be greater than the proportion of values the areas represent.
+The stroke of the circle illustrates the team's region or medal won during the regional competition as explained in the context section.
+The curved lines between teams represent that those teams have encountered each other in the playoffs.
+Each row represents an end position in the play-offs.
+The higher a team is, the further it has come.
+From the top to the bottom the rows represent the winner of the final (place 1), loser of the final (place 2), losers of the semi-finals (place 3), losers of the quarter-finals (place 4), losers of the 1/8-finales (place 5).
+The last row represents the teams that did not make it to the play-offs (place 6).
+A connection is always made between a team that is in a higher row and a team that is in a lower row.
+The highest team of the two teams has won the confrontation, and therefor went at least one round further in the play-offs.
+The two up most teams are the winners of the play-offs in their own conference and played the final.
 A more popular visualization of a play-off is the team double elimination bracket.
-Although this is a more common representation, we wanted to create a more compact
-version by eliminating the recurring representation of a team. We also considered 
-enclosed circles and a sunburst.
-<figure of 2 alternatives https://thetuftersblog.wordpress.com/2016/03/07/team-ranking-play-off-infovis/>
+Although this is a more common representation , and there are some nice desings like in this NBA 2012 play-offs visualisation~\cite{tournamentladder} where a sunburst visualisation is used, we wanted to create a more compact version by eliminating the recurring representation of a team.
+We considered enclosed circles, as used in this visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
 Although the enclosed circles is a compact representation as well, we have continued 
-with our own design as we believe this view is more original and innovative. We 
-didn't encouter a similar view when searching for alternatives to visualize a 
-play-off or championship. A new visualization has the advantage that it will 
-trigger persons as well.
+with our own design as we believe this view is more original and innovative.
+Furthermore we are convinced that our view gives a good illustration of the end positions.
+It's generally assumed that the team that is on top of the ``ladder'' is the winning team.
+We didn't encouter a similar view when searching for alternatives to visualize a play-off or championship.
+A new visualization has the advantage that it will trigger users as well.
+
 <Talk about the importance of estethics seen in the lesson>
-Our initial bubble view was too compact. It connected teams that played against
+Our initial play-off view was too heavy. It connected teams that played against
 each other directly. When searching for alternatives to reflect on our choice we
@@
 <https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/>
-<comparison of original bubble view with white house paths>
+<comparison of original play-off view with white house paths>
 This visualization gives a cleaner view of the competitors for a specific team
@@
 fixated when the users clicks on a team.
-He/she then automatically gets forwarded to the statistics view where the bubble 
+He/she then automatically gets forwarded to the statistics view where the play-off 
 view is still represented in the right upper corner to make sure the user keeps
 an overview. The statistics view will be explained in more detail in the next section.
+
 < Hierarchy = sorting = facilitate trends of winners
-position have a meaning in the bubble view
+position have a meaning in the play-off view
 when selecting circle => zoom/more detail
@@
 
-We considered a map view, but didn't do it because geographical position has no added value
+We considered a map view to just show the teams and there regular competition ranking.
+But this was a view that has no additional value in comparison with the play-off view except for the geographical position, which has no importance for our visualisation.
+Furthermore we've found a visualisation of the NBA with a map view~\cite{mapviewvisualization} similar to our idea, so we decided to not implement this map view.
+
 This visualization helped us discover data quality issues => cut off year 1984
@@
 \subsection{The statistics view}
-The statistics view gives a user a clear overview of how a statistic has been influenced compared to the year before. The view consists of 3 parts:
+The statistics view gives a user a clear overview of how a statistic has been influenced in the selected season compared to the season before.
+The statistics view is showed in Figure~\ref{fig:statisticsview}.
+The view consists of 3 coordinated parts:
 \begin{itemize}
-    \item Context section: the top part should inform the user of his/her
-        context
-    \item Arrow section: hhis section illustrates why a statistic changed
-    \item Small multiples section: hmall multiples of team statistics and how they evolve over time
+    \item Context section: the top part informs the user of the context of the selected team
+    \item Arrow section: this section illustrates why a statistic changed compared to the previous season
+    \item Small multiples section: small multiples of team statistics and how they evolve over time
 \end{itemize}
 
-\subsection{Improving the result}
-Talk about how we improved the data-ink ration in the end..
-
 \subsubsection{Context section}
-The context section should inform the user at all times of the current selections 
-that have been made. The statistics view will always be based on a specific team
-and season that have been selected or that are initated. This should be clear based
-on the textual information in the top left. Next to that, in the top right a small 
-bubble view is shown. It indicates the team selected and how well it performed in 
-the play-offs during that season, as explained in the bubble view section. When changing teams and/or season, this view gets 
-updated as well as the textual information.
-<Explain that the context supports the short term memory of human beings rfr. Monkey Illusion>
+The context section informs the user at all times of the current selection of a team  
+that have been made in the play-off view.
+The statistics view will always be based on a specific team and season that have been selected or that are initated.
+This should be clear based on the textual information in the top left.
+Next to that, in the top right a small play-off view is shown.
+It indicates the team selected and how well it performed in  the play-offs during that season, as explained in the play-off view section.
+When changing teams and/or season, this view gets updated as well as the textual information.
+<Explain that the context supports the short term memory of human beings rfr.
+Monkey Illusion>
 
@@
 This should give a visitor more insight into why a statistic changed over time.
-The selected team is represented by a circle in the same way as explained in the bubble view.
+The selected team is represented by a circle in the same way as explained in the play-off view.
 The change for a statistic is represented by three arrows:
@@
 <Sketch of the small multiples>
+
+\subsection{Improving the result}
+Talk about how we improved the data-ink ration in the end..
 
@@
 Our main problem was to integrate these views as a whole.
-When we created drafts of two views being the bubble view and the statitics view, we suddenly noted patterns.
+When we created drafts of two views being the play-off view and the statitics view, we suddenly noted patterns.
 The Golden State Warriors became NBA champion a couple of years after a great uplift in SRS score.

### 2016-05-18 paper/paper.tex
links voor related work opgezocht

@@
 
+Many visualizations of NBA data are focused on the "shot-chart".  This of individual players, multiple players together or all the teams
+This is almost a standard in visualizations about NBA data where even the code for making short charts is easily found ~\cite{shotchartpython}~\cite{shotchartr}~\cite{shotchartapp}.
+
+This paper (http://inf.ufrgs.br/~lsguedes/skooth/Publications_&_Works_files/nbavis_final.pdf) describes
+
 %Related work/References
 
+%General
+%https://en.wikipedia.org/wiki/National_Basketball_Association
 %http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado_MultipleViews_AVI00.pdf
-%http://lemonchiu.github.io/NBA-Visualization/
+%http://www.interactiondesign.us/courses/2011_AD690/PDFs/Shneiderman_1996.pdf
+%http://workshop.sportvis.com/
+%http://www2.parc.com/istl/groups/uir/publications/items/UIR-1986-02-Mackinlay-TOG-Automating.pdf
+%http://www.cc.gatech.edu/~john.stasko/papers/infovis06.pdf
+
+%Shot charts
 %http://peterbeshai.com/buckets/app
+%http://vorped.com/1-nba/2015-2016/player/6667/emmanuel-mudiay/
+%https://github.com/amatlin/NBAvis
+%https://thedatagame.com.au/2015/09/27/how-to-create-nba-shot-charts-in-r/
+%http://savvastjortjoglou.com/nba-shot-sharts.html?utm_source=Python+Weekly+Newsletter&utm_campaign=5185ff0538-Python_Weekly_Issue_202_July_30_2015&utm_medium=email&utm_term=0_9e26887fc5-5185ff0538-312727397
+%http://www.sloansportsconference.com/wp-content/uploads/2012/02/Goldsberry_Sloan_Submission.pdf
+%http://vis.berkeley.edu/courses/cs294-10-sp10/wiki/images/archive/a/a2/20100511091342!Stephen_Chu_-_Information_Visualization_in_the_NBA_-_The_Shot_Chart_16.pdf
+%http://toddwschneider.com/posts/ballr-interactive-nba-shot-charts-with-r-and-shiny/
+
+%Standings
 %https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/
@@
 %http://www.nbaplayoffsbracket.com/2016/index.php
+%http://nyloncalculus.com/2015/08/07/as-nba-empires-rise-and-fall/
 %http://inf.ufrgs.br/~lsguedes/skooth/Publications_&_Works_files/nbavis_final.pdf
-%https://en.wikipedia.org/wiki/National_Basketball_Association
+%http://research.microsoft.com/pubs/64282/tvcg2007-adaptivitree.pdf
+%http://www.student.kuleuven.be/~s0217391/mume/deliverables/paper.pdf
+
+%Other
+%https://graphics.stanford.edu/wikis/cs448b-10-fall/FP-RuthEric-HeddlestonKate?action=AttachFile&do=get&target=448bPaper.pdf
+%http://www.scribblelive.com/blog/2014/03/20/the-power-of-sports-data-visualizing-passes-between-nba-players-offers-new-game-insights/
+%http://lemonchiu.github.io/NBA-Visualization/
 

### 2016-05-18 paper/paper.tex
Start related work

@@
 Like in every other sports branch, a lot of data and statistics about players, teams, games and seasons are available, mainly in large tables.
-Despite these tables contains a lot of information, it is very hard if not impossible, to interpret, compare efficiently and to gather insights in it.
+Despite these tables contains a lot of information, it is very hard if not impossible, to interpret, compare efficiently and gather insights in it.
 In our work we try to present this data in a whole different way.
@@
 rfr to picture in 1st ppt sl 60>
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/data}
+  \caption{An overview of the data.}~\label{fig:data}
+\end{figure}
+
 \section{Related Work}\label{sec:literature}
 
-Many visualizations of NBA data are focused on the "shot-chart".  This of individual players, multiple players together or all the teams
-This is almost a standard in visualizations about NBA data where even the code for making short charts is easily found ~\cite{shotchartpython}~\cite{shotchartr}~\cite{shotchartapp}.
+On the IEEE Visualization Conference in 2013, the first workshop on sports data visualization was held \cite{ieeevis}.
+There is shown that the area of sport produces large amounts of data, which can be used for building visualizations and making analyses of the data.
+In the past visualizations of sports existed already, but there were not a lot of visualizations, and they were simple in the sense that they just represented the data without trying to give insights or to approach the data on a different way.
+There are some, but limited, sports-specific visualizations, like the basketball shot chart.
+This is a visualization frequentely seen when looking for visualizations about the NBA.
+Some shot chart vizualisations try to be different than others, but the general design and idea stays roughly the same.
+% 2 goeie papers die hierin verwerkt kunnen worden
+%http://www.sloansportsconference.com/wp-content/uploads/2012/02/Goldsberry_Sloan_Submission.pdf
+%http://vis.berkeley.edu/courses/cs294-10-sp10/wiki/images/archive/a/a2/20100511091342!Stephen_Chu_-_Information_Visualization_in_the_NBA_-_The_Shot_Chart_16.pdf
 
-This paper (http://inf.ufrgs.br/~lsguedes/skooth/Publications_&_Works_files/nbavis_final.pdf) describes
+With our visualization we try to make it easier to interpret a collection of data about the NBA, and try to let users find patterns and insights in the data.
+A same kind of incentive was the start for \cite{cox2006sportsvis} who created a visualization to help people discover meaning in statistics generated during sporting events in the case of baseball.
 
 %Related work/References
-
-%General
-%https://en.wikipedia.org/wiki/National_Basketball_Association
-%http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado_MultipleViews_AVI00.pdf
-%http://www.interactiondesign.us/courses/2011_AD690/PDFs/Shneiderman_1996.pdf
-%http://workshop.sportvis.com/
-%http://www2.parc.com/istl/groups/uir/publications/items/UIR-1986-02-Mackinlay-TOG-Automating.pdf
-%http://www.cc.gatech.edu/~john.stasko/papers/infovis06.pdf
 
@@
 %http://savvastjortjoglou.com/nba-shot-sharts.html?utm_source=Python+Weekly+Newsletter&utm_campaign=5185ff0538-Python_Weekly_Issue_202_July_30_2015&utm_medium=email&utm_term=0_9e26887fc5-5185ff0538-312727397
-%http://www.sloansportsconference.com/wp-content/uploads/2012/02/Goldsberry_Sloan_Submission.pdf
-%http://vis.berkeley.edu/courses/cs294-10-sp10/wiki/images/archive/a/a2/20100511091342!Stephen_Chu_-_Information_Visualization_in_the_NBA_-_The_Shot_Chart_16.pdf
 %http://toddwschneider.com/posts/ballr-interactive-nba-shot-charts-with-r-and-shiny/
@@
 - Use of the keyboard is possible (space bar to scroll + arrows, arrows for timeline)
+- This papers are usefulle to refer to
+http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado_MultipleViews_AVI00.pdf
+http://www.interactiondesign.us/courses/2011_AD690/PDFs/Shneiderman_1996.pdf
+http://www2.parc.com/istl/groups/uir/publications/items/UIR-1986-02-Mackinlay-TOG-Automating.pdf
 >
@@
 \printbibliography
+%TODO References should be in alphabetical order and represented by numbers in the text.
 

### 2016-05-20 paper/paper.tex
Little changes
I don't remember

@@
 
-The National Basketball Association (NBA) is the most famous and perhaps the best professional basketball league in the world.
-Like in every other sports branch, a lot of data and statistics about players, teams, games and seasons are available, mainly in large tables.
-Despite these tables contains a lot of information, it is very hard if not impossible, to interpret, compare efficiently and gather insights in it.
-In our work we try to present this data in a whole different way.
-Therefore we have focused ourselves on some parts and not the entire available data.
+The National Basketball Association (NBA) is the most famous and perhaps the
+best professional basketball league in the world. Like in every other sports
+branch, a lot of data and statistics about players, teams, games and seasons
+are available, mainly in large tables. Despite these tables contains a lot of
+information, it is very hard if not impossible, to interpret, compare
+efficiently and gather insights in it. In our work we try to present this data
+in a whole different way. Therefore we have focused ourselves on some parts and
+not the entire available data.
 
@@
 advantages and limitations. In section~\ref{sec:literature} we give an overview
-of related literature and web resources, including related visualizations.  This
-includes both visualizations of NBA or other sports data, as well as
+of related literature and web resources, including related visualizations.
+This includes both visualizations of NBA or other sports data, as well as
 visualizations tackling an isue encountered during the development of our
@@
 \section{Goal and Target Audience}\label{sec:goal} 
-The visualization's goal can be best summarized by the following sentence: ``Why
-did a certain NBA-team win the Championship?'' The visualization focuses on exposing relations between team
-performance and their player roster over several years. It allows users to find
-explanations for major improvements and declines in team performance.  The
-visualization allows exploration of NBA data by lay persons.  More specifically,
-the visualization does not focus on premade explanations for phenomenons
-visible in the NBA data. By providing easy and intuitive access to the data, the
-visualization allows users to draw their own conclusions. The visualization's
-target audience are lay persons. Specifically, fans of NBA are the core of our
-target audience. This means the visualization assumes most common basketball
-terms are known to the audience and does not provide basic terminology
-explanations. 
-< We also give basic insights in why a statistic changes (arrow views)
-we try to help them explore patters and trigger them for questions
-Do we create awareness?
->
+The visualization's goal can be best summarized by the following sentence:
+``Why did a certain NBA-team win the Championship?'' The visualization focuses
+on exposing relations between team performance and their player roster over
+several years. It allows users to find explanations for major improvements and
+declines in team performance.  The visualization allows exploration of NBA data
+by lay persons.  More specifically, the visualization does not focus on premade
+explanations for phenomenons visible in the NBA data. By providing easy and
+intuitive access to the data, the visualization allows users to draw their own
+conclusions. The visualization's target audience are lay persons. Specifically,
+fans of NBA are the core of our target audience. This means the visualization
+assumes most common basketball terms are known to the audience and does not
+provide basic terminology explanations.
+
+< We also give basic insights in why a
+statistic changes (arrow views) we try to help them explore patters and trigger
+them for questions Do we create awareness?  >
 
@@
 
-On the IEEE Visualization Conference in 2013, the first workshop on sports data visualization was held \cite{ieeevis}.
-There is shown that the area of sport produces large amounts of data, which can be used for building visualizations and making analyses of the data.
-In the past visualizations of sports existed already, but there were not a lot of visualizations, and they were simple in the sense that they just represented the data without trying to give insights or to approach the data on a different way.
-There are some, but limited, sports-specific visualizations, like the basketball shot chart.
-This is a visualization frequentely seen when looking for visualizations about the NBA.
-Some shot chart vizualisations try to be different than others, but the general design and idea stays roughly the same.
+On the IEEE Visualization Conference in 2013, the first workshop on sports data
+visualization was held \cite{ieeevis}. There is shown that the area of sport
+produces large amounts of data, which can be used for building visualizations
+and making analyses of the data. In the past visualizations of sports existed
+already, but there were not a lot of visualizations, and they were simple in
+the sense that they just represented the data without trying to give insights
+or to approach the data on a different way. There are some, but limited,
+sports-specific visualizations, like the basketball shot chart. This is a
+visualization frequentely seen when looking for visualizations about the NBA.
+Some shot chart vizualisations try to be different than others, but the general
+design and idea stays roughly the same. 
 % 2 goeie papers die hierin verwerkt kunnen worden
@@
 
-With our visualization we try to make it easier to interpret a collection of data about the NBA, and try to let users find patterns and insights in the data.
-A same kind of incentive was the start for \cite{cox2006sportsvis} who created a visualization to help people discover meaning in statistics generated during sporting events in the case of baseball.
+With our visualization we try to make it easier to interpret a collection of
+data about the NBA, and try to let users find patterns and insights in the
+data. A same kind of incentive was the start for \cite{cox2006sportsvis} who
+created a visualization to help people discover meaning in statistics generated
+during sporting events in the case of baseball.
 
@@
 
-<Explain that this complies with the infovis seeking mantra: Overview, zoom&filter, details on demand
+<Explain that this complies with the infovis seeking mantra: Overview, zoom\&filter, details on demand
 play-off view = overview
@@
 - This papers are usefulle to refer to
-http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado_MultipleViews_AVI00.pdf
-http://www.interactiondesign.us/courses/2011_AD690/PDFs/Shneiderman_1996.pdf
+http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado\_MultipleViews\_AVI00.pdf
+http://www.interactiondesign.us/courses/2011\_AD690/PDFs/Shneiderman\_1996.pdf
 http://www2.parc.com/istl/groups/uir/publications/items/UIR-1986-02-Mackinlay-TOG-Automating.pdf
@@
 \subsubsection{Context section}
-The context section contains a legend that explains the stroke colors used in the bubble section.
-The stroke is gold, silver or bronze if the team ended respectively first, second or third in the regular competition.
-Teams that didn't end in a podium place during the regular competition have a red or blue stroke color indicating the team's region, respectively the western or eastern conference.
-The context section also contains a link to another page (``About'') where the user can get an explanation of the visualization with its different views, can get some background information about the NBA and the terminolgy used, and can check the source of our used data.
-If the user hovers over or selects a team in the bubble section, the context section shows more information about that team, as shown in Figure~\ref{fig:playoffviewteamselected}.
-This includes the Simple Rating System score (SRS), the end position in the play-offs and the end position in the regular competition.
+The context section contains a legend that explains the stroke colors used in
+the bubble section. The stroke is gold, silver or bronze if the team ended
+respectively first, second or third in the regular competition. Teams that
+didn't end in a podium place during the regular competition have a red or blue
+stroke color indicating the team's region, respectively the western or eastern
+conference. The context section also contains a link to another page
+(``About'') where the user can get an explanation of the visualization with its
+different views, can get some background information about the NBA and the
+terminolgy used, and can check the source of our used data. If the user hovers
+over or selects a team in the bubble section, the context section shows more
+information about that team, as shown in
+Figure~\ref{fig:playoffviewteamselected}. This includes the Simple Rating
+System score (SRS), the end position in the play-offs and the end position in
+the regular competition.
 
 \subsubsection{Bubble section}
-This section presents the roster and end positions of the play-offs for a selected season.
-Every team is respresented by a circle with the logo of the team inside it.
-Our initial desing filled the circle with the logo of the team and colored the circle red or blue depending on the region where the team competed.
-The shade of colour was darker if their final ranking was higher in their region.
-According to the Ranking of Perceptual Tasks~\cite{perceptualranking} this is a good property for ordered values like the regular competition ranking.
-But the combination of the logo's of the teams with the colors made it to heavy and less easy to compare the end ranking of different teams.
-As an alternative we tried the same view without the logo's, but this is inconvenient because you don't know which team is represented by which circle.
-Therefore we've chosen to only use the logo's of the teams as we thought this is more user friendly and the extra information of the ranking doesn't outweight.
-The size of the circle represents the SRS score using an own interpretation of perceptual scaling because the (Flannery) Appearance Compensation is not usable with a domain of both negative and positive values.
-An exponential function with exponent 1.5 was used.
-This way the proportion between the area of a bigger and the area of a smaller circle will be greater than the proportion of values the areas represent.
-The stroke of the circle illustrates the team's region or medal won during the regional competition as explained in the context section.
-The curved lines between teams represent that those teams have encountered each other in the playoffs.
-Each row represents an end position in the play-offs.
-The higher a team is, the further it has come.
-From the top to the bottom the rows represent the winner of the final (place 1), loser of the final (place 2), losers of the semi-finals (place 3), losers of the quarter-finals (place 4), losers of the 1/8-finales (place 5).
-The last row represents the teams that did not make it to the play-offs (place 6).
-A connection is always made between a team that is in a higher row and a team that is in a lower row.
-The highest team of the two teams has won the confrontation, and therefor went at least one round further in the play-offs.
-The two up most teams are the winners of the play-offs in their own conference and played the final.
-A more popular visualization of a play-off is the team double elimination bracket.
-Although this is a more common representation , and there are some nice desings like in this NBA 2012 play-offs visualisation~\cite{tournamentladder} where a sunburst visualisation is used, we wanted to create a more compact version by eliminating the recurring representation of a team.
-We considered enclosed circles, as used in this visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
-Although the enclosed circles is a compact representation as well, we have continued 
-with our own design as we believe this view is more original and innovative.
-Furthermore we are convinced that our view gives a good illustration of the end positions.
-It's generally assumed that the team that is on top of the ``ladder'' is the winning team.
-We didn't encouter a similar view when searching for alternatives to visualize a play-off or championship.
-A new visualization has the advantage that it will trigger users as well.
+This section presents the roster and end positions of the play-offs for a
+selected season. Every team is respresented by a circle with the logo of the
+team inside it. Our initial desing filled the circle with the logo of the team
+and colored the circle red or blue depending on the region where the team
+competed. The shade of colour was darker if their final ranking was higher in
+their region. According to the Ranking of Perceptual
+Tasks~\cite{perceptualranking} this is a good property for ordered values like
+the regular competition ranking. But the combination of the logo's of the teams
+with the colors made it to heavy and less easy to compare the end ranking of
+different teams. As an alternative we tried the same view without the logo's,
+but this is inconvenient because you don't know which team is represented by
+which circle. Therefore we've chosen to only use the logo's of the teams as we
+thought this is more user friendly and the extra information of the ranking
+doesn't outweight. The size of the circle represents the SRS score using an own
+interpretation of perceptual scaling because the (Flannery) Appearance
+Compensation is not usable with a domain of both negative and positive values.
+An exponential function with exponent 1.5 was used. This way the proportion
+between the area of a bigger and the area of a smaller circle will be greater
+than the proportion of values the areas represent. The stroke of the circle
+illustrates the team's region or medal won during the regional competition as
+explained in the context section. The curved lines between teams represent that
+those teams have encountered each other in the playoffs. Each row represents an
+end position in the play-offs. The higher a team is, the further it has come.
+From the top to the bottom the rows represent the winner of the final (place
+1), loser of the final (place 2), losers of the semi-finals (place 3), losers
+of the quarter-finals (place 4), losers of the 1/8-finales (place 5). The last
+row represents the teams that did not make it to the play-offs (place 6).  A
+connection is always made between a team that is in a higher row and a team
+that is in a lower row. The highest team of the two teams has won the
+confrontation, and therefor went at least one round further in the play-offs.
+The two up most teams are the winners of the play-offs in their own conference
+and played the final. 
+A more popular visualization of a play-off is the team double elimination
+bracket.  Although this is a more common representation , and there are some
+nice desings like in this NBA 2012 play-offs
+visualisation~\cite{tournamentladder} where a sunburst visualisation is used,
+we wanted to create a more compact version by eliminating the recurring
+representation of a team. We considered enclosed circles, as used in this
+visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
+Although the enclosed circles is a compact representation as well, we have
+continued with our own design as we believe this view is more original and
+innovative. Furthermore we are convinced that our view gives a good
+illustration of the end positions. It's generally assumed that the team that is
+on top of the ``ladder'' is the winning team. We didn't encouter a similar view
+when searching for alternatives to visualize a play-off or championship. A new
+visualization has the advantage that it will trigger users as well.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/presidentcandidatesvisualization}
+  \caption{The path of president candidates in their road to become president.}~\label{fig:whitehousepath}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.9\columnwidth]{figures/playoffviewfirstdesign}
+  \caption{First design of the play-off view.}~\label{fig:firstdesignplayoffview}
+\end{figure}
+
+\begin{figure}
+\centering
+  \missingfigure{}
+%  \includegraphics[width=0.9\columnwidth]{figures/playoffviewfirstimplemantation}
+  \caption{First implementation of the play-off view.}~\label{fig:firstimplementationplayoffview}
+\end{figure}
 
@@
 
-We considered a map view to just show the teams and there regular competition ranking.
-But this was a view that has no additional value in comparison with the play-off view except for the geographical position, which has no importance for our visualisation.
-Furthermore we've found a visualisation of the NBA with a map view~\cite{mapviewvisualization} similar to our idea, so we decided to not implement this map view.
+We considered a map view to just show the teams and there regular competition
+ranking. But this was a view that has no additional value in comparison with
+the play-off view except for the geographical position, which has no importance
+for our visualisation. Furthermore we've found a visualisation of the NBA with
+a map view~\cite{mapviewvisualization} similar to our idea, so we decided to
+not implement this map view.
 
@@
 \subsection{The statistics view}
-The statistics view gives a user a clear overview of how a statistic has been influenced in the selected season compared to the season before.
-The statistics view is showed in Figure~\ref{fig:statisticsview}.
-The view consists of 3 coordinated parts:
+The statistics view gives a user a clear overview of how a statistic has been
+influenced in the selected season compared to the season before. The statistics
+view is showed in Figure~\ref{fig:statisticsview}. The view consists of 3
+coordinated parts:
 \begin{itemize}
@@
 \subsubsection{Context section}
-The context section informs the user at all times of the current selection of a team  
-that have been made in the play-off view.
-The statistics view will always be based on a specific team and season that have been selected or that are initated.
-This should be clear based on the textual information in the top left.
-Next to that, in the top right a small play-off view is shown.
-It indicates the team selected and how well it performed in  the play-offs during that season, as explained in the play-off view section.
-When changing teams and/or season, this view gets updated as well as the textual information.
+The context section informs the user at all times of the current selection of a
+team  that have been made in the play-off view. The statistics view will always
+be based on a specific team and season that have been selected or that are
+initated. This should be clear based on the textual information in the top
+left. Next to that, in the top right a small play-off view is shown. It
+indicates the team selected and how well it performed in  the play-offs during
+that season, as explained in the play-off view section. When changing teams
+and/or season, this view gets updated as well as the textual information.
 <Explain that the context supports the short term memory of human beings rfr.
@@
 \subsubsection{Arrow section of selected statistic}
-This section illustrates a change for a chosen statistic for the selected team compared with the season before.
-This should give a visitor more insight into why a statistic changed over time.
-The selected team is represented by a circle in the same way as explained in the play-off view.
-The change for a statistic is represented by three arrows:
+This section illustrates a change for a chosen statistic for the selected team
+compared with the season before. This should give a visitor more insight into
+why a statistic changed over time. The selected team is represented by a circle
+in the same way as explained in the play-off view. The change for a statistic
+is represented by three arrows:
 \begin{itemize}
@@
 contribution to that arrow. Note that these shirts don't change in size. The box on 
-the right enables the user to change between 
-statistics. Next to that it informs the user of which statistic has been chosen. When 
-the user hits the circle in the middle representing the teams, he is guided 
-to the team view. This view will be explained in more detail in the next section.
+the right enables the user to change between statistics.
+Next to that it informs the user of which statistic has been chosen.
+When the user hits the circle in the middle representing the teams, he is guided to the team view.
+This view will be explained in more detail in the next section.
 <filtering: 
-- time & team
+- time \& team
 - based on exploration (small multiples)
@@
 <this vis helps us comparre statistics over time and relations between them
-eg. srs score & play-off ranking>
+eg. srs score \& play-off ranking>
 
@@
 overall performance.
-This view shows each player of the team on its field positions. Shirts aligned next to 
-each other will thus share the same position. The size of a shirt would illustrate a players performance on this spot. The PER score is initiated. This view should give
+This view shows each player of the team on its field positions.
+Shirts aligned next to each other will thus share the same position.
+The size of a shirt would illustrate a players performance on this spot.
+The PER score is initiated.
+This view should give
 a user insight on how strong or weak a particular team is for a specific field position.
-A user can then scroll through time, see how field positions evolve and what the influence
-is over time.
+A user can then scroll through time, see how field positions evolve and what the influence is over time.
 Initially we designed this view more elaborate. When the user would click on a shirt,
 specific player information would be shown. Next to some general text information, 
-we would have visualised a number of player statistics as small multiples. A bell curve
-should have informed how a player scores compared to his team, other players in the 
-league on this position or all other players in the league. We considered other charts for
-the small multiples as well. A box plot or bullet was less clear for us to indicate where
-a player stands compared to the group he was compared with. Alternatives for the small 
-multiples were evaluated as well. One good alternative was the shooting signature.\cite{peterbeshai}
+we would have visualised a number of player statistics as small multiples.
+A bell curve should have informed how a player scores compared to his team, other players in the league on this position or all other players in the league.
+We considered other charts for the small multiples as well.
+A box plot or bullet was less clear for us to indicate where a player stands compared to the group he was compared with.
+Alternatives for the small multiples were evaluated as well.
+One good alternative was the shooting signature.\cite{peterbeshai}
 However, creating such a visualization is again a project on itself and due to time 
-constraints we left it aside. Other possibilities such as a heatmap or spider web didn't
-fulfill our need to give the visitor a clear overview of how a player performed compared
-to its peergroup.
+constraints we left it aside. Other possibilities such as a heatmap or spider web didn't fulfill our need to give the visitor a clear overview of how a player performed compared to its peergroup.
 <Sketch of the small multiples>
@@
 \section{Possible improvements}
-As with most projects, the result is never 'finished', meaning that there are always improvements
-possible but due to timing or resource constraints the haven't all been implemented. In what 
-follows, we suggest a number of features that might improve the visualization:
+As with most projects, the result is never 'finished', meaning that there are always improvements possible but due to timing or resource constraints the haven't all been implemented.
+In what  follows, we suggest a number of features that might improve the visualization:
 \begin{itemize}
-    \item Smooth transitions between states could improve the user experience. For example when changing
-    year, a transformation of the arrow in the statistics view would help the user see a difference between years.
-    At the moment the arrows change at the blink of an eye, making it difficult for us humans with a short term memory
-    to make the comparison.
-    \item To further support our short term memory we could also 'save' the previous state. Eg. in the field view we could
-    visualize ghost shirts to indicate which players left the team or position and highlight players that joined the team. 
+    \item Smooth transitions between states could improve the user experience.
+For example when changing year, a transformation of the arrow in the statistics view would help the user see a difference between years.
+    At the moment the arrows change at the blink of an eye, making it difficult for us humans with a short term memory to make the comparison.
+    \item To further support our short term memory we could also 'save' the previous state.
+Eg.
+in the field view we could visualize ghost shirts to indicate which players left the team or position and highlight players that joined the team.
     This would help the user see how the team changes over time.
-    \item At the moment no attention has been payed to the process & provenance of the exploration. A user thus needs to 
-    memorize his action to be able to share or reproduce his actions. This could be improved by adding URL parameters. This
-    way a user at least can share or save a state of an analysis. One step further is to provide a play-feature. The visitor could 
-    them share his exploration and replay it. A simple implemention would ask the visitor for a start date and an end date and
+    \item At the moment no attention has been payed to the process \& provenance of the exploration.
+A user thus needs to  memorize his action to be able to share or reproduce his actions.
+This could be improved by adding URL parameters.
+This
+    way a user at least can share or save a state of an analysis.
+One step further is to provide a play-feature.
+The visitor could them share his exploration and replay it.
+A simple implemention would ask the visitor for a start date and an end date and
     would then 'play' the changes of a particular team and characteristic over time.
-    \item We could support users further in exploring for patterns by focussing more on comparisons. We could integrate a view 
-    where a visitor is able to compare two teams in the same year, or in two different years.
+    \item We could support users further in exploring for patterns by focussing more on comparisons.
+We could integrate a view where a visitor is able to compare two teams in the same year, or in two different years.
 \end{itemize}
@@
 This due to a increase inside the team and because of new players joining the team in 2009.
-In
-the years following, the team kept on attracting talented players and increased the inherent team score (with ups and downs) to become NBA champion in 2015.
+In the years following, the team kept on attracting talented players and increased the inherent team score (with ups and downs) to become NBA champion in 2015.
 With the team view, one can even notice that the team gets stronger in center team positions during this period.
-The visualization however is missing at the moment if this reinforcement of the field positions
-is due to the attraction of talented people or by utilizing available talent within the team.
+The visualization however is missing at the moment if this reinforcement of the field positions is due to the attraction of talented people or by utilizing available talent within the team.
 Rfr to possible improvements.
@@
 \subsection{Each component should be informative}
-In order to have a good visualization as a whole, each component should be a clear and 
-informative visualization on itself. This is a lesson learned from our work, but also when
-evaluating the work of peers. We started noticing that most components in a visualization 
-lack references and hence are not informative enough. Eg. each small multiple of a statisctic
-in the statistics view on itself gives sufficient information to stand on itselfs. The bar in
-the small multiple indicates the score in the selected season and enables the user to evaluate
-that score over time. Have there been better scores, or worse scores. Is this score part of 
-an upward movement during years? 
-For this reason we can argue that the circle representing the teams SRS score in the statisctics
-view is not the best choice. We could have chosen a bar chart with an indication of the best and
-worst score for that specific team over time or for the best and worst team in that season.
+In order to have a good visualization as a whole, each component should be a clear and informative visualization on itself.
+This is a lesson learned from our work, but also when evaluating the work of peers.
+We started noticing that most components in a visualization lack references and hence are not informative enough.
+Eg.
+each small multiple of a statisctic in the statistics view on itself gives sufficient information to stand on itselfs.
+The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
+Have there been better scores, or worse scores.
+Is this score part of an upward movement during years?
+For this reason we can argue that the circle representing the teams SRS score in the statisctics view is not the best choice.
+We could have chosen a bar chart with an indication of the best and worst score for that specific team over time or for the best and worst team in that season.
 

### 2016-05-21 paper/paper.tex
Reworked the paper, moved some stuff around, related work should be rewritten, then I think we're pretty much done

@@
-\documentclass{sigchi}
+\documentclass[chi_draft]{sigchi}
 % Use this command to override the default ACM copyright statement
@@
 \begin{abstract}
-    In this paper we describe an interactive visualization of the NBA competition over the last 30 years.
-With three different views, we give the opportunity for the user to explore various kinds of data about NBA seasons including standings, team statistics, transfers and player statistics.
-The d3.js framework was used to develop this visualization.
-    %TODO add results
-
+    Visualization of sports data is very common in recent years as access to
+    data has increased. The abundance of data and audience allows for new and
+    innovative presentations. In this paper we describe an interactive
+    visualization of the NBA competition over the last 30 years. With three
+    different views, we allow the user to explore data about NBA seasons
+    including standings, team statistics, transfers and player statistics. The
+    d3.js framework was used to develop this visualization. 
 \end{abstract}
@@
 \category{H.5.m.}{Information Interfaces and Presentation
-  (e.g. HCI)}{Miscellaneous} \category{See
-  \url{http://acm.org/about/class/1998/} for the full list of ACM
-  classifiers. This section is required.}{}{}
+  (e.g. HCI)}{Miscellaneous} \category{}{}{}
 
@@
 
-The National Basketball Association (NBA) is the most famous and perhaps the
-best professional basketball league in the world. Like in every other sports
+The National Basketball Association (NBA) is the most famous professional
+basketball league in the world\cite{nbawiki}. As in every other sports
 branch, a lot of data and statistics about players, teams, games and seasons
-are available, mainly in large tables. Despite these tables contains a lot of
-information, it is very hard if not impossible, to interpret, compare
-efficiently and gather insights in it. In our work we try to present this data
-in a whole different way. Therefore we have focused ourselves on some parts and
-not the entire available data.
+are available, mainly in large tables. Despite the large amount of data
+available, gathering useful insights from this data can be surprisingly
+challenging. In our work we try to present this data in an innovative and
+intuitive way, allowing the user to draw meaningful conclusions.
 
-This paper describes the visualization of NBA data made by the ``The Tufters''
-team for the course ``Information Visualization''. In section~\ref{sec:goal} we
-describe the goal of the visualization and the target audience. In
-section~\ref{sec:data} we describe the data used, including its origins,
-advantages and limitations. In section~\ref{sec:literature} we give an overview
-of related literature and web resources, including related visualizations.
-This includes both visualizations of NBA or other sports data, as well as
-visualizations tackling an isue encountered during the development of our
-visualization. In section~\ref{sec:visualization} we describe the visualization
-itself. We give an overview of the different stages of development of the
-visualization, as well as the major design decisions.
+In section~\ref{sec:goal} we describe the goal of the visualization and the
+target audience. In section~\ref{sec:data} we describe the data used, including
+its origins, advantages and limitations. In section~\ref{sec:literature} we give
+an overview of related literature and web resources, including related
+visualizations.  This includes both visualizations of NBA or other sports data,
+as well as visualizations tackling an isue encountered during the development of
+our visualization. In section~\ref{sec:visualization} we describe the
+visualization itself. We give an overview of the different stages of development
+of the visualization, as well as the major design decisions.
 Section~\ref{sec:discussion} discusses potential improvements of the final
@@
 conclusions. The visualization's target audience are lay persons. Specifically,
-fans of NBA are the core of our target audience. This means the visualization
-assumes most common basketball terms are known to the audience and does not
-provide basic terminology explanations.
-
-< We also give basic insights in why a
-statistic changes (arrow views) we try to help them explore patters and trigger
-them for questions Do we create awareness?  >
+basketball fans are the core of our target audience. It can be of particular
+interest with the recent surge in fantasy leagues, for which data analysis is
+paramount to a player's success~\cite{fantasy,fantasyskill}.
 
@@
 such as field goals, percentage of shots scored, shooting distance, minutes
-played per season, number of personal fouls made, salaries and much more. These
+played per season, number of personal fouls made, salaries and much more. All of
+this data is available as time-series. These
 statistics are avalable for each player individually for each season. Aggregated
@@
 during the playoffs, and the rest of the data was combined into json format. The
-final preprocessing step combined all data into one json file.
-
-<Tell something more about the data in terms of tall/wide data
-Mention that it's a time series
-Tell something more about pre-processing steps (why?, did we change it over time?)
-rfr to picture in 1st ppt sl 60>
+final preprocessing step combined all data into one json file. A model of the
+available data can be seen on figure~\ref{fig:data}.
 
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/data}
-  \caption{An overview of the data.}~\label{fig:data}
+  \includegraphics[width=1.0\columnwidth]{figures/data}
+  \caption{An overview of the data.}
+  \label{fig:data}
 \end{figure}
 
+\todo{Rework realted work}
 \section{Related Work}\label{sec:literature}
@@
 during sporting events in the case of baseball.
+
+(From bubble view)
+A more traditional visualization of a play-off is the team double elimination
+bracket.  Although this is a more common representation , and there are some
+nice desings like in this NBA 2012 play-offs
+visualisation~\cite{tournamentladder} where a sunburst visualisation is used,
+we wanted to create a more compact version by eliminating the recurring
+representation of a team. We considered enclosed circles, as used in this
+visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
+Although the enclosed circles is a compact representation as well, we have
+continued with our own design as we believe this view is more original and
+innovative. Furthermore we are convinced that our view gives a good
+illustration of the end positions. It's generally assumed that the team that is
+on top of the ``ladder'' is the winning team. We didn't encouter a similar view
+when searching for alternatives to visualize a play-off or championship. A new
+visualization has the advantage that it will trigger users as well.
+
+(From field view)
+One good alternative was the shooting signature.\cite{peterbeshai} However,
+creating such a visualization is again a project on itself and due to time
+constraints we left it aside. Other possibilities such as a heatmap or spider
+web didn't fulfill our need to give the visitor a clear overview of how a player
+performed compared to its peergroup.
 
@@
     \item The play-off view: a compact view on the play-offs per season
-    \item The statistics (zoom) view: a more detailed view on a statistic of a selected team
-    \item The team view: a detailed view on how good or bad a team scores on a specific field position
+    \item The statistics (zoom) view: a more detailed view on a statistic of a
+        selected team
+    \item The team view: a detailed view on how good or bad a team scores on a
+        specific field position
 \end{itemize}
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/playoffviewwithcontext}
+  \includegraphics[width=1.0\columnwidth]{figures/playoffviewwithcontext}
   \caption{The play-off view without a team selected.}~\label{fig:playoffviewnoteam}
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/playoffviewteamselected}
+  \includegraphics[width=1.0\columnwidth]{figures/playoffviewteamselected}
   \caption{The play-off view with a team selected.}~\label{fig:playoffviewteam}
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/statisticsview}
+  \includegraphics[width=1.0\columnwidth]{figures/statisticsview}
   \caption{The statistics view.}~\label{fig:statisticsview}
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/teamview}
-  \caption{The team view.}~\label{fig:teamview}
+  \includegraphics[width=1.0\columnwidth]{figures/teamview}
+  \caption{The team view.}
+  \label{fig:teamview}
 \end{figure}
 
-<Explain that this complies with the infovis seeking mantra: Overview, zoom\&filter, details on demand
-play-off view = overview
-statistics view= details on demand
-Talk about th organization of the 3 views + the choice for a tiled approach
-- Each page is a 'tile' that we force with fixed scrolling
-- Use of the keyboard is possible (space bar to scroll + arrows, arrows for timeline)
-- This papers are usefulle to refer to
-http://courses.ischool.berkeley.edu/i247/f05/readings/Baldonado\_MultipleViews\_AVI00.pdf
-http://www.interactiondesign.us/courses/2011\_AD690/PDFs/Shneiderman\_1996.pdf
-http://www2.parc.com/istl/groups/uir/publications/items/UIR-1986-02-Mackinlay-TOG-Automating.pdf
->
+This technique follows the standards of information visualization. It provides
+an overview, allows zooming and filtering and provides details on
+demand\cite{mantra,multipleviews,automatingdesign}. The play-off view provides
+the full overview which allows filtering by team. The statistics view provides
+the ability to zoom into one specific team. There some details are provided
+instantly, while others can be procured by hovering over certain elements.
 
-All three are discussed in more detail below. The three views are available to
-the user on a single webpage. The user start on the play-off view and is able to
-navigate to the statistics and team views by scrolling down or clicking on
-teams. A different season can be selected with the left and right arrows or by clicking on the timeline. The timeline is a fixed part of the visualization, hence visible in the three different views.
-%The webpage is built in such a way that it gives a hint to the user at the bottom
-%or the top of a page when another view is available. By showing a different 
-%background color and an arrow, we invite the visitor to explore other views. The 
-%user is also at liberty to scroll through these views. However, a fixed scroll is
-%implemeted so users will only end in complete views when moving around. Note that 
-%the user also will be guided through all 3 views when performing actions described 
-%in the section below.
+The organization of these three views is in a vertical layout with fixed
+scrolling. Navigation with the keyboard is supported as well.
+
+The choice was made to use a thematic background instead of opting for the
+traditional minimalistic approach to information visualization. This was done to
+improve the attractiveness of the visualisation on the one hand and the users'
+memory of it on the other hand\cite{aesthetics}.
+
+All three views are discussed in more detail below. The three views are
+available to the user on a single webpage. The user start on the play-off view
+and is able to navigate to the statistics and team views by scrolling down or
+clicking on teams. A different season can be selected with the left and right
+arrows or by clicking on the timeline. The timeline is a fixed part of the
+visualization. It is visible in the three different views.
 
 \subsection{The play-off view}
-This view informs the user of the roster and end positions in the NBA play-offs for a selected season, in general or for a selected team. The view consists of two parts: 
+This view informs the user of the roster and end positions in the NBA play-offs
+for a selected season, in general or for a selected team. The view consists of
+two parts: 
 \begin{itemize}
-    \item Context section: the top part informs the user about the selected team and shows a legend of the play-off view.
-    \item Bubble section: this section represents the play-offs of the selected season
+    \item Context section: the top part informs the user about the selected team
+    and shows a legend of the play-off view.
+    \item Bubble section: this section represents the play-offs of the selected
+        season.
 \end{itemize}
-The play-off view is shown in Figure~\ref{fig:playoffviewwithcontext}.
+The play-off view is shown in figure~\ref{fig:playoffviewwithcontext}.
 
 \subsubsection{Context section}
-The context section contains a legend that explains the stroke colors used in
-the bubble section. The stroke is gold, silver or bronze if the team ended
-respectively first, second or third in the regular competition. Teams that
-didn't end in a podium place during the regular competition have a red or blue
-stroke color indicating the team's region, respectively the western or eastern
-conference. The context section also contains a link to another page
-(``About'') where the user can get an explanation of the visualization with its
-different views, can get some background information about the NBA and the
-terminolgy used, and can check the source of our used data. If the user hovers
-over or selects a team in the bubble section, the context section shows more
-information about that team, as shown in
-Figure~\ref{fig:playoffviewteamselected}. This includes the Simple Rating
-System score (SRS), the end position in the play-offs and the end position in
-the regular competition.
+The context section contains a legend for the bubble section. The bubble's
+borders are gold, silver or bronze if the team ended respectively first, second
+or third in the regular competition. Teams that didn't end in a podium place
+during the regular competition have a red or blue stroke color indicating the
+team's region, respectively the western or eastern conference. The context
+section also contains a link to an ``about'' page where the user can get an
+explanation of the visualization with its different views, can get some
+background information about the NBA and the terminolgy used, and can check the
+source of our used data. If the user hovers over or selects a team in the bubble
+section, the context section shows more information about that team, as shown in
+Figure~\ref{fig:playoffviewteamselected}. The information shown includes the
+Simple Rating System score (SRS), the end position in the play-offs and the end
+position in the regular competition.
 
@@
 selected season. Every team is respresented by a circle with the logo of the
-team inside it. Our initial desing filled the circle with the logo of the team
-and colored the circle red or blue depending on the region where the team
-competed. The shade of colour was darker if their final ranking was higher in
-their region. According to the Ranking of Perceptual
-Tasks~\cite{perceptualranking} this is a good property for ordered values like
-the regular competition ranking. But the combination of the logo's of the teams
-with the colors made it to heavy and less easy to compare the end ranking of
-different teams. As an alternative we tried the same view without the logo's,
-but this is inconvenient because you don't know which team is represented by
-which circle. Therefore we've chosen to only use the logo's of the teams as we
-thought this is more user friendly and the extra information of the ranking
-doesn't outweight. The size of the circle represents the SRS score using an own
+team inside it. The size of the circle represents the SRS score using an own
 interpretation of perceptual scaling because the (Flannery) Appearance
@@
 between the area of a bigger and the area of a smaller circle will be greater
-than the proportion of values the areas represent. The stroke of the circle
-illustrates the team's region or medal won during the regional competition as
-explained in the context section. The curved lines between teams represent that
-those teams have encountered each other in the playoffs. Each row represents an
-end position in the play-offs. The higher a team is, the further it has come.
-From the top to the bottom the rows represent the winner of the final (place
-1), loser of the final (place 2), losers of the semi-finals (place 3), losers
-of the quarter-finals (place 4), losers of the 1/8-finales (place 5). The last
-row represents the teams that did not make it to the play-offs (place 6).  A
-connection is always made between a team that is in a higher row and a team
-that is in a lower row. The highest team of the two teams has won the
-confrontation, and therefor went at least one round further in the play-offs.
-The two up most teams are the winners of the play-offs in their own conference
-and played the final. 
-A more popular visualization of a play-off is the team double elimination
-bracket.  Although this is a more common representation , and there are some
-nice desings like in this NBA 2012 play-offs
-visualisation~\cite{tournamentladder} where a sunburst visualisation is used,
-we wanted to create a more compact version by eliminating the recurring
-representation of a team. We considered enclosed circles, as used in this
-visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
-Although the enclosed circles is a compact representation as well, we have
-continued with our own design as we believe this view is more original and
-innovative. Furthermore we are convinced that our view gives a good
-illustration of the end positions. It's generally assumed that the team that is
-on top of the ``ladder'' is the winning team. We didn't encouter a similar view
-when searching for alternatives to visualize a play-off or championship. A new
-visualization has the advantage that it will trigger users as well.
+than the proportion of values the areas represent.
+
+The curved lines between teams represent that those teams have encountered each
+other in the playoffs. Each row represents an end position in the play-offs. The
+higher a team is, the further it has come.  From the top to the bottom the rows
+represent the winner of the final (rank 1), loser of the final (rank 2), losers
+of the semi-finals (rank 3), losers of the quarter-finals (rank 4), losers of
+the eighth finals (rank 5). The last row represents the teams that did not make
+it to the play-offs (rank 6).
+
+This simple presentation intuitively puts the winning teams higher than the
+losing teams. The winner of a certain game can be easily deriving from their
+relative vertical position. It is also clear that each team neads to have played
+and won a series of matches against one other team at every level below
+them~\ref{fig:playoffviewteamselected}. This view is further reinforced by
+highlighting a team and its matches and opponents when hovering over it.
+Additional team information is then also shown in the context part.
 
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/presidentcandidatesvisualization}
+  \includegraphics[width=1.0\columnwidth]{figures/presidentcandidatesvisualization}
   \caption{The path of president candidates in their road to become president.}~\label{fig:whitehousepath}
@@
 \centering
-  \includegraphics[width=0.9\columnwidth]{figures/playoffviewfirstdesign}
+  \includegraphics[width=1.0\columnwidth]{figures/playoffviewfirstdesign}
   \caption{First design of the play-off view.}~\label{fig:firstdesignplayoffview}
@@
 
-<Talk about the importance of estethics seen in the lesson>
-Our initial play-off view was too heavy. It connected teams that played against
-each other directly. When searching for alternatives to reflect on our choice we
-discovered a similar visualization in a totally different context. 
-<https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/>
-<comparison of original play-off view with white house paths>
-This visualization gives a cleaner view of the competitors for a specific team
+\subsubsection{Previous designs}
+Our initial design used the team circles' lightness to indicate their final
+ranking in their region. The hue was used to show the league they were part of.
+According to the Ranking of Perceptual Tasks~\cite{perceptualranking} this is a
+good property for ordered values like the regular competition ranking. But the
+combination of the logo's of the teams with the colors resulted in an overload
+of informaiton and made it more difficult to compare the end ranking of
+different teams.
+
+As an alternative we tried the same view without the logo's. This was deemed
+inconvenient because you cannot instantly see which team is represented by which
+circle. Therefore we've chosen to only use the teams' logos.
+
+Our initial play-off view was too dense because of the connections between teams
+as well. It connected teams that played against each other directly. When
+searching for alternatives to reflect on our choice we discovered a similar
+visualization in a totally different context\cite{whitehousepath}.  This
+visualization gives a cleaner view of the competitors for a specific team
 compared to our original sketch. Hence we decided to adapt our visualization. We
-added an intermediate step between two teams to clearer indicate how teams competed
-to become NBA champion.
-Next to the static view, the user can interact with the visualization. When he 
-hoovers over a circle of a team, only the games played by that team are shown by highliting the 
-competitors (their circles) and the curved lines that connect them. This view gets 
-fixated when the users clicks on a team.
-He/she then automatically gets forwarded to the statistics view where the play-off 
-view is still represented in the right upper corner to make sure the user keeps
-an overview. The statistics view will be explained in more detail in the next section.
+added an intermediate step between two teams to more clearly indicate how teams
+competed to become NBA champion.
 
-< Hierarchy = sorting = facilitate trends of winners
+\todo{Hierarchy = sorting = facilitate trends of winners
 position have a meaning in the play-off view
@@
 - size of circles
-- hue to highligt selections or indicate increase decline arrows next view
+- hue to highligt selections or indicate increase decline arrows next view}
 
-We considered a map view to just show the teams and there regular competition
-ranking. But this was a view that has no additional value in comparison with
-the play-off view except for the geographical position, which has no importance
-for our visualisation. Furthermore we've found a visualisation of the NBA with
-a map view~\cite{mapviewvisualization} similar to our idea, so we decided to
-not implement this map view.
-
-This visualization helped us discover data quality issues => cut off year 1984
->
+We initially also considered a map view to show the teams and their regular
+competition ranking on a map of the USA. However, this view had no additional
+value compared to the play-off view except for the geographical position, which
+has no importance for our visualisation. Furthermore we've found a visualisation
+of the NBA with a map view~\cite{mapviewvisualization} similar to our idea, so
+we decided to not implement this map view.
 
@@
 \begin{itemize}
-    \item Context section: the top part informs the user of the context of the selected team
-    \item Arrow section: this section illustrates why a statistic changed compared to the previous season
-    \item Small multiples section: small multiples of team statistics and how they evolve over time
+    \item Context section: the top part informs the user of the context of the
+        selected team
+    \item Arrow section: this section illustrates why a statistic changed
+        compared to the previous season
+    \item Selected statistics section: team statistics and how they evolve over
+        time
 \end{itemize}
@@
 \subsubsection{Context section}
-The context section informs the user at all times of the current selection of a
-team  that have been made in the play-off view. The statistics view will always
-be based on a specific team and season that have been selected or that are
-initated. This should be clear based on the textual information in the top
-left. Next to that, in the top right a small play-off view is shown. It
-indicates the team selected and how well it performed in  the play-offs during
-that season, as explained in the play-off view section. When changing teams
-and/or season, this view gets updated as well as the textual information.
-<Explain that the context supports the short term memory of human beings rfr.
-Monkey Illusion>
+The context section is partly identical to that of the play-offs view. The
+team's information is also displayed in the top left. In the top right, the
+bubble section of the play-offs view is displayed in a reduced, minimalistic
+form. This view is repeated to create a sense of continuity between the
+different views and provide context of the current position in the
+visualization. The team shown in the statistics view is highlighted on this
+smaller bubble section as well. Replicating this section also alleviates
+short-term memory issues. It prevents the need to scroll back and forth between
+the different views constantly.
 
@@
 This section illustrates a change for a chosen statistic for the selected team
-compared with the season before. This should give a visitor more insight into
-why a statistic changed over time. The selected team is represented by a circle
-in the same way as explained in the play-off view. The change for a statistic
-is represented by three arrows:
+compared to the previous season. This gives a user more insight into why a
+statistic changed over time and how this change has impacted team performance.
+The selected team is represented by a circle in the same way as explained in the
+play-off view. The change for a statistic is represented by three arrows:
 \begin{itemize}
-    \item Arrow on the left pointing towards the circle: indicating the influence of players who joined the team (team member current season, not previous season)
-    \item Arrow on the right of the circle: indicating how the team inherent changed (team member this season and previous season)
-    \item Arrow on the left pointing away from the circle: indicating the influence of players who left the team (team member previous season, not current season)
+    \item Arrow on the left pointing towards the circle: indicating the
+        influence of players who joined the team 
+    \item Arrow on the right of the circle: indicating how the team internally
+        changed, e.g. by training its players.
+    \item Arrow on the left pointing away from the circle: indicating the
+        influence of players who left the team 
 \end{itemize}
-<figure of the statistics view>
-The shirts next to the arrows are the shirts of players that had the biggest 
-contribution to that arrow. Note that these shirts don't change in size. The box on 
-the right enables the user to change between statistics.
-Next to that it informs the user of which statistic has been chosen.
-When the user hits the circle in the middle representing the teams, he is guided to the team view.
-This view will be explained in more detail in the next section.
-<filtering: 
-- time \& team
-- based on exploration (small multiples)
-- dynamically: changes impact multiple parts>
+\todo{add statistics view image here}
+The jerseys next to the arrows are the jerseys of players that had the biggest
+contribution to that arrow. On the left, different statistics , including player
+PER can be chosen from. Below this, player information is shown when hovering
+over a player's jersey.
 
-\subsubsection{Small multiples of different statistics}
-Below the arrow section small multiples are shown of multiple statistics over time. 
-The line charts enables a user to identify peaks or drops that possibly influenced the 
-outcome of a team is this season, following seasons or previous seasons.
-A user interacts with the visualization by shifting the bar indicating the current year displayed.
-By doing this the view gets updated and the user can see the impact on the team in the context section and why the statistic changes in the arrow section.
-An alternative to scroll through time is to use the timeline at the bottom of the page.
-<this vis helps us comparre statistics over time and relations between them
-eg. srs score \& play-off ranking>
+\subsubsection{Selected statistics}
+Below the arrow section some selected statistics are shown over
+time.  The line charts enables a user to identify peaks or drops that
+influenced the outcome of a team is this season, following seasons or previous
+seasons. A user interacts with the visualization by shifting the bar indicating
+the current year displayed. By doing this the view gets updated and the user can
+see the impact on the team in the context section and why the statistic changes
+in the arrow section. This view provides a lot of additional information to
+users by allowing them to see correlations between statistics and phenomena
+quickly and easily.
 
 \subsection{The team view}
-This view as well as the statistics view should give the user the ability to search
-for explanations why a team is is performing better or worse over seasons. On the 
-other hand, a user can see what impact a change in team characteristics has on its
-overall performance.
-This view shows each player of the team on its field positions.
-Shirts aligned next to each other will thus share the same position.
-The size of a shirt would illustrate a players performance on this spot.
-The PER score is initiated.
-This view should give
-a user insight on how strong or weak a particular team is for a specific field position.
-A user can then scroll through time, see how field positions evolve and what the influence is over time.
-Initially we designed this view more elaborate. When the user would click on a shirt,
-specific player information would be shown. Next to some general text information, 
-we would have visualised a number of player statistics as small multiples.
-A bell curve should have informed how a player scores compared to his team, other players in the league on this position or all other players in the league.
-We considered other charts for the small multiples as well.
-A box plot or bullet was less clear for us to indicate where a player stands compared to the group he was compared with.
-Alternatives for the small multiples were evaluated as well.
-One good alternative was the shooting signature.\cite{peterbeshai}
-However, creating such a visualization is again a project on itself and due to time 
-constraints we left it aside. Other possibilities such as a heatmap or spider web didn't fulfill our need to give the visitor a clear overview of how a player performed compared to its peergroup.
-<Sketch of the small multiples>
+This view shows each player of the team on its field positions. Players are
+represented by their jerseys. Jerseys aligned next to each other share the same
+position, which is indicated as well. The size of a jersey would illustrate a
+players performance on this spot. This view as well as the statistics view
+should give the user the ability to search for explanations why a team is is
+performing better or worse over seasons. On the other hand, a user can see what
+impact a change in team characteristics has on its overall performance. This
+view should give a user insight on how strong or weak a particular team is for a
+specific field position. A user can then scroll through time to see how field
+positions evolve and what the influence is over time.
 
-\subsection{Improving the result}
+\subsubsection{Previous designs}
+Initially we designed this view more elaborate. When the user would click on a
+jersey, specific player information would be shown. A number of player
+statistics would be shown as small multiples. A bell curve should have informed
+how a player scores compared to his team, other players in the league on this
+position or all other players in the league. We considered other charts for the
+small multiples as well. A box plot or bullet was less clear for us to indicate
+where a player stands compared to the group he was compared with. Alternatives
+for the small multiples were evaluated as well. This part of the visualization
+was deemed less important than the others and has not been fully developed
+because of time and resource constraints. \todo{sketch of small multiples
+here}
+
+\subsection{Improving the result}\todo{Is this necessary? I don't know what to
+write here}
 Talk about how we improved the data-ink ration in the end..
 
-\section{Possible improvements}
-As with most projects, the result is never 'finished', meaning that there are always improvements possible but due to timing or resource constraints the haven't all been implemented.
-In what  follows, we suggest a number of features that might improve the visualization:
+\section{Future work}
+As with most projects, the result is never 'finished', meaning that there are
+always improvements possible but due to timing or resource constraints they
+haven't all been implemented. We suggest a number of features that might improve
+the visualization:
 \begin{itemize}
     \item Smooth transitions between states could improve the user experience.
-For example when changing year, a transformation of the arrow in the statistics view would help the user see a difference between years.
-    At the moment the arrows change at the blink of an eye, making it difficult for us humans with a short term memory to make the comparison.
-    \item To further support our short term memory we could also 'save' the previous state.
-Eg.
-in the field view we could visualize ghost shirts to indicate which players left the team or position and highlight players that joined the team.
-    This would help the user see how the team changes over time.
-    \item At the moment no attention has been payed to the process \& provenance of the exploration.
-A user thus needs to  memorize his action to be able to share or reproduce his actions.
-This could be improved by adding URL parameters.
-This
-    way a user at least can share or save a state of an analysis.
-One step further is to provide a play-feature.
-The visitor could them share his exploration and replay it.
-A simple implemention would ask the visitor for a start date and an end date and
-    would then 'play' the changes of a particular team and characteristic over time.
-    \item We could support users further in exploring for patterns by focussing more on comparisons.
-We could integrate a view where a visitor is able to compare two teams in the same year, or in two different years.
+        For example when changing year, a transformation of the arrow in the
+        statistics view would help the user see a difference between years.  At
+        the moment the arrows change at the blink of an eye, making it difficult
+        for us humans with a short term memory to make the comparison. This
+        would have been especially useful in the transitions between different
+        years on the bubble view.
+    \item To further support our short term memory we could also 'save' the
+        previous state. Eg. in the field view we could visualize ghost jerseys
+        to indicate which players left the team or position and highlight
+        players that joined the team.  This would help the user see how the team
+        changes over time.
+    \item At the moment no attention has been payed to the process, provenance
+        and history of the user's exploration session. A user thus needs to
+        memorize his action to be able to share or reproduce his actions. This
+        could be improved by adding URL parameters. This way a user at least can
+        share or save a state of an analysis. One step further is to provide a
+        play-feature. The visitor could them share his exploration and replay
+        it. A simple implemention would ask the visitor for a start date and an
+        end date and would then 'play' the changes of a particular team and
+        characteristic over time.
+    \item We could support users further in looking for patterns by focussing
+        more on comparisons. We could integrate a view where a visitor is able
+        to compare two teams in the same year, or in two different years.
 \end{itemize}
@@
 \subsection{Technology}
-To create the visualization, the d3 javascript framework~\cite{d3.js} was used in combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3 is based on HTML, CSS, Javascript and SVG.
-Next to that,the use of D3 was a requirement for the project. This allows easy
-access to the visualization as most modern browsers are capable of handling
-these webstandards. 
-The choice not to use the d3 framework for the entire visualization was made to
-ease the layout configuration of the visualization. Instead, html was used to do
-the global layout of the visualization. We opted to structure the site with multiple
-divs as containers. To each container we allocate a svg with a specific visualization.
-Eg. The statistic view is build out of x divs filled with different svg's created
-with the d3 framework.
-<sketch of how the page is structured>
+To create the visualization, the d3 javascript framework~\cite{d3.js} was used
+in combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3
+is based on HTML, CSS, Javascript and SVG. This allows easy access to the visualization as
+most modern browsers are capable of handling these webstandards.  The choice not
+to use the d3 framework for the entire visualization was made to ease the layout
+configuration of the visualization. Instead, html was used to do the global
+layout of the visualization. We opted to structure the site with multiple divs
+as containers. To each container we allocate a svg with a specific
+visualization.
+
+\todo{sketch of how the page is structured here}
 When working with d3.js we encountered some obstacles:
-- A synchrounous call need to be combined with synchronous calls.
-More specific, when a visitor uses the timeline to scroll through time, the whiping of previous visualizations was not synced with the creation of the visualisation.
-This resulted in whiping sections when a visualization was not created yet.
-Hence the view resulted in multiple figures overlapping.
+\begin{itemize}
+        \item A synchrounous call need to be combined with synchronous calls.
+            More specific, when a visitor uses the timeline to scroll through
+            time, the whiping of previous visualizations was not synced with the
+            creation of the visualisation. This resulted in whiping sections
+            when a visualization was not created yet. Hence the view resulted in
+            multiple figures overlapping.
+\end{itemize}
 
@@
 At the beginning of this project we were able to design the 3 separate views
-quite quickly.
-Our main problem was to integrate these views as a whole.
-When we created drafts of two views being the play-off view and the statitics view, we suddenly noted patterns.
-The Golden State Warriors became NBA champion a couple of years after a great uplift in SRS score.
-This due to a increase inside the team and because of new players joining the team in 2009.
-In the years following, the team kept on attracting talented players and increased the inherent team score (with ups and downs) to become NBA champion in 2015.
-With the team view, one can even notice that the team gets stronger in center team positions during this period.
-The visualization however is missing at the moment if this reinforcement of the field positions is due to the attraction of talented people or by utilizing available talent within the team.
-Rfr to possible improvements.
-Being able to visualize this story, enabled us to support visistors in searching their own stories.
-They are able to explore the data for patters.
-On top of that, we provide the insight on how the changes were impacted by changes in teamplayers.
+quite quickly. Our main problem was to integrate these views as a whole. When we
+created drafts of two views being the play-off view and the statitics view, we
+suddenly noted patterns. The Golden State Warriors became NBA champion a couple
+of years after a great uplift in SRS score. This was because of new players
+joining the team in 2009. In the years following, the team kept on attracting
+talented players and increased the inherent team score to become NBA champion in
+2015. With the team view, one can even notice that the team gets stronger in
+center team positions during this period. Users are able to explore the data for
+patters. On top of that, we provide the insight on how the changes were impacted
+by changes in teamplayers.
 
 \subsection{Each component should be informative}
-In order to have a good visualization as a whole, each component should be a clear and informative visualization on itself.
-This is a lesson learned from our work, but also when evaluating the work of peers.
-We started noticing that most components in a visualization lack references and hence are not informative enough.
-Eg.
-each small multiple of a statisctic in the statistics view on itself gives sufficient information to stand on itselfs.
-The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
-Have there been better scores, or worse scores.
-Is this score part of an upward movement during years?
-For this reason we can argue that the circle representing the teams SRS score in the statisctics view is not the best choice.
-We could have chosen a bar chart with an indication of the best and worst score for that specific team over time or for the best and worst team in that season.
+In order to have a good visualization as a whole, each component should be a
+clear and informative visualization on itself. During the project, we started noticing that most
+components in a visualization lack references and hence are not informative
+enough. Eg. each small multiple of a statisctic in the statistics view on itself
+gives sufficient information to stand on itselfs. The bar in the small multiple
+indicates the score in the selected season and enables the user to evaluate that
+score over time. Have there been better scores, or worse scores. Is this score
+part of an upward movement during years? This should be weighed against the
+duplication of data and use of screen space. Coordinates views alleviate this
+need significantly as well, as each component is strongly supported by other
+components.
 
@@
 Before exploring what is out there in the d3.js world, we made our own sketches.
-This is how we came up with the 3 custom views that makes our visualization.
-When comparing our design with alternatives we could not find something that satisfied our needs.
-Nonetheless, they gave us inspiration to finetune our designs.
-Although we are satisfied with our result and we are convinced it was the right choice to reach our goal, it had some draw backs.
-We noted that other teams had more freedom in exploring multiple different visualizations which enabled them to evaluate different options with trial and error.
-We could have prepared our data in a format that was more compatible with out of the box examples, but this would have had too much of an impact on our result.
-<Talk about redesign>
-Next to that, although we've tried to write our code as general as possible, we're not convinced this could be shared in such a way that other people building a visualization could easily re-use our work to visualize other sports.
+This is how we came up with the 3 custom views that make up our visualization.
+When comparing our design with alternatives we could not find something that
+satisfied our needs. Nonetheless, they gave us inspiration to finetune our
+designs. Although we are satisfied with our result and we are convinced it was
+the right choice to reach our goal, it had some draw backs. We noted that other
+teams had more freedom in exploring multiple different visualizations which
+enabled them to evaluate different options with trial and error. Much of the
+time developing was spent on the technical details for these custom
+visualizations. This was especially true for the bubble component of the
+playoffs view. No out of the box solution for this could be found so everything
+from the determination of the positions of the bubbles to how to connect them
+had to be developed by us.
 
 \section{Conclusion}\label{sec:conclusion}
+We believe the visualization we have created can be of tremendous value for fans
+and fantasy basketball players. Personally we have gained several insights into
+the performance of basketball teams by using the visualization and we believe
+others could learn even more. 
+The most interesting conclusion we were able to make concerns the importance of
+individual players in a basketball team. We were able to find strong
+correlations between having two to four very strong players and the team's
+performance. It seems having one star player is not enough. On the other hand,
+having more than 5 good players no longer has much influence either.
+
+While the visualization has a lot of potential for improvement, we think the
+innovative play-offs representation combined with detailed team information in
+an integrated fashion has much value.
 
@@
 
+\paragraph{NBA} National Basketball Association - The largest, most well-known
+basketball of the USA.
+
 \paragraph{PER} Player Efficiency Rating - An all-in-one basketball rating,
 boiling down all of a player's contributions into one number\cite{per}.
+
+\paragraph{SRS} Simple Rating System - a rating that takes into account average
+point differential and strength of schedule. The rating is denominated in points
+above/below average, where zero is average\cite{srs}.
+
+\paragraph{Playoffs} The National Basketball Association (NBA) playoffs are a
+best-of-seven elimination tournament among 16 teams in the Eastern Conference
+and Western Conference (called divisions, pre-1970), ultimately deciding the
+winner of the NBA Finals\cite{playoffs}.
+
+\paragraph{Jersey} The shirt worn by basketball players.
 

### 2016-05-21 paper/paper.tex
Reworked related work

@@
 
-\todo{Rework realted work}
 \section{Related Work}\label{sec:literature}
+Recently, there has been an increase in the interest in visualizations of sports
+data. This is best exemplified by theworkshop on sports data visualization at
+the IEEE Visualization Conference in 2013\cite{ieeevis}.
 
-On the IEEE Visualization Conference in 2013, the first workshop on sports data
-visualization was held \cite{ieeevis}. There is shown that the area of sport
-produces large amounts of data, which can be used for building visualizations
-and making analyses of the data. In the past visualizations of sports existed
-already, but there were not a lot of visualizations, and they were simple in
-the sense that they just represented the data without trying to give insights
-or to approach the data on a different way. There are some, but limited,
-sports-specific visualizations, like the basketball shot chart. This is a
-visualization frequentely seen when looking for visualizations about the NBA.
-Some shot chart vizualisations try to be different than others, but the general
-design and idea stays roughly the same. 
-% 2 goeie papers die hierin verwerkt kunnen worden
-%http://www.sloansportsconference.com/wp-content/uploads/2012/02/Goldsberry_Sloan_Submission.pdf
-%http://vis.berkeley.edu/courses/cs294-10-sp10/wiki/images/archive/a/a2/20100511091342!Stephen_Chu_-_Information_Visualization_in_the_NBA_-_The_Shot_Chart_16.pdf
+There are a number of traditional basketball visualizations. One of the most
+popular ones is the shot chart. This visualizes the shots of a player or team
+around the hoop. It provides insight into the shooting ability of a
+player\cite{goldsberry,stephenchu}.
 
-With our visualization we try to make it easier to interpret a collection of
-data about the NBA, and try to let users find patterns and insights in the
-data. A same kind of incentive was the start for \cite{cox2006sportsvis} who
-created a visualization to help people discover meaning in statistics generated
-during sporting events in the case of baseball.
+This view features heavily in Peter Beshai's Buckets\cite{peterbeshai}. This is
+an extensive visualization of NBA player data. It provides not only the player's
+shot chart for various years, but his shooting signature and several line graphs
+as well. It is strongly focused on individual player's shooting abilities.
 
-(From bubble view)
-A more traditional visualization of a play-off is the team double elimination
-bracket.  Although this is a more common representation , and there are some
-nice desings like in this NBA 2012 play-offs
-visualisation~\cite{tournamentladder} where a sunburst visualisation is used,
-we wanted to create a more compact version by eliminating the recurring
-representation of a team. We considered enclosed circles, as used in this
-visualisation of the wimbledon tournament roster~\cite{enclosedcircles}.
-Although the enclosed circles is a compact representation as well, we have
-continued with our own design as we believe this view is more original and
-innovative. Furthermore we are convinced that our view gives a good
-illustration of the end positions. It's generally assumed that the team that is
-on top of the ``ladder'' is the winning team. We didn't encouter a similar view
-when searching for alternatives to visualize a play-off or championship. A new
-visualization has the advantage that it will trigger users as well.
+Visualizations focusing on basketball teams and their performance are
+significantly more difficult to find. For the playoffs, there is the traditional
+playoffs bracket~\cite{tournamentladder}. This bracket structure has the
+advantage of being well known and as such, not requiring much explanation. It is
+flawed however. It duplicates much information by displaying teams multiple
+times.
 
-(From field view)
-One good alternative was the shooting signature.\cite{peterbeshai} However,
-creating such a visualization is again a project on itself and due to time
-constraints we left it aside. Other possibilities such as a heatmap or spider
-web didn't fulfill our need to give the visitor a clear overview of how a player
-performed compared to its peergroup.
+An interesting alternative is the sunburst tournament
+visualization\cite{sunburst}. As the name suggests, this visualizes the
+tournament as a sunburst, with the winning team taking the part of the circle
+closer to the center. This does not eliminate the problem of the recurring teams
+however.
+
+For the wimbledon tennis tournament, an enclosed circles visualization was
+created~\cite{enclosedcircles}. This visualization eliminates the repeating
+occurrences of the same team. However, it wastes a lot of space.
+
+England's Permier League results were visualized by Ami Sedghi for the
+Guardian\ref{premierleague}. This visualization shows the team's rankings as
+line graphs and allows highlighting of individual teams.
+
+Tan et al. created an adaption of the traditional tournament bracket to better
+support moden interaction\cite{adaptivitree}. The tool, AdaptiviTree, changes
+the brackets shape and adds colored lines to show available information as the
+user pciks his own bracket.
+
+An alternative to the traditional bracket structure can be found in the
+visualization of the race to the white house in 2012 made for the New York Times
+by Mike Bostock and Shan Carter\cite{whitehousepath}. They use a tree-like
+structure where each split represents a state being one by one candidate or the
+other. The interactivity is very important here as it allows users to highlight
+one possible path out of many.
+
+The entire history of the NBA and the best teams were analyzed by
+fivethirtyeight~\cite{fivethirtyeight}. They calculated an elo score for all
+teams an visualized them on line graphs.
+
+Pagno et al. provide a number of different visualizations\cite{starplots}. The
+most interesting one is the starplot. This plot several statistics for a team on
+an axis pointing radially outward from a common center. These points are then
+connected. It allows the user to get basic information about a team or player at
+a glance. Because of this speed of conveying information, it is also a prevalent
+visualization in sports games.
 
@@
 \subsection{Technology}
-To create the visualization, the d3 javascript framework~\cite{d3.js} was used
-in combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3
-is based on HTML, CSS, Javascript and SVG. This allows easy access to the visualization as
-most modern browsers are capable of handling these webstandards.  The choice not
-to use the d3 framework for the entire visualization was made to ease the layout
-configuration of the visualization. Instead, html was used to do the global
-layout of the visualization. We opted to structure the site with multiple divs
-as containers. To each container we allocate a svg with a specific
-visualization.
+To create the visualization, the d3 javascript framework~\cite{d3} was used in
+combination with html5~\cite{html5} and jquery~\cite{jquery}.  Note that D3 is
+based on HTML, CSS, Javascript and SVG. This allows easy access to the
+visualization as most modern browsers are capable of handling these
+webstandards.  The choice not to use the d3 framework for the entire
+visualization was made to ease the layout configuration of the visualization.
+Instead, html was used to do the global layout of the visualization. We opted to
+structure the site with multiple divs as containers. To each container we
+allocate a svg with a specific visualization.
 

### 2016-05-24 paper/paper.tex
Some changes paper

@@
 
-\usepackage[
-    backend=biber,
-    style=draft
-]{biblatex}
+\usepackage[backend=biber, style=draft]{biblatex}
+%\usepackage{custom}
 
@@
 \begin{abstract}
+\todo{Abstracts should be about 150 words}
     Visualization of sports data is very common in recent years as access to
@@
 available, gathering useful insights from this data can be surprisingly
-challenging. In our work we try to present this data in an innovative and
+challenging. In our work we try to present this data in an innovative but still
 intuitive way, allowing the user to draw meaningful conclusions.
 
+\todo{reference to sections should use the name and not a number because the numbers are not shown, so they are not usable}
 In section~\ref{sec:goal} we describe the goal of the visualization and the
@@
 visualizations.  This includes both visualizations of NBA or other sports data,
-as well as visualizations tackling an isue encountered during the development of
+as well as visualizations tackling an issue encountered during the development of
 our visualization. In section~\ref{sec:visualization} we describe the
@@
 The visualization's goal can be best summarized by the following sentence:
-``Why did a certain NBA-team win the Championship?'' The visualization focuses
+``Why did a certain team win the NBA Championship?'' The visualization focuses
 on exposing relations between team performance and their player roster over
@@
 The data visualized is a subset of the data available on basketball statistics
-site basketball-reference\cite{basketball-reference}. A wide range of data is
-available on this site. This includes common basketball performance statistics,
-such as field goals, percentage of shots scored, shooting distance, minutes
-played per season, number of personal fouls made, salaries and much more. All of
-this data is available as time-series. These
-statistics are avalable for each player individually for each season. Aggregated
-data for entire teams and for entire seasons is also available. Additionally,
-individual game statistics are also available, including playoff games.  In our
-visualization we only use data from 1984 onwards. This is done because of rule
-changes in the NBA which changed the number of teams competing and the
+site basketball-reference\cite{basketball-reference}.
+It is a wide range of data including common basketball performance statistics, such as field goals, percentage of shots scored, shooting distance, minutes played per season, number of personal fouls made, salaries and much more .
+All of this data is available as time-series.
+These statistics are available for each player individually for each season.
+Aggregated data for entire teams and for entire seasons is also available.
+Additionally, individual game statistics are also available, including playoff games.
+In our visualization we only use data from 1984 onwards.
+This is done because of rule changes in the NBA which changed the number of teams competing and the
 competition's structure. To simplify implementation of our visualization, only
@@
 
+\todo{Figure is still in dutch}
 \begin{figure}
@@
 Recently, there has been an increase in the interest in visualizations of sports
-data. This is best exemplified by theworkshop on sports data visualization at
+data. This is best exemplified by the workshop on sports data visualization at
 the IEEE Visualization Conference in 2013\cite{ieeevis}.
@@
 the brackets shape and adds colored lines to show available information as the
-user pciks his own bracket.
+user picks his own bracket.
 
@@
 The entire history of the NBA and the best teams were analyzed by
-fivethirtyeight~\cite{fivethirtyeight}. They calculated an elo score for all
+fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score \todo{explanation of the elo score?} for all
 teams an visualized them on line graphs.
@@
 
-This technique follows the standards of information visualization. It provides
+The technique used, follows the standards of information visualization. It provides
 an overview, allows zooming and filtering and provides details on
@@
 
-The organization of these three views is in a vertical layout with fixed
-scrolling. Navigation with the keyboard is supported as well.
+All three views are discussed in more detail below. The three views are
+available to the user on a single webpage, where they are organized in a 
+vertical layout with fixed scrolling. Navigation with the keyboard is 
+supported as well. The user start on the play-off view and is able to 
+navigate to the statistics and team views by scrolling down or clicking 
+on teams. A different season can be selected with the left and right
+arrows or by clicking on the timeline. The timeline is a fixed part of the
+visualization. It is visible in the three different views. Navigation 
+with the keyboard for switching between the three different views is 
+supported as well.
 
-The choice was made to use a thematic background instead of opting for the
-traditional minimalistic approach to information visualization. This was done to
-improve the attractiveness of the visualisation on the one hand and the users'
-memory of it on the other hand\cite{aesthetics}.
-
-All three views are discussed in more detail below. The three views are
-available to the user on a single webpage. The user start on the play-off view
-and is able to navigate to the statistics and team views by scrolling down or
-clicking on teams. A different season can be selected with the left and right
-arrows or by clicking on the timeline. The timeline is a fixed part of the
-visualization. It is visible in the three different views.
+We did not opt for the traditional minimalistic approach used in information 
+visualizations. The data-ink design principles were kept in mind, but we also 
+focused on aesthetics instead of maximizing the data-ink ratio. So the choice 
+was made to use a thematic background. This was done to improve the 
+attractiveness of the visualisation on the one hand and the users' memory of 
+it on the other hand\cite{aesthetics}.
 
@@
 This view informs the user of the roster and end positions in the NBA play-offs
-for a selected season, in general or for a selected team. The view consists of
+for a selected season, in general or for a selected team. The play-off view is 
+shown in figure~\ref{fig:playoffviewwithcontext}.The view consists of
 two parts: 
@@
 \end{itemize}
-The play-off view is shown in figure~\ref{fig:playoffviewwithcontext}.
 
@@
 borders are gold, silver or bronze if the team ended respectively first, second
-or third in the regular competition. Teams that didn't end in a podium place
+or third in the regular competition. Teams that didn't end on a podium place
 during the regular competition have a red or blue stroke color indicating the
 team's region, respectively the western or eastern conference. The context
-section also contains a link to an ``about'' page where the user can get an
+section also contains a link to the ``About'' page where the user can get an
 explanation of the visualization with its different views, can get some
 background information about the NBA and the terminolgy used, and can check the
-source of our used data. If the user hovers over or selects a team in the bubble
-section, the context section shows more information about that team, as shown in
-Figure~\ref{fig:playoffviewteamselected}. The information shown includes the
-Simple Rating System score (SRS), the end position in the play-offs and the end
-position in the regular competition.
+source of the data used in the visualization. If the user hovers over or selects 
+a team in the bubble section, the context section shows more information about that 
+team, as shown in Figure~\ref{fig:playoffviewteamselected}. The information shown 
+includes the Simple Rating System score (SRS), the end position in the play-offs 
+and the end position in the regular competition.
 
@@
 
-The curved lines between teams represent that those teams have encountered each
-other in the playoffs. Each row represents an end position in the play-offs. The
-higher a team is, the further it has come.  From the top to the bottom the rows
-represent the winner of the final (rank 1), loser of the final (rank 2), losers
-of the semi-finals (rank 3), losers of the quarter-finals (rank 4), losers of
-the eighth finals (rank 5). The last row represents the teams that did not make
-it to the play-offs (rank 6).
+The curved lines between teams represent the fact that those teams have 
+encountered each other in the playoffs. Each row represents an end position 
+in the play-offs. The higher a team is, the further it has come.  From the 
+top to the bottom the rows represent the winner of the final (rank 1), loser 
+of the final (rank 2), losers of the semi-finals (rank 3), losers of the 
+quarter-finals (rank 4), losers of the eighth finals (rank 5). The last row 
+represents the teams that did not make it to the play-offs (rank 6).
 
 This simple presentation intuitively puts the winning teams higher than the
-losing teams. The winner of a certain game can be easily deriving from their
-relative vertical position. It is also clear that each team neads to have played
-and won a series of matches against one other team at every level below
-them~\ref{fig:playoffviewteamselected}. This view is further reinforced by
-highlighting a team and its matches and opponents when hovering over it.
-Additional team information is then also shown in the context part.
+losing teams. The winner and loser of a certain game can be easily derived from 
+their relative vertical position. It is also clear that for a team to be on a 
+higher level, the team needs to have played and won a series of matches against 
+one other team at every level below them~\ref{fig:playoffviewteamselected}. 
+This view is further reinforced by highlighting a team and its matches and 
+opponents when hovering over it. Additional team information is then also shown 
+in the context part.
 
@@
 combination of the logo's of the teams with the colors resulted in an overload
-of informaiton and made it more difficult to compare the end ranking of
+of information and made it more difficult to compare the end ranking of
 different teams.
 
-As an alternative we tried the same view without the logo's. This was deemed
-inconvenient because you cannot instantly see which team is represented by which
-circle. Therefore we've chosen to only use the teams' logos.
+As an alternative we tried the same view without the logo's. It turned out this 
+was inconvenient because you could not instantly see which team is represented 
+by which circle. Therefore we've chosen to only use the teams' logos.
 
 Our initial play-off view was too dense because of the connections between teams
-as well. It connected teams that played against each other directly. When
+as well. Teams that played against each other were connected directly. When
 searching for alternatives to reflect on our choice we discovered a similar
-visualization in a totally different context\cite{whitehousepath}.  This
-visualization gives a cleaner view of the competitors for a specific team
-compared to our original sketch. Hence we decided to adapt our visualization. We
-added an intermediate step between two teams to more clearly indicate how teams
-competed to become NBA champion.
+visualization in a totally different context\cite{whitehousepath} as explained in 
+the section \ref{sec:literature}. With the idea of this visualization in mind, our 
+visualization gives a cleaner view of the competitors for a specific team compared 
+to our original sketch. Hence we decided to adapt our visualization. We added an 
+intermediate step between two teams to more clearly indicate how teams competed to 
+become the NBA champion. \todo{meaning of last sentence is not clear}
 
@@
         selected team
-    \item Arrow section: this section illustrates why a statistic changed
+    \item Arrow section: this section illustrates why a statistic has changed
         compared to the previous season
@@
 
-\subsubsection{Arrow section of selected statistic}
-This section illustrates a change for a chosen statistic for the selected team
-compared to the previous season. This gives a user more insight into why a
-statistic changed over time and how this change has impacted team performance.
-The selected team is represented by a circle in the same way as explained in the
-play-off view. The change for a statistic is represented by three arrows:
+\subsubsection{Arrow section}
+This section illustrates a change for a chosen statistic for the selected team in 
+the selected season compared to the previous season. This gives a user more 
+insight into why a statistic changed over time and how this change has affected 
+team performance. The selected team is represented by a circle in the same way as 
+explained in the play-off view. To give the user a better interpretation of the SRS 
+value of the circle, three other reference circles are drawn. These are 
+transparant with a lightgrey border. They show the maximum, average and minimum SRS 
+value of the NBA in the selected season. The change for the chosen statistic is 
+represented by three arrows:
 \begin{itemize}
     \item Arrow on the left pointing towards the circle: indicating the
-        influence of players who joined the team 
+        influence of players who joined the team.
     \item Arrow on the right of the circle: indicating how the team internally
-        changed, e.g. by training its players.
+        changed, e.g. by training its players
     \item Arrow on the left pointing away from the circle: indicating the
@@
 \todo{add statistics view image here}
-The jerseys next to the arrows are the jerseys of players that had the biggest
-contribution to that arrow. On the left, different statistics , including player
-PER can be chosen from. Below this, player information is shown when hovering
+The value of the change in statistic is shown next to the arrow. The number of 
+players responsible for the change is shown between brackets. Their are shown 
+jerseys next to the arrows as well. These are the jerseys of the two players that 
+had the biggest contribution to the influence represented by that arrow.  
+On the left, there is a possibility to chose a statistic. The possibilities 
+are the PER value, the age, the effective field goal percentage, the offensive rating 
+and the defensive rating. Below this, player information is shown when a user hovers
 over a player's jersey.
 
-\subsubsection{Selected statistics}
-Below the arrow section some selected statistics are shown over
-time.  The line charts enables a user to identify peaks or drops that
-influenced the outcome of a team is this season, following seasons or previous
-seasons. A user interacts with the visualization by shifting the bar indicating
-the current year displayed. By doing this the view gets updated and the user can
-see the impact on the team in the context section and why the statistic changes
-in the arrow section. This view provides a lot of additional information to
-users by allowing them to see correlations between statistics and phenomena
-quickly and easily.
+\subsubsection{Selected statistics section}
+Below the arrow section, multiple line and area charts show how the selected 
+team has performed over the years. The different statistics shown are the 
+total PER of the whole team, the league rank, the playoff rank, the average age 
+of the team, the audience, the SRS-value of the team, the offensive rating and 
+the defensive rating of the team. A user interacts with the visualization by 
+shifting the bar indicating the current year displayed or the timeline. By 
+doing this the view gets updated and the user can see the impact on the team 
+in the context section and why the statistic changes in the arrow section. 
+These charts enable the user to identify peaks or drops in a certain statistic. 
+By interacting with the visualization, previous seasons can be inspected to try 
+to find explanations or following seasons can be inspected to see if a team 
+responds in a specific way or what the consequences are. This view provides 
+a lot of additional information to users by allowing them to see correlations 
+between statistics and phenomena quickly and easily.
+
+\subsubsection{Previous designs}
+This view has not changed much compared to the initial design. The only difference 
+was the position of the arrow and jerseys for the players who stayed, which were 
+originally in the team circle, and so the logo was not showed. 
+The design of the arrows did not change. The incoming and outgoing 
+arrows are maybe more intuitevly than the arrow for the stayed players, but we did 
+not find any alternatives we could use.
 
 \subsection{The team view}
+The team view gives a user a detailed view on the strength of the team on the 
+different field positions. The team view is showed in Figure~\ref{fig:teamview}. 
+The view consists of 3 coordinated parts:
+\begin{itemize}
+    \item Context section: the top part informs the user of the context of the
+        selected team
+    \item Field section: this section illustrates why a statistic has changed
+        compared to the previous season
+    \item Selected statistics section: team statistics and how they evolve over
+        time
+\end{itemize}
+The context section and the selected statistics section are fixed parts in the 
+statistics and team view, and are therefore not explained again for this view.
+
+\subsubsection{Field section}
 This view shows each player of the team on its field positions. Players are
 represented by their jerseys. Jerseys aligned next to each other share the same
-position, which is indicated as well. The size of a jersey would illustrate a
-players performance on this spot. This view as well as the statistics view
-should give the user the ability to search for explanations why a team is is
-performing better or worse over seasons. On the other hand, a user can see what
-impact a change in team characteristics has on its overall performance. This
-view should give a user insight on how strong or weak a particular team is for a
-specific field position. A user can then scroll through time to see how field
-positions evolve and what the influence is over time.
+position, which is indicated as well. The size of a jersey illustrates a
+players performance according to the chosen statistic. The possible statistics 
+are shown on the left and are the same as those in the arrow section of the 
+statistics view. As well as the statistics view, this view should give the user 
+the ability to search for explanations why a team is performing better or worse 
+over seasons. On the other hand, a user can see what impact a change in team 
+characteristics has on its overall performance. This view should give a user 
+insight on how strong or weak a particular team is for a specific field position. 
+A user can then scroll through time to see how field positions evolve and what 
+the influence is over time.
 
@@
 Initially we designed this view more elaborate. When the user would click on a
-jersey, specific player information would be shown. A number of player
-statistics would be shown as small multiples. A bell curve should have informed
-how a player scores compared to his team, other players in the league on this
+jersey, specific player information would be shown together with a number of player
+statistics in the form of small multiples. A bell curve should have informed
+how a player scores compared to his team, other players in the league on his
 position or all other players in the league. We considered other charts for the
 small multiples as well. A box plot or bullet was less clear for us to indicate
-where a player stands compared to the group he was compared with. Alternatives
-for the small multiples were evaluated as well. This part of the visualization
-was deemed less important than the others and has not been fully developed
-because of time and resource constraints. \todo{sketch of small multiples
+where a player stands compared to the group he was compared with. This part of the 
+visualization was not our main focus. Because of the lesser importnace and because of time and resource constraints, it has not been fully developed. \todo{sketch of small multiples
 here}
@@
 \subsection{Improving the result}\todo{Is this necessary? I don't know what to
-write here}
+write here.}
 Talk about how we improved the data-ink ration in the end..
@@
 As with most projects, the result is never 'finished', meaning that there are
-always improvements possible but due to timing or resource constraints they
-haven't all been implemented. We suggest a number of features that might improve
-the visualization:
+always extensions and iprovements possible but due to timing or resource constraints they haven't all been implemented. We suggest a number of features that might improve the visualization:
 \begin{itemize}
@@
         for us humans with a short term memory to make the comparison. This
-        would have been especially useful in the transitions between different
+        would also have been very useful in the transitions between different
         years on the bubble view.
@@
 allocate a svg with a specific visualization.
+\todo{sketch of how the page is structured here}
 
-\todo{sketch of how the page is structured here}
-When working with d3.js we encountered some obstacles:
+When working with d3.js we encountered one main obstacle:
 \begin{itemize}
@@
 \section{Lessons learned}\label{sec:discussion}
-We've learned a couple of lessons during the creation of this visualization:
+We've learned a couple of things during the creation of this visualization:
 \begin{itemize}
@@
 In order to have a good visualization as a whole, each component should be a
-clear and informative visualization on itself. During the project, we started noticing that most
-components in a visualization lack references and hence are not informative
-enough. Eg. each small multiple of a statisctic in the statistics view on itself
-gives sufficient information to stand on itselfs. The bar in the small multiple
-indicates the score in the selected season and enables the user to evaluate that
-score over time. Have there been better scores, or worse scores. Is this score
+clear and informative visualization on itself.
+During the project, we started noticing that most components in a visualization lack references and hence are not informative enough.
+Eg.
+each small multiple of a statisctic in the statistics view on itself gives sufficient information to stand on itselfs.
+The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
+Have there been better scores, or worse scores.
+Is this score
 part of an upward movement during years? This should be weighed against the
-duplication of data and use of screen space. Coordinates views alleviate this
+duplication of data and use of screen space. Coordinated views alleviate this
 need significantly as well, as each component is strongly supported by other
@@
 designs. Although we are satisfied with our result and we are convinced it was
-the right choice to reach our goal, it had some draw backs. We noted that other
-teams had more freedom in exploring multiple different visualizations which
-enabled them to evaluate different options with trial and error. Much of the
-time developing was spent on the technical details for these custom
-visualizations. This was especially true for the bubble component of the
-playoffs view. No out of the box solution for this could be found so everything
-from the determination of the positions of the bubbles to how to connect them
-had to be developed by us.
+the right choice to reach our goal, it had some draw backs.
+If we would have not started with our custom view, we would have had more freedom in exploring multiple different visualizations and we would be able to evaluate different options with trial and error.
+Instead much of the time developing was spent on the technical details for these custom visualizations.
+This was especially true for the bubble component of the play-off view.
+No out of the box solution for this could be found so everything from the determination of the positions of the bubbles to how to connect them had to be developed by us.
 
@@
 %\bibliographystyle{SIGCHI-Reference-Format}
+
 \printbibliography
-%TODO References should be in alphabetical order and represented by numbers in the text.
+\todo{References should be in alphabetical order and represented by numbers in the text. Style should be set to style{SIGCHI-Reference-Format}, this moment it is set to the style draft}
 

### 2016-05-25 paper/paper.tex
Corrections and additions paper

@@
 \usepackage{textcomp}
+\usepackage{nameref}
 
@@
 
-\todo{reference to sections should use the name and not a number because the numbers are not shown, so they are not usable}
-In section~\ref{sec:goal} we describe the goal of the visualization and the
-target audience. In section~\ref{sec:data} we describe the data used, including
-its origins, advantages and limitations. In section~\ref{sec:literature} we give
-an overview of related literature and web resources, including related
+In the \nameref{sec:goal} section we describe the goal of the visualization and the
+target audience.
+In the \nameref{sec:data} section, we describe the data used, including its origins, advantages and limitations.
+The \nameref{sec:literature} section gives an overview of related literature and web resources, including related
 visualizations. This includes both visualizations of NBA or other sports data,
 as well as visualizations tackling an issue encountered during the development of
-our visualization. In section~\ref{sec:visualization} we describe the
-visualization itself. We give an overview of the different stages of development
+our visualization. In the \nameref{sec:visualization} section, the visualization itself is described. We give an overview of the different stages of development
 of the visualization, as well as the major design decisions.
-Section~\ref{sec:discussion} discusses potential improvements of the final
-visualization and lessons learned from the project. We conclude in
-section~\ref{sec:conclusion}. Appendix~\ref{sec:terminology} contains
-definitions for common basketball terms used in this paper.
+The \nameref{sec:lessonslearned} section describes some interesting lessons we have learned from the project and in the \nameref{sec:futurework} section we discuss potential  improvements of the final visualization.
+We conclude in the
+\nameref{sec:conclusion} section. Appendix~\ref{sec:terminology} contains definitions 
+for common basketball terms used in this paper.
 
@@
 final preprocessing step combined all data into one json file. A model of the
-available data can be seen on figure~\ref{fig:data}.
+available data can be seen on figure~\ref{fig:data} \todo{is die afbeelding nog wel juist? Is die nuttig?}.
 
-\todo{Figure is still in dutch}
 \begin{figure}
@@
 There are a number of traditional basketball visualizations. One of the most
-popular ones is the shot chart. This visualizes the shots of a player or team
-around the hoop. It provides insight into the shooting ability of a
+popular ones is the shot chart.
+This visualizes the shots made by a player or team.
+Different methods can be used, going from a scatterplot to hexagonal charts where the shots are grouped into hexagonal regions, to a heatmap.
+This is often done for the the area in and around the three-point line, but it can als be done for the whole attack side of the field, as seen in the Vorped visualisation\cite{vorped}.
+Because shots of the own side only happen rarely, it's not shown in shot charts.
+Shot chart provides insight into the shooting ability of a
 player\cite{goldsberry,stephenchu}.
@@
 England's Permier League results were visualized by Ami Sedghi for the
-Guardian\ref{premierleague}. This visualization shows the team's rankings as
+Guardian\cite{premierleague}. This visualization shows the team's rankings as
 line graphs and allows highlighting of individual teams.
@@
 The entire history of the NBA and the best teams were analyzed by
-fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score \todo{explanation of the elo score?} for all
-teams an visualized them on line graphs.
+fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score\cite{eloraring} 
+for all teams an visualized them on line graphs.
+
+In the article written by Johannes Becker\cite{nbaempires}, the Simple Rating System 
+score (SRS score) is used to compare teams over the last 40 seasons in the NBA. The 
+SRS score is represented by a color hue. For comparing quantative values, this is 
+not the most ideal characteristic to use as explained by John Mackinglay\cite{automatingdesign}.
+Nevertheless, some interesting findings are found and presented.
+By giving a green stroke to the rectangle of the champion of a season,
+it's easy to see which teams have won the NBA championship frequently. This way 
+they have divided the last 40 seasons in three eras being the Los Angeles Lakers 
+versus the Boston Celtics era, the Michael Jordan era and the Tim Duncan era.
 
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=1.0\columnwidth]{figures/playoffviewwithcontext}
-  \caption{The play-off view without a team selected.}~\label{fig:playoffviewnoteam}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1.0\columnwidth]{figures/playoffviewteamselected}
-  \caption{The play-off view with a team selected.}~\label{fig:playoffviewteam}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1.0\columnwidth]{figures/statisticsview}
-  \caption{The statistics view.}~\label{fig:statisticsview}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1.0\columnwidth]{figures/teamview}
-  \caption{The team view.}
-  \label{fig:teamview}
-\end{figure}
-
 The technique used, follows the standards of information visualization. It provides
@@
 
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\columnwidth]{figures/playoffviewwithcontext}
+  \caption{The play-off view without a team selected.}~\label{fig:playoffviewnoteam}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\columnwidth]{figures/playoffviewteamselected}
+  \caption{The play-off view with a team selected.}~\label{fig:playoffviewteam}
+\end{figure}
+
 \subsection{The play-off view}
@@
 team, as shown in Figure~\ref{fig:playoffviewteamselected}. The information shown 
-includes the Simple Rating System score (SRS), the end position in the play-offs 
-and the end position in the regular competition.
+includes the SRS score, the end position in the play-offs and the end position in 
+the regular competition.
 
@@
 ranking in their region. The hue was used to show the league they were part of.
-According to the Ranking of Perceptual Tasks~\cite{perceptualranking} this is a
+According to the Ranking of Perceptual Tasks~\cite{automatingdesign} this is a
 good property for ordered values like the regular competition ranking. But the
@@
 we decided to not implement this map view.
+
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\columnwidth]{figures/statisticsview}
+  \caption{The statistics view.}~\label{fig:statisticsview}
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\columnwidth]{figures/teamview}
+  \caption{The team view.}
+  \label{fig:teamview}
+\end{figure}
+
 \subsection{The team view}
@@
 
-\section{Future work}
+\subsection{Technology}
+To create the visualization, the d3 javascript framework~\cite{d3} was used in
+combination with html5~\cite{html5} and jquery~\cite{jquery}. Note that D3 is
+based on HTML, CSS, Javascript and SVG. This allows easy access to the
+visualization as most modern browsers are capable of handling these
+webstandards. The choice not to use the d3 framework for the entire
+visualization was made to ease the layout configuration of the visualization.
+Instead, html was used to do the global layout of the visualization. We opted to
+structure the site with multiple divs as containers. To each container we
+allocate a svg with a specific visualization.
+\todo{sketch of how the page is structured here}
+
+When working with d3.js we encountered one main obstacle:
+\begin{itemize}
+        \item A synchrounous call need to be combined with synchronous calls.
+            More specific, when a visitor uses the timeline to scroll through
+            time, the whiping of previous visualizations was not synced with the
+            creation of the visualisation. This resulted in whiping sections
+            when a visualization was not created yet. Hence the view resulted in
+            multiple figures overlapping.
+\end{itemize}
+
+\section{Lessons learned}\label{sec:lessonslearned}
+We've learned a couple of things during the creation of this visualization:
+\begin{itemize}
+    \item Find a story to tell with your visualization
+    \item Each separate component should be a sufficient informative visualization on itself
+    \item Creating a custom visualization costs time and opportunities
+\end{itemize}
+In what follows we will discuss our experience in more detail.
+
+\subsection{Story telling}
+At the beginning of this project we were able to design the 3 separate views
+quite quickly. Our main concern was, when integrating these views as a whole, would 
+we be able to see the relations and consequences we thought we would see and if 
+we could draw the conclusions we hoped for. The relation we thought would be visible 
+was that the total PER value of a team would influence the end positions in the NBA championship, both the league rank as the play-off rank. We expected that teams 
+that had a big increase in total PER value, due to a good transferpolicy in a 
+season, would benefit in one of the following seasons.
+It was not possible to draw this conclusion and according to us the reason for this is that a whole NBA team contains an average of 15 players.
+Some players may play a lot more than others and
+are therefore much more important to the team.
+This observation is why we could not draw our hoped conclusion, but at the same time we found it so interesting and we wondered if we could find an example to support this observation.
+The story is about
+the Chicago Bulls, the team of Michael Jordan who is seen as the greatest basketball 
+player of all time.
+In 1984, Jordan arrives at the Chicago Bulls and in his first season he is already the top player of his team, as seen by the size of his jersey for the PER statistic.
+In the season 1990-1991 and the two following seasons, the
+Chicago Bulls win the NBA championship. The ``three-peat'' as it is called was done 
+with almost the same core team. Next to Michael Jordan, there were around five 
+players that were part of the team and were roughly the best on their position in 
+the team, for example Scottie Pippen. We can see that in the seasons before this 
+``three-peat'' this team was formed and already played with each other. After that 
+last season form those three consecutive wins, Michael Jordan retires and the team 
+is not able anymore to achieve the same.
+In the middle of the season 1994-1995, Michael Jordan came out of retirement.
+What follows is a similar story to the previous one.
+In the season 1995-1996 and the two following seasons, the Chicago
+Bulls again achieve to win the NBA chamionship three times in a row, again with a  team consisting of around five core players. Once again is Michael Jordan the star 
+player of the team, but also Scottie Pippen is still one of the best players of the team. So to become the champion, it seems like a team needs to have a default team 
+with around five core players and one or two star players.
+
+%Our main problem was to integrate these views as a whole. When we
+%created drafts of two views being the play-off view and the statitics view, we
+%suddenly noted patterns. The Golden State Warriors became NBA champion a couple
+%of years after a great uplift in SRS score. This was because of new players
+%joining the team in 2009. In the years following, the team kept on attracting
+%talented players and increased the inherent team score to become NBA champion in
+%2015. With the team view, one can even notice that the team gets stronger in
+%center team positions during this period. Users are able to explore the data for
+%patters. On top of that, we provide the insight on how the changes were impacted
+%by changes in teamplayers.
+
+\subsection{Each component should be informative}
+In order to have a good visualization as a whole, each component should be a
+clear and informative visualization on itself.
+During the project, we started noticing that most components in a visualization lack references and hence are not informative enough.
+Eg.
+each small multiple of a statisctic in the statistics view on itself gives sufficient information to stand on itselfs.
+The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
+Have there been better scores, or worse scores.
+Is this score
+part of an upward movement during years? This should be weighed against the
+duplication of data and use of screen space. Coordinated views alleviate this
+need significantly as well, as each component is strongly supported by other
+components.
+
+\subsection{Custom visualizations}
+Before exploring what is out there in the d3.js world, we made our own sketches.
+This is how we came up with the 3 custom views that make up our visualization.
+When comparing our design with alternatives we could not find something that
+satisfied our needs. Nonetheless, they gave us inspiration to finetune our
+designs. Although we are satisfied with our result and we are convinced it was
+the right choice to reach our goal, it had some draw backs.
+If we would have not started with our custom view, we would have had more freedom in exploring multiple different visualizations and we would be able to evaluate different options with trial and error.
+Instead much of the time developing was spent on the technical details for these custom visualizations.
+This was especially true for the bubble component of the play-off view.
+No out of the box solution for this could be found so everything from the determination of the positions of the bubbles to how to connect them had to be developed by us.
+
+\section{Future work}\label{sec:futurework}
 As with most projects, the result is never 'finished', meaning that there are
@@
         to compare two teams in the same year, or in two different years.
+    \item We could focus more on the core team, i.e. five or six players who seems 
+        to be the most important for the results of a team.
 \end{itemize}
-
-\subsection{Technology}
-To create the visualization, the d3 javascript framework~\cite{d3} was used in
-combination with html5~\cite{html5} and jquery~\cite{jquery}. Note that D3 is
-based on HTML, CSS, Javascript and SVG. This allows easy access to the
-visualization as most modern browsers are capable of handling these
-webstandards. The choice not to use the d3 framework for the entire
-visualization was made to ease the layout configuration of the visualization.
-Instead, html was used to do the global layout of the visualization. We opted to
-structure the site with multiple divs as containers. To each container we
-allocate a svg with a specific visualization.
-\todo{sketch of how the page is structured here}
-
-When working with d3.js we encountered one main obstacle:
-\begin{itemize}
-        \item A synchrounous call need to be combined with synchronous calls.
-            More specific, when a visitor uses the timeline to scroll through
-            time, the whiping of previous visualizations was not synced with the
-            creation of the visualisation. This resulted in whiping sections
-            when a visualization was not created yet. Hence the view resulted in
-            multiple figures overlapping.
-\end{itemize}
-
-\section{Lessons learned}\label{sec:discussion}
-We've learned a couple of things during the creation of this visualization:
-\begin{itemize}
-    \item Find a story to tell with your visualization
-    \item Each separate component should be a sufficient informative visualization on itself
-    \item Creating a custom visualization costs time and opportunities
-\end{itemize}
-In what follows we will discuss our experience in more detail.
-
-\subsection{Story telling}
-At the beginning of this project we were able to design the 3 separate views
-quite quickly. Our main problem was to integrate these views as a whole. When we
-created drafts of two views being the play-off view and the statitics view, we
-suddenly noted patterns. The Golden State Warriors became NBA champion a couple
-of years after a great uplift in SRS score. This was because of new players
-joining the team in 2009. In the years following, the team kept on attracting
-talented players and increased the inherent team score to become NBA champion in
-2015. With the team view, one can even notice that the team gets stronger in
-center team positions during this period. Users are able to explore the data for
-patters. On top of that, we provide the insight on how the changes were impacted
-by changes in teamplayers.
-
-\subsection{Each component should be informative}
-In order to have a good visualization as a whole, each component should be a
-clear and informative visualization on itself.
-During the project, we started noticing that most components in a visualization lack references and hence are not informative enough.
-Eg.
-each small multiple of a statisctic in the statistics view on itself gives sufficient information to stand on itselfs.
-The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
-Have there been better scores, or worse scores.
-Is this score
-part of an upward movement during years? This should be weighed against the
-duplication of data and use of screen space. Coordinated views alleviate this
-need significantly as well, as each component is strongly supported by other
-components.
-
-\subsection{Custom visualizations}
-Before exploring what is out there in the d3.js world, we made our own sketches.
-This is how we came up with the 3 custom views that make up our visualization.
-When comparing our design with alternatives we could not find something that
-satisfied our needs. Nonetheless, they gave us inspiration to finetune our
-designs. Although we are satisfied with our result and we are convinced it was
-the right choice to reach our goal, it had some draw backs.
-If we would have not started with our custom view, we would have had more freedom in exploring multiple different visualizations and we would be able to evaluate different options with trial and error.
-Instead much of the time developing was spent on the technical details for these custom visualizations.
-This was especially true for the bubble component of the play-off view.
-No out of the box solution for this could be found so everything from the determination of the positions of the bubbles to how to connect them had to be developed by us.
 

### 2016-05-26 paper/paper.tex
Change in figures
Small reference update

@@
 
-\usepackage[backend=biber, style=draft]{biblatex}
+\usepackage[backend=biber, style=numeric, firstinits=true, url=false, sorting=none]{biblatex}
 %\usepackage{custom}
@@
 The entire history of the NBA and the best teams were analyzed by
-fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score\cite{eloraring} 
+fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score\cite{elorating} 
 for all teams an visualized them on line graphs.
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=1.0\columnwidth]{figures/presidentcandidatesvisualization}
-  \caption{The path of president candidates in their road to become president.}~\label{fig:whitehousepath}
-\end{figure}
+%\begin{figure}
+%\centering
+%  \includegraphics[width=1.0\columnwidth]{figures/presidentcandidatesvisualization}
+%  \caption{The path of president candidates in their road to become president.}~%\label{fig:whitehousepath}
+%\end{figure}
 
@@
 \centering
-  \missingfigure{}
-%  \includegraphics[width=0.9\columnwidth]{figures/playoffviewfirstimplemantation}
+%  \missingfigure{}
+  \includegraphics[width=0.9\columnwidth]{figures/playoffviewfirstimplemantation}
   \caption{First implementation of the play-off view.}~\label{fig:firstimplementationplayoffview}
@@
 \printbibliography
+
 \todo{References should be in alphabetical order and represented by numbers in the text. Style should be set to style{SIGCHI-Reference-Format}, this moment it is set to the style draft}

### 2016-05-29 paper/paper.tex
Fixed citations mostly
Pretty decent now

@@
-\documentclass[chi_draft]{sigchi}
+\documentclass[]{sigchi}
 % Use this command to override the default ACM copyright statement
@@
 
-\usepackage[backend=biber, style=numeric, firstinits=true, url=false, sorting=none]{biblatex}
+\usepackage[backend=biber, style=numeric, sorting=nty, firstinits=true, url=false]{biblatex}
 %\usepackage{custom}
@@
 \begin{abstract}
-\todo{Abstracts should be about 150 words}
     Visualization of sports data is very common in recent years as access to
@@
 final preprocessing step combined all data into one json file. A model of the
-available data can be seen on figure~\ref{fig:data} \todo{is die afbeelding nog wel juist? Is die nuttig?}.
+available data can be seen on figure~\ref{fig:data}.
 
@@
 to our original sketch. Hence we decided to adapt our visualization. We added an 
-intermediate step between two teams to more clearly indicate how teams competed to 
-become the NBA champion. \todo{meaning of last sentence is not clear}
+intermediate connection point between two teams to more clearly indicate how teams competed to 
+become the NBA champion.
 
-\todo{Hierarchy = sorting = facilitate trends of winners
-position have a meaning in the play-off view
-when selecting circle => zoom/more detail
-Bubble principle of 
-- Symmetry (whole = play-off)
-- conectedness arc = games played
-Pre attentive characteristics
-- size of circles
-- hue to highligt selections or indicate increase decline arrows next view}
+%\todo{Hierarchy = sorting = facilitate trends of winners
+%position have a meaning in the play-off view
+%when selecting circle => zoom/more detail
+%Bubble principle of 
+%- Symmetry (whole = play-off)
+%- conectedness arc = games played
+%Pre attentive characteristics
+%- size of circles
+%- hue to highligt selections or indicate increase decline arrows next view}
 
@@
 \end{itemize}
-\todo{add statistics view image here}
 The value of the change in statistic is shown next to the arrow. The number of 
@@
 Initially we designed this view more elaborate. When the user would click on a
-jersey, specific player information would be shown together with a number of player
-statistics in the form of small multiples. A bell curve should have informed
-how a player scores compared to his team, other players in the league on his
-position or all other players in the league. We considered other charts for the
-small multiples as well. A box plot or bullet was less clear for us to indicate
-where a player stands compared to the group he was compared with. This part of the 
-visualization was not our main focus. Because of the lesser importnace and because of time and resource constraints, it has not been fully developed. \todo{sketch of small multiples
-here}
+jersey, specific player information would be shown together with a number of
+player statistics in the form of small multiples. A bell curve should have
+informed how a player scores compared to his team, other players in the league
+on his position or all other players in the league. We considered other charts
+for the small multiples as well. A box plot or bullet was less clear for us to
+indicate where a player stands compared to the group he was compared with. This
+part of the visualization was not our main focus. Because of the lesser
+importance and because of time and resource constraints, it has not been fully
+developed. A sketch including the small multiples view on the bottom right is
+shown on Figure~\ref{fig:smallmultsketch}.
 
-\subsection{Improving the result}\todo{Is this necessary? I don't know what to
-write here.}
-Talk about how we improved the data-ink ration in the end..
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\columnwidth]{figures/smallmultsketch}
+  \caption{A sketch of a previous design including the small multiples on the
+  bottom right.}
+  \label{fig:smallmultsktech}
+\end{figure}
 
@@
 allocate a svg with a specific visualization.
-\todo{sketch of how the page is structured here}
+%\todo{sketch of how the page is structured here}
 
@@
 
-\todo{References should be in alphabetical order and represented by numbers in the text. Style should be set to style{SIGCHI-Reference-Format}, this moment it is set to the style draft}
-
 \appendix

### 2016-05-29 paper/paper.tex
test

@@
 \def\emptyauthor{}
-\def\plainkeywords{Information Visualization; CHI; NBA}
+\def\plainkeywords{Information Visualization; CHI; NBA; D3.js}
 \def\plaingeneralterms{}
@@
     including standings, team statistics, transfers and player statistics. The
-    d3.js framework was used to develop this visualization. 
+    d3.js framework was used to develop this visualization. The result can be 
+    consulted on tankske.github.io/InfoVisNBA/.
 \end{abstract}
@@
 The \nameref{sec:lessonslearned} section describes some interesting lessons we have learned from the project and in the \nameref{sec:futurework} section we discuss potential  improvements of the final visualization.
-We conclude in the
+We reflect on our work in the
 \nameref{sec:conclusion} section. Appendix~\ref{sec:terminology} contains definitions 
@@
 \section{Goal and Target Audience}\label{sec:goal} 
-The visualization's goal can be best summarized by the following sentence:
+The visualization's goal is to answer the following question:
 ``Why did a certain team win the NBA Championship?'' The visualization focuses
@@
 intuitive access to the data, the visualization allows users to draw their own
-conclusions. The visualization's target audience are lay persons. Specifically,
+conclusions. Specifically,
 basketball fans are the core of our target audience. It can be of particular
-interest with the recent surge in fantasy leagues, for which data analysis is
+interest with the recent surge in fantasy leagues, a type of online game where participants assemble imaginary or virtual teams of real players of a professional sport, for which data analysis is
 paramount to a player's success~\cite{fantasy,fantasyskill}.
@@
 Aggregated data for entire teams and for entire seasons is also available.
-Additionally, individual game statistics are also available, including playoff games.
+Additionally, individual game statistics are available, including playoff games.
 In our visualization we only use data from 1984 onwards.
@@
 Different methods can be used, going from a scatterplot to hexagonal charts where the shots are grouped into hexagonal regions, to a heatmap.
-This is often done for the the area in and around the three-point line, but it can als be done for the whole attack side of the field, as seen in the Vorped visualisation\cite{vorped}.
+This is often done for the area in and around the three-point line, but it can als be done for the whole attack side of the field, as seen in the Vorped visualisation\cite{vorped}.
 Because shots of the own side only happen rarely, it's not shown in shot charts.
@@
 playoffs bracket~\cite{tournamentladder}. This bracket structure has the
-advantage of being well known and as such, not requiring much explanation. It is
-flawed however. It duplicates much information by displaying teams multiple
-times.
+advantage of being well known and as such, does not require much explanation. 
+However, it duplicates much information by displaying teams repeatedly.
 
@@
 tournament as a sunburst, with the winning team taking the part of the circle
-closer to the center. This does not eliminate the problem of the recurring teams
-however.
+closer to the center. This does however not eliminate the problem of the recurring teams.
 
@@
 created~\cite{enclosedcircles}. This visualization eliminates the repeating
-occurrences of the same team. However, it wastes a lot of space.
+occurrences of the same team. However, it uses a lot of space.
 
@@
 Tan et al. created an adaption of the traditional tournament bracket to better
-support moden interaction\cite{adaptivitree}. The tool, AdaptiviTree, changes
+support modern interaction\cite{adaptivitree}. The tool, AdaptiviTree, changes
 the brackets shape and adds colored lines to show available information as the
@@
 
-An alternative to the traditional bracket structure can be found in the
-visualization of the race to the white house in 2012 made for the New York Times
+The play-off view in our visualization is an alternative to the traditional bracket structure inspired by the
+visualization of the race to the white house made in 2012 for the New York Times
 by Mike Bostock and Shan Carter\cite{whitehousepath}. They use a tree-like
@@
 fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score\cite{elorating} 
-for all teams an visualized them on line graphs.
+for all teams and visualized them on line graphs.
 
@@
 Pagno et al. provide a number of different visualizations\cite{starplots}. The
-most interesting one is the starplot. This plot several statistics for a team on
+most interesting one being the starplot. This plot several statistics for a team on
 an axis pointing radially outward from a common center. These points are then
@@
 with the keyboard for switching between the three different views is 
-supported as well.
+supported as well with standard browser navigation by pressing the space bar or
+the up-and-down arrows.
 
@@
 between the area of a bigger and the area of a smaller circle will be greater
-than the proportion of values the areas represent.
+than the proportion of values the areas represent to improve human perception.
 
@@
 This view is further reinforced by highlighting a team and its matches and 
-opponents when hovering over it. Additional team information is then also shown 
-in the context part.
+opponents when hovering over it. To emphasize the matches won and lost, the colors green
+and red where used. Additional team information is also shown in the context part
+when hoovering over a team.
 
@@
 
-As an alternative we tried the same view without the logo's. It turned out this 
-was inconvenient because you could not instantly see which team is represented 
+As an alternative we tried the same view without the logo's. This turned out to be 
+inconvenient for the user as one could not instantly see which team is represented 
 by which circle. Therefore we've chosen to only use the teams' logos.
 
-Our initial play-off view was too dense because of the connections between teams
-as well. Teams that played against each other were connected directly. When
+Our initial play-off view was too dense because teams that played against each other were connected directly. When
 searching for alternatives to reflect on our choice we discovered a similar
@@
 visualization gives a cleaner view of the competitors for a specific team compared 
-to our original sketch. Hence we decided to adapt our visualization. We added an 
-intermediate connection point between two teams to more clearly indicate how teams competed to 
-become the NBA champion.
+to our original sketch. Hence we decided to adapt our visualization.
 
@@
 We initially also considered a map view to show the teams and their regular
-competition ranking on a map of the USA. However, this view had no additional
-value compared to the play-off view except for the geographical position, which
-has no importance for our visualisation. Furthermore we've found a visualisation
+competition ranking on a map of the USA. However, the geographical position gave
+no additional information that would contribute to meaningful insights. Furthermore we've found a visualisation
 of the NBA with a map view~\cite{mapviewvisualization} similar to our idea, so
@@
 influenced in the selected season compared to the season before. The statistics
-view is showed in Figure~\ref{fig:statisticsview}. The view consists of 3
+view is shown in Figure~\ref{fig:statisticsview}. The view consists of 3
 coordinated parts:
@@
         selected team
-    \item Arrow section: this section illustrates why a statistic has changed
+    \item Arrow section: this section illustrates how transfers change statistics
         compared to the previous season
@@
 visualization. The team shown in the statistics view is highlighted on this
-smaller bubble section as well. Replicating this section also alleviates
-short-term memory issues. It prevents the need to scroll back and forth between
+smaller bubble section as well. Replicating this section alleviates
+short-term memory limitations. It prevents the need to scroll back and forth between
 the different views constantly.
@@
 transparant with a lightgrey border. They show the maximum, average and minimum SRS 
-value of the NBA in the selected season. The change for the chosen statistic is 
+value of the NBA teams in the selected season. The change for the chosen statistic is 
 represented by three arrows:
@@
 \end{itemize}
-The value of the change in statistic is shown next to the arrow. The number of 
-players responsible for the change is shown between brackets. Their are shown 
-jerseys next to the arrows as well. These are the jerseys of the two players that 
+The value of the change in statistic is shown next to the arrow. As well as the number of 
+players responsible for the change between brackets. Their are max. two
+jerseys next to the arrows as well. These are the jerseys of the players that 
 had the biggest contribution to the influence represented by that arrow.  
@@
 are the PER value, the age, the effective field goal percentage, the offensive rating 
-and the defensive rating. Below this, player information is shown when a user hovers
+and the defensive rating. Below this menu, player information is shown when a user hovers
 over a player's jersey.
@@
 By interacting with the visualization, previous seasons can be inspected to try 
-to find explanations or following seasons can be inspected to see if a team 
-responds in a specific way or what the consequences are. This view provides 
+and find explanations. The user could als look at the following seasons to see how a team 
+performed related to the evolution of a statistic. This view provides 
 a lot of additional information to users by allowing them to see correlations 
@@
 \subsection{The team view}
-The team view gives a user a detailed view on the strength of the team on the 
-different field positions. The team view is showed in Figure~\ref{fig:teamview}. 
+The team view gives a user a detailed view on the strength of the team at the 
+different field positions. The team view is shown in Figure~\ref{fig:teamview}. 
 The view consists of 3 coordinated parts:
 \begin{itemize}
-    \item Context section: the top part informs the user of the context of the
+    \item Context sect
+    ion: the top part informs the user of the context of the
         selected team
@@
 \subsubsection{Field section}
-This view shows each player of the team on its field positions. Players are
+This view shows each player of the team at its field positions. Players are
 represented by their jerseys. Jerseys aligned next to each other share the same
-position, which is indicated as well. The size of a jersey illustrates a
+position, which is indicated as well with a rectangle. The size of a jersey illustrates a
 players performance according to the chosen statistic. The possible statistics 
@@
 part of the visualization was not our main focus. Because of the lesser
-importance and because of time and resource constraints, it has not been fully
+importance and because of time and resource constraints, it has not been
 developed. A sketch including the small multiples view on the bottom right is
@@
 \begin{itemize}
-        \item A synchrounous call need to be combined with synchronous calls.
+        \item Asynchrounous calls had to be combined with synchronous calls.
             More specific, when a visitor uses the timeline to scroll through
@@
 At the beginning of this project we were able to design the 3 separate views
-quite quickly. Our main concern was, when integrating these views as a whole, would 
-we be able to see the relations and consequences we thought we would see and if 
-we could draw the conclusions we hoped for. The relation we thought would be visible 
+quite quickly. Our main concern was to integrate these views as a whole and to be 
+able to let a user see relations between the teams' performance and their statistics to draw
+conclusions of his/her own. The relation we thought would be visible 
 was that the total PER value of a team would influence the end positions in the NBA championship, both the league rank as the play-off rank. We expected that teams 
-that had a big increase in total PER value, due to a good transferpolicy in a 
-season, would benefit in one of the following seasons.
-It was not possible to draw this conclusion and according to us the reason for this is that a whole NBA team contains an average of 15 players.
-Some players may play a lot more than others and
-are therefore much more important to the team.
-This observation is why we could not draw our hoped conclusion, but at the same time we found it so interesting and we wondered if we could find an example to support this observation.
-The story is about
-the Chicago Bulls, the team of Michael Jordan who is seen as the greatest basketball 
+that had a big increase in total PER value, due to a good transfer policy during 
+season, would benefit from it in one of the following seasons.
+We were not able to conclude this from the data as some players may play a lot more than others and are therefore having a big influence on the team.
+Howeverm this led us to a story about
+the Chicago Bulls The team of Michael Jordan who is seen as the greatest basketball 
 player of all time.
@@
 Bulls again achieve to win the NBA chamionship three times in a row, again with a  team consisting of around five core players. Once again is Michael Jordan the star 
-player of the team, but also Scottie Pippen is still one of the best players of the team. So to become the champion, it seems like a team needs to have a default team 
-with around five core players and one or two star players.
+player of the team, but also Scottie Pippen is still one of the best players of the team.
+So to become the champion, it seems like a team needs to have a default team with around five core players and one or two star players.
+By having this story it was easier for us to evaluate our design We could focus more on relevant parts of the visualization and cut less relevant parts from the project eg the details player information.
 
@@
 The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
-Have there been better scores, or worse scores.
+Have there been better scores, or worse scores?
 Is this score
@@
 duplication of data and use of screen space. Coordinated views alleviate this
-need significantly as well, as each component is strongly supported by other
+need significantly, as each component is strongly supported by other
 components.
@@
 the right choice to reach our goal, it had some draw backs.
-If we would have not started with our custom view, we would have had more freedom in exploring multiple different visualizations and we would be able to evaluate different options with trial and error.
+If we did not started with our custom view, we would have had more freedom in exploring multiple different visualizations and we would be able to evaluate different options with trial and error.
 Instead much of the time developing was spent on the technical details for these custom visualizations.
@@
 As with most projects, the result is never 'finished', meaning that there are
-always extensions and iprovements possible but due to timing or resource constraints they haven't all been implemented. We suggest a number of features that might improve the visualization:
+always extensions and improvements possible but due to timing or resource constraints they haven't all been implemented. We suggest a number of features that might improve the visualization:
 \begin{itemize}
-    \item Smooth transitions between states could improve the user experience.
+    \item Smooth/animated transitions between states could improve the user experience.
         For example when changing year, a transformation of the arrow in the
@@
         more on comparisons. We could integrate a view where a visitor is able
-        to compare two teams in the same year, or in two different years.
+        to compare two teams in the same year, or in two different years by duplicating current views next to each other.
     \item We could focus more on the core team, i.e. five or six players who seems 
         to be the most important for the results of a team.
+    \item As mentioned before, we could elaborate individual player information and visualize their statistics relevant to their field position.
+
 \end{itemize}
@@
 \section{Conclusion}\label{sec:conclusion}
-We believe the visualization we have created can be of tremendous value for fans
+We believe the visualization can be of tremendous value for fans
 and fantasy basketball players. Personally we have gained several insights into
@@
 innovative play-offs representation combined with detailed team information in
-an integrated fashion has much value.
+an integrated fashion enables users to gain insights into why certain teams won the NBA Championship.
 

### 2016-05-30 paper/paper.tex
Last major update

@@
 \begin{abstract}
-    Visualization of sports data is very common in recent years as access to
-    data has increased. The abundance of data and audience allows for new and
-    innovative presentations. In this paper we describe an interactive
+    Visualization of sports data is very common in recent years as access to this
+    kind of data has increased. The abundance of data and audience allows for 
+    new and innovative presentations. In this paper we describe an interactive
     visualization of the NBA competition over the last 30 years. With three
@@
     including standings, team statistics, transfers and player statistics. The
-    d3.js framework was used to develop this visualization. The result can be 
-    consulted on tankske.github.io/InfoVisNBA/.
+    d3.js framework was used to develop this visualization.
 \end{abstract}
@@
 
-The National Basketball Association (NBA) is the most famous professional
-basketball league in the world\cite{nbawiki}. As in every other sports
-branch, a lot of data and statistics about players, teams, games and seasons
+The National Basketball Association (NBA)~\cite{nbawiki} is the most famous professional
+basketball league in the world. As in every other sports
+branch these days, a lot of data and statistics about players, teams, games and seasons
 are available, mainly in large tables. Despite the large amount of data
 available, gathering useful insights from this data can be surprisingly
-challenging. In our work we try to present this data in an innovative but still
-intuitive way, allowing the user to draw meaningful conclusions.
+challenging. In our visualization we try to present this data in an innovative but still
+intuitive way, allowing the user to overcome this challenge and to draw meaningful conclusions.
 
-In the \nameref{sec:goal} section we describe the goal of the visualization and the
+In this paper we elaborate on the creation and operation of the visualization.
+The section \nameref{sec:goal} describes the goal of the visualization and the
 target audience.
-In the \nameref{sec:data} section, we describe the data used, including its origins, advantages and limitations.
-The \nameref{sec:literature} section gives an overview of related literature and web resources, including related
+In the section \nameref{sec:data}, the data used, including its origins, advantages and limitations are described.
+The section \nameref{sec:literature} gives an overview of related literature and web resources, including related
 visualizations. This includes both visualizations of NBA or other sports data,
 as well as visualizations tackling an issue encountered during the development of
-our visualization. In the \nameref{sec:visualization} section, the visualization itself is described. We give an overview of the different stages of development
+our visualization. In the section \nameref{sec:visualization}, the visualization itself is described. We give an overview of the different stages of development
 of the visualization, as well as the major design decisions.
-The \nameref{sec:lessonslearned} section describes some interesting lessons we have learned from the project and in the \nameref{sec:futurework} section we discuss potential  improvements of the final visualization.
-We reflect on our work in the
-\nameref{sec:conclusion} section. Appendix~\ref{sec:terminology} contains definitions 
-for common basketball terms used in this paper.
+The section \nameref{sec:lessonslearned} describes some interesting lessons we have learned from the project and in the section \nameref{sec:futurework} we discuss potential  improvements of the final visualization.
+We reflect on our work in the section \nameref{sec:conclusion}.
+Appendix~\ref{sec:terminology} contains definitions for common basketball terms used in this paper.
+The visualization can
+be consulted on tankske.github.io/InfoVisNBA/.
 
@@
 ``Why did a certain team win the NBA Championship?'' The visualization focuses
-on exposing relations between team performance and their player roster over
-several years. It allows users to find explanations for major improvements and
+on exposing relations between team performance and their player selection over
+several seasons. It allows users to find explanations for major improvements and
 declines in team performance.  The visualization allows exploration of NBA data
 by lay persons.  More specifically, the visualization does not focus on premade
-explanations for phenomenons visible in the NBA data. By providing easy and
+explanations for phenomena visible in the NBA data. By providing easy and
 intuitive access to the data, the visualization allows users to draw their own
@@
 The data visualized is a subset of the data available on basketball statistics
-site basketball-reference\cite{basketball-reference}.
-It is a wide range of data including common basketball performance statistics, such as field goals, percentage of shots scored, shooting distance, minutes played per season, number of personal fouls made, salaries and much more .
+site basketball-reference~\cite{basketball-reference}.
+It is a wide range of data including common basketball performance statistics, such as field goals, percentage of shots scored, shooting distance, minutes played per season, number of personal fouls made, salaries and much more.
 All of this data is available as time-series.
@@
 In our visualization we only use data from 1984 onwards.
-This is done because of rule changes in the NBA which changed the number of teams competing and the
+This is done because of rule changes in the NBA, which changed the number of teams competing and the
 competition's structure. To simplify implementation of our visualization, only
@@
 team's roster and individual player statistics for each year, including the 
-PER~(Player Efficiency Rating)\cite{per}.
+Player Efficiency Rating (PER)~\cite{per}.
 
@@
 provided download capabilities. Most of the data was downloaded in csv format,
-while some tables had to be manually scraped. This process was automated using
+while some tables had to be scraped manually. This process was automated using
 Python scripts. The data was then combined in a preprocessing step. In this
-step, each team's playoff rankings were calculated based on the matches played
-during the playoffs, and the rest of the data was combined into json format. The
+step, each team's play-off rankings were calculated based on the matches played
+during the play-offs, and the rest of the data was combined into json format. The
 final preprocessing step combined all data into one json file. A model of the
-available data can be seen on figure~\ref{fig:data}.
+available data can be seen in Figure~\ref{fig:data}.
 
@@
 data. This is best exemplified by the workshop on sports data visualization at
-the IEEE Visualization Conference in 2013\cite{ieeevis}.
+the IEEE Visualization Conference in 2013~\cite{ieeevis}.
 
-There are a number of traditional basketball visualizations. One of the most
-popular ones is the shot chart.
-This visualizes the shots made by a player or team.
+There are a number of traditional basketball visualizations, with the shot chart 
+being one of the most popular.
+A shot chart  visualizes the shots made by a player or team.
 Different methods can be used, going from a scatterplot to hexagonal charts where the shots are grouped into hexagonal regions, to a heatmap.
-This is often done for the area in and around the three-point line, but it can als be done for the whole attack side of the field, as seen in the Vorped visualisation\cite{vorped}.
-Because shots of the own side only happen rarely, it's not shown in shot charts.
+This is often done for the area in and around the three-point line, but it can also be done for the whole attack side of the field, as seen in the visualization of Vorped~\cite{vorped}.
+Because shots of the own side only happen rarely, it's usually not shown in shot charts.
 Shot chart provides insight into the shooting ability of a
-player\cite{goldsberry,stephenchu}.
+player~\cite{goldsberry,stephenchu}.
 
-This view features heavily in Peter Beshai's Buckets\cite{peterbeshai}. This is
+This view features heavily in Peter Beshai's Buckets~\cite{peterbeshai}. This is
 an extensive visualization of NBA player data. It provides not only the player's
@@
 Visualizations focusing on basketball teams and their performance are
-significantly more difficult to find. For the playoffs, there is the traditional
-playoffs bracket~\cite{tournamentladder}. This bracket structure has the
+less common to find. For the play-offs, there is the traditional tournament bracket structure~\cite{tournamentladder}. This bracket structure has the
 advantage of being well known and as such, does not require much explanation. 
@@
 An interesting alternative is the sunburst tournament
-visualization\cite{sunburst}. As the name suggests, this visualizes the
+visualization~\cite{sunburst}. As the name suggests, this visualizes the
 tournament as a sunburst, with the winning team taking the part of the circle
-closer to the center. This does however not eliminate the problem of the recurring teams.
+closer to the center. This kind of visualization has the same drawback as the traditional bracket structure has.
 
-For the wimbledon tennis tournament, an enclosed circles visualization was
-created~\cite{enclosedcircles}. This visualization eliminates the repeating
-occurrences of the same team. However, it uses a lot of space.
+Peter Cook created the Player Bubbles visualization~\cite{enclosedcircles}, where the tennis tournament Wimbledon is visualized as enclosed circles.
+A player is represented by a bubble that encloses all the bubbles of players he has won of.
+This eliminates the need of repeating the same player twice, but on the other hand it uses a lot of space and it makes it hard to understand the structure of the tournament.
 
-England's Permier League results were visualized by Ami Sedghi for the
-Guardian\cite{premierleague}. This visualization shows the team's rankings as
+In the Guardian, the England's Premier League results were visualized by Ami 
+Sedghi~\cite{premierleague}. This visualization shows the team's rankings as
 line graphs and allows highlighting of individual teams.
 
-Tan et al. created an adaption of the traditional tournament bracket to better
-support modern interaction\cite{adaptivitree}. The tool, AdaptiviTree, changes
+Tan et al. created an adaptation of the traditional tournament bracket to better
+support modern interaction~\cite{adaptivitree}. The tool, AdaptiviTree, changes
 the brackets shape and adds colored lines to show available information as the
@@
 
-The play-off view in our visualization is an alternative to the traditional bracket structure inspired by the
-visualization of the race to the white house made in 2012 for the New York Times
-by Mike Bostock and Shan Carter\cite{whitehousepath}. They use a tree-like
+An alternative to the traditional bracket structure can be found in the
+visualization of the race to the white house in 2012 made for the New York Times
+by Mike Bostock and Shan Carter~\cite{whitehousepath}. They use a tree-like
 structure where each split represents a state being one by one candidate or the
 other. The interactivity is very important here as it allows users to highlight
-one possible path out of many.
+one possible path out of many. The play-off view in our visualization is inspired on this visualization as explained in the section \nameref{sec:visualization}.
 
 The entire history of the NBA and the best teams were analyzed by
-fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score\cite{elorating} 
+fivethirtyeight~\cite{fivethirtyeight}. They calculated an ELO-score~\cite{elorating} 
 for all teams and visualized them on line graphs.
 
-In the article written by Johannes Becker\cite{nbaempires}, the Simple Rating System 
+In the article written by Johannes Becker~\cite{nbaempires}, the Simple Rating System 
 score (SRS score) is used to compare teams over the last 40 seasons in the NBA. The 
 SRS score is represented by a color hue. For comparing quantative values, this is 
-not the most ideal characteristic to use as explained by John Mackinglay\cite{automatingdesign}.
-Nevertheless, some interesting findings are found and presented.
+not the most ideal characteristic to use as explained by Jock Mackinglay~\cite{automatingdesign}.
+Nevertheless, some interesting insights are found and presented.
 By giving a green stroke to the rectangle of the champion of a season,
@@
 
-Pagno et al. provide a number of different visualizations\cite{starplots}. The
-most interesting one being the starplot. This plot several statistics for a team on
-an axis pointing radially outward from a common center. These points are then
+Pagno et al.
+provide a number of different visualizations~\cite{starplots}, with the starplot being one of the most interesting.
+It plots several statistics for a team on an axis pointing radially outward from a common center.
+These points are then
 connected. It allows the user to get basic information about a team or player at
@@
 an overview, allows zooming and filtering and provides details on
-demand\cite{mantra,multipleviews,automatingdesign}. The play-off view provides
+demand~\cite{mantra,multipleviews}. The play-off view provides
 the full overview which allows filtering by team. The statistics view provides
-the ability to zoom into one specific team. There some details are provided
+the ability to zoom in on one specific team. There some details are provided
 instantly, while others can be procured by hovering over certain elements.
@@
 available to the user on a single webpage, where they are organized in a 
-vertical layout with fixed scrolling. Navigation with the keyboard is 
-supported as well. The user start on the play-off view and is able to 
-navigate to the statistics and team views by scrolling down or clicking 
-on teams. A different season can be selected with the left and right
+vertical layout with fixed scrolling. The user start on the play-off view. Navigation with the keyboard for switching between the three different views is 
+supported as well, with standard browser navigation by pressing the space bar or
+the up-and-down arrows. A different season can be selected with the left and right
 arrows or by clicking on the timeline. The timeline is a fixed part of the
-visualization. It is visible in the three different views. Navigation 
-with the keyboard for switching between the three different views is 
-supported as well with standard browser navigation by pressing the space bar or
-the up-and-down arrows.
+visualization. It is visible in the three different views.
 
@@
 was made to use a thematic background. This was done to improve the 
-attractiveness of the visualisation on the one hand and the users' memory of 
-it on the other hand\cite{aesthetics}.
+attractiveness of the visualization on the one hand and the users' memory of 
+it on the other hand~\cite{aesthetics}.
 
@@
 for a selected season, in general or for a selected team. The play-off view is 
-shown in figure~\ref{fig:playoffviewwithcontext}.The view consists of
+shown in Figure~\ref{fig:playoffviewnoteam}. The view consists of
 two parts: 
@@
 team's region, respectively the western or eastern conference. The context
-section also contains a link to the ``About'' page where the user can get an
-explanation of the visualization with its different views, can get some
-background information about the NBA and the terminolgy used, and can check the
-source of the data used in the visualization. If the user hovers over or selects 
-a team in the bubble section, the context section shows more information about that 
-team, as shown in Figure~\ref{fig:playoffviewteamselected}. The information shown 
-includes the SRS score, the end position in the play-offs and the end position in 
-the regular competition.
+section also contains a link to the ``About'' page where the user can get an explanation of the visualization with its different views, some background information about the NBA and the terminology used, and the link to the source of the data used in the visualization.
+If the user hovers over or selects a team in the bubble section, the context section shows more information about that team, as shown in Figure~\ref{fig:playoffviewteam}.
+The information shown includes the SRS score, the end position in the play-offs and the end position in the regular competition.
 
@@
 This section presents the roster and end positions of the play-offs for a
-selected season. Every team is respresented by a circle with the logo of the
-team inside it. The size of the circle represents the SRS score using an own
+selected season. Every team is represented by a circle with the team's logo inside it. The size of the circle represents the SRS score using an own
 interpretation of perceptual scaling because the (Flannery) Appearance
@@
 An exponential function with exponent 1.5 was used. This way the proportion
-between the area of a bigger and the area of a smaller circle will be greater
+between the area of a bigger and a smaller circle will be greater
 than the proportion of values the areas represent to improve human perception.
@@
 The curved lines between teams represent the fact that those teams have 
-encountered each other in the playoffs. Each row represents an end position 
-in the play-offs. The higher a team is, the further it has come.  From the 
-top to the bottom the rows represent the winner of the final (rank 1), loser 
+encountered each other in the play-offs. Each row represents an end position 
+in the play-offs. The higher a team stands, the further it has come.  From the 
+top to the bottom, the rows represent the winner of the final (rank 1), loser 
 of the final (rank 2), losers of the semi-finals (rank 3), losers of the 
-quarter-finals (rank 4), losers of the eighth finals (rank 5). The last row 
-represents the teams that did not make it to the play-offs (rank 6).
+quarter-finals (rank 4), losers of the first round in the play-offs (rank 5). The last row represents the teams that did not make it to the play-offs (rank 6).
 
@@
 higher level, the team needs to have played and won a series of matches against 
-one other team at every level below them~\ref{fig:playoffviewteamselected}. 
-This view is further reinforced by highlighting a team and its matches and 
-opponents when hovering over it. To emphasize the matches won and lost, the colors green
-and red where used. Additional team information is also shown in the context part
-when hoovering over a team.
+one other team at every level below them, which can be seen in Figure~\ref{fig:playoffviewnoteam}.
+This view is further reinforced by highlighting a team, its matches and opponents when hovering over it.
+To emphasize the matches won and lost, the colors green and red where used.
+Additional team information is also shown in the context part when hovering over a team.
 
@@
 Our initial design used the team circles' lightness to indicate their final
-ranking in their region. The hue was used to show the league they were part of.
-According to the Ranking of Perceptual Tasks~\cite{automatingdesign} this is a
-good property for ordered values like the regular competition ranking. But the
-combination of the logo's of the teams with the colors resulted in an overload
-of information and made it more difficult to compare the end ranking of
-different teams.
+ranking in their region. The hue was red or blue indicating the teams' region.
+According to the Ranking of Perceptual Tasks by Jock Mackinlay~\cite{automatingdesign} this is a good property for ordered values like the end positions in the regular competition ranking.
+But the combination of the teams' logo's with the different hues and ligthnesses resulted in an overload of information and made it more difficult to compare different teams based on their end ranking.
+Our initial design is shown in Figure~\ref{fig:firstdesignplayoffview}
 
-As an alternative we tried the same view without the logo's. This turned out to be 
-inconvenient for the user as one could not instantly see which team is represented 
-by which circle. Therefore we've chosen to only use the teams' logos.
+As an alternative we tried the same view without the logo's, , as shown in Figure~\ref{fig:playoffviewfirstimplemantation}.
+This turned out to be inconvenient for the user as one could not instantly see which team is represented by which circle.
+Recognizing the different teams is a fundamental part of our visualization, while the comparison of end ranking of different teams would not have great importance.
+Therefore we've chosen to only use the teams' logos.
 
-Our initial play-off view was too dense because teams that played against each other were connected directly. When
-searching for alternatives to reflect on our choice we discovered a similar
-visualization in a totally different context\cite{whitehousepath} as explained in 
-the section \ref{sec:literature}. With the idea of this visualization in mind, our 
-visualization gives a cleaner view of the competitors for a specific team compared 
-to our original sketch. Hence we decided to adapt our visualization.
+Our initial play-off view was too dense because teams that played against each other were connected directly.
+This can also be seen in Figure~\ref{fig:playoffviewfirstimplemantation}.
+When searching for alternatives to reflect on our choice we discovered a similar visualization by Mike Bostock and Shan Carter~\cite{whitehousepath} in a totally different context as explained in the section \nameref{sec:literature}.
+With the idea of this visualization in mind, our visualization gives a cleaner view of the competitors for a specific team compared to our original sketch.
+Hence we decided to adapt our visualization.
 
@@
 competition ranking on a map of the USA. However, the geographical position gave
-no additional information that would contribute to meaningful insights. Furthermore we've found a visualisation
-of the NBA with a map view~\cite{mapviewvisualization} similar to our idea, so
-we decided to not implement this map view.
+no additional information that would contribute to meaningful insights.
+Furthermore we've found a visualization of the NBA with a map view made by Leeman Chiu~\cite{mapviewvisualization} similar to our idea, so we decided to not implement this map view.
 
@@
 \subsubsection{Context section}
-The context section is partly identical to that of the play-offs view. The
-team's information is also displayed in the top left. In the top right, the
+The context section is partly identical to that of the play-off's view. The
+team's information is again displayed in the top left. In the top right, the
 bubble section of the play-offs view is displayed in a reduced, minimalistic
 form. This view is repeated to create a sense of continuity between the
-different views and provide context of the current position in the
+different views and to provide context of the current position in the
 visualization. The team shown in the statistics view is highlighted on this
-smaller bubble section as well. Replicating this section alleviates
-short-term memory limitations. It prevents the need to scroll back and forth between
-the different views constantly.
+smaller bubble section the same way as in the large play-off view..
+Furthermore, replicating this section alleviates short-term memory limitations.
+It prevents the need to scroll back and forth between the different views constantly.
 
 \subsubsection{Arrow section}
-This section illustrates a change for a chosen statistic for the selected team in 
-the selected season compared to the previous season. This gives a user more 
+This section illustrates how transfers change a chosen statistic for the selected team in the selected season compared to the previous season. This gives a user more 
 insight into why a statistic changed over time and how this change has affected 
-team performance. The selected team is represented by a circle in the same way as 
-explained in the play-off view. To give the user a better interpretation of the SRS 
-value of the circle, three other reference circles are drawn. These are 
-transparant with a lightgrey border. They show the maximum, average and minimum SRS 
-value of the NBA teams in the selected season. The change for the chosen statistic is 
-represented by three arrows:
+team performance.
+The selected team is represented by a circle in the same way as in the play-off view.
+To give the user a better interpretation of the SRS value of the team, three other reference circles are drawn.
+These are transparent with a lightgrey border and show the maximum, average and minimum SRS value of the NBA teams in the selected season.
+The change for the chosen statistic is represented by three arrows:
 \begin{itemize}
     \item Arrow on the left pointing towards the circle: indicating the
-        influence of players who joined the team.
+        influence of players who joined the team
     \item Arrow on the right of the circle: indicating how the team internally
@@
 \end{itemize}
-The value of the change in statistic is shown next to the arrow. As well as the number of 
-players responsible for the change between brackets. Their are max. two
+The value of the change in statistic is shown next to the arrow, as well as the number of 
+players responsible for the change, which is shown between brackets. There are maximum two
 jerseys next to the arrows as well. These are the jerseys of the players that 
-had the biggest contribution to the influence represented by that arrow.  
-On the left, there is a possibility to chose a statistic. The possibilities 
-are the PER value, the age, the effective field goal percentage, the offensive rating 
-and the defensive rating. Below this menu, player information is shown when a user hovers
-over a player's jersey.
+had the biggest contribution to the change represented by their corresponding arrow.  
+On the left, there is the possibility to choose a statistic.
+The different possibilities are the PER value, the age, the effective field goal percentage, the offensive rating and the defensive rating.
+Below this menu, player information is shown when a user hovers over a player's jersey.
 
@@
 team has performed over the years. The different statistics shown are the 
-total PER of the whole team, the league rank, the playoff rank, the average age 
+total PER of the whole team, the league rank, the play-off rank, the average age 
 of the team, the audience, the SRS-value of the team, the offensive rating and 
@@
 in the context section and why the statistic changes in the arrow section. 
-These charts enable the user to identify peaks or drops in a certain statistic. 
+These charts enable the user to identify peaks or drops in a certain team statistic. 
 By interacting with the visualization, previous seasons can be inspected to try 
-and find explanations. The user could als look at the following seasons to see how a team 
+and find explanations. The user could also look at the following seasons to see how a team 
 performed related to the evolution of a statistic. This view provides 
@@
 This view has not changed much compared to the initial design. The only difference 
-was the position of the arrow and jerseys for the players who stayed, which were 
-originally in the team circle, and so the logo was not showed. 
+is the position of the arrow and jerseys for the players who stayed, which was originally located in the team circle, and so the team's logo was not shown. 
 The design of the arrows did not change. The incoming and outgoing 
-arrows are maybe more intuitevly than the arrow for the stayed players, but we did 
+arrows are maybe more intuitively than the arrow for the stayed players, but we did 
 not find any alternatives we could use.
@@
 \begin{itemize}
-    \item Context sect
-    ion: the top part informs the user of the context of the
+    \item Context section: the top part informs the user of the context of the
         selected team
-    \item Field section: this section illustrates why a statistic has changed
-        compared to the previous season
+    \item Field section: this section illustrates the strength of the whole team on
+        the different field positions
     \item Selected statistics section: team statistics and how they evolve over
@@
 \subsubsection{Field section}
-This view shows each player of the team at its field positions. Players are
+This view shows each player of the team at its field position. Players are
 represented by their jerseys. Jerseys aligned next to each other share the same
-position, which is indicated as well with a rectangle. The size of a jersey illustrates a
-players performance according to the chosen statistic. The possible statistics 
+position, which is indicated with a rectangle. The size of a jersey illustrates a
+player's performance according to the chosen statistic. The possible statistics 
 are shown on the left and are the same as those in the arrow section of the 
@@
 A user can then scroll through time to see how field positions evolve and what 
-the influence is over time.
+the influence of this field changes are over time.
 
@@
 developed. A sketch including the small multiples view on the bottom right is
-shown on Figure~\ref{fig:smallmultsketch}.
+shown in Figure~\ref{fig:smallmultsktech}.
 
@@
 structure the site with multiple divs as containers. To each container we
-allocate a svg with a specific visualization.
+allocate an svg with a specific visualization.
 %\todo{sketch of how the page is structured here}
@@
 \begin{itemize}
-        \item Asynchrounous calls had to be combined with synchronous calls.
+        \item Asynchronous calls had to be combined with synchronous calls.
             More specific, when a visitor uses the timeline to scroll through
-            time, the whiping of previous visualizations was not synced with the
-            creation of the visualisation. This resulted in whiping sections
+            time, the wiping of previous visualizations was not synced with the
+            creation of the new visualization. This resulted in wiping sections
             when a visualization was not created yet. Hence the view resulted in
@@
 was that the total PER value of a team would influence the end positions in the NBA championship, both the league rank as the play-off rank. We expected that teams 
-that had a big increase in total PER value, due to a good transfer policy during 
-season, would benefit from it in one of the following seasons.
-We were not able to conclude this from the data as some players may play a lot more than others and are therefore having a big influence on the team.
-Howeverm this led us to a story about
-the Chicago Bulls The team of Michael Jordan who is seen as the greatest basketball 
-player of all time.
+that had a big increase in total PER value, due to a good transfer policy during a season, would benefit from it in one of the following seasons.
+We were not able to conclude this from the data as some players may play a lot more than others and are therefore of much bigger importance for the team.
+This interesting observation led us to a story about the Chicago Bulls, the team of Michael Jordan, who is considered to be the greatest basketball player of all time.
 In 1984, Jordan arrives at the Chicago Bulls and in his first season he is already the top player of his team, as seen by the size of his jersey for the PER statistic.
-In the season 1990-1991 and the two following seasons, the
-Chicago Bulls win the NBA championship. The ``three-peat'' as it is called was done 
-with almost the same core team. Next to Michael Jordan, there were around five 
-players that were part of the team and were roughly the best on their position in 
-the team, for example Scottie Pippen. We can see that in the seasons before this 
-``three-peat'' this team was formed and already played with each other. After that 
-last season form those three consecutive wins, Michael Jordan retires and the team 
-is not able anymore to achieve the same.
+In the season 1990-1991 and the two following seasons, the Chicago Bulls win the NBA championship.
+The ``three-peat'' as it is called, was done with almost the same core team.
+Next to Michael Jordan, there were around five players that were part of the team and were roughly the best on their position in the team, for example Scottie Pippen.
+We can see that in the seasons before this ``three-peat'' this team was formed and already played with each other.
+After that last season from those three consecutive wins, Michael Jordan retires and the team failed in winning the NBA championship again.
 In the middle of the season 1994-1995, Michael Jordan came out of retirement.
 What follows is a similar story to the previous one.
-In the season 1995-1996 and the two following seasons, the Chicago
-Bulls again achieve to win the NBA chamionship three times in a row, again with a  team consisting of around five core players. Once again is Michael Jordan the star 
-player of the team, but also Scottie Pippen is still one of the best players of the team.
-So to become the champion, it seems like a team needs to have a default team with around five core players and one or two star players.
-By having this story it was easier for us to evaluate our design We could focus more on relevant parts of the visualization and cut less relevant parts from the project eg the details player information.
+In the season 1995-1996 and the two following seasons, the Chicago Bulls again achieve to win the NBA championship three times in a row, again with a  team consisting of around five core players.
+Once again is Michael Jordan the star player of the team, but also Scottie Pippen is still one of the best players of the team.
+So it seems that to become the champion, a team needs to have a default team with around five core players from which one or two are star players.
+By having this story it was easier for us to evaluate our design.
+We could focus more on important parts of the visualization and cut less relevant parts from the project, e.g.
+the detailed player information.
 
@@
 clear and informative visualization on itself.
-During the project, we started noticing that most components in a visualization lack references and hence are not informative enough.
-Eg.
-each small multiple of a statisctic in the statistics view on itself gives sufficient information to stand on itselfs.
+During the project, we started noticing that most components in a visualization lack references and hence are not informative enough, e.g.
+each small multiple of a statistic in the statistics view gives sufficient information to stand on itself.
 The bar in the small multiple indicates the score in the selected season and enables the user to evaluate that score over time.
@@
 Is this score
-part of an upward movement during years? This should be weighed against the
-duplication of data and use of screen space. Coordinated views alleviate this
-need significantly, as each component is strongly supported by other
-components.
+part of an upward movement during years? This should be weighted against the
+duplication of data and the use of screen space. Coordinated views alleviate this
+need significantly, as each component is strongly supported by other components.
 
@@
 Before exploring what is out there in the d3.js world, we made our own sketches.
-This is how we came up with the 3 custom views that make up our visualization.
+This is how we came up with the three custom views that make up our visualization.
 When comparing our design with alternatives we could not find something that
-satisfied our needs. Nonetheless, they gave us inspiration to finetune our
-designs. Although we are satisfied with our result and we are convinced it was
-the right choice to reach our goal, it had some draw backs.
+satisfied our needs. Nonetheless, they gave us inspiration to fine-tune our
+design. Although we are satisfied with our result and we are convinced it was
+the right choice to reach our goal, it had some drawbacks.
 If we did not started with our custom view, we would have had more freedom in exploring multiple different visualizations and we would be able to evaluate different options with trial and error.
-Instead much of the time developing was spent on the technical details for these custom visualizations.
+Instead much of the developing time was spent on the technical details for these custom visualizations.
 This was especially true for the bubble component of the play-off view.
-No out of the box solution for this could be found so everything from the determination of the positions of the bubbles to how to connect them had to be developed by us.
+No out of the box solution for this could be found so everything from the determination of the positions of the bubbles to how to connect them we had to develop ourselves.
 
 \section{Future work}\label{sec:futurework}
-As with most projects, the result is never 'finished', meaning that there are
-always extensions and improvements possible but due to timing or resource constraints they haven't all been implemented. We suggest a number of features that might improve the visualization:
+As with most projects, the result is never `finished', meaning that there are
+always extensions and improvements possible. We suggest a number of features that might improve the visualization:
 \begin{itemize}
     \item Smooth/animated transitions between states could improve the user experience.
-        For example when changing year, a transformation of the arrow in the
-        statistics view would help the user see a difference between years.  At
-        the moment the arrows change at the blink of an eye, making it difficult
-        for us humans with a short term memory to make the comparison. This
-        would also have been very useful in the transitions between different
+        For example when changing seasons, an animated transition of the arrow in 
+        the statistics view would help the user see the difference between the 
+        seasons. At the moment the arrows change at the blink of an eye, making it  
+        difficult for us, humans with a short term memory, to make the comparison. 
+        This would also have been very useful in the transitions between different 
         years on the bubble view.
-    \item To further support our short term memory we could also 'save' the
-        previous state. Eg. in the field view we could visualize ghost jerseys
+    \item To further support our short term memory we could also `save' the
+        previous state. E.g. in the field view we could visualize ghost jerseys
         to indicate which players left the team or position and highlight
@@
     \item At the moment no attention has been payed to the process, provenance
-        and history of the user's exploration session. A user thus needs to
+        and history of the user's exploration session. A user needs to
         memorize his action to be able to share or reproduce his actions. This
-        could be improved by adding URL parameters. This way a user at least can
+        could be improved by adding URL parameters. This way a user can
         share or save a state of an analysis. One step further is to provide a
-        play-feature. The visitor could them share his exploration and replay
-        it. A simple implemention would ask the visitor for a start date and an
-        end date and would then 'play' the changes of a particular team and
-        characteristic over time.
-    \item We could support users further in looking for patterns by focussing
+        play-feature. Then the visitor could share his exploration and replay
+        it. A simple implementation would ask the visitor for a start date and an
+        end date and would then `play' the changes of the selected team and 
+        characteristic.
+    \item We could support users further in looking for patterns by focusing
         more on comparisons. We could integrate a view where a visitor is able
-        to compare two teams in the same year, or in two different years by duplicating current views next to each other.
+        to compare two teams in the same year or in two different years, for example   
+        by duplicating current views next to each other.
     \item We could focus more on the core team, i.e. five or six players who seems 
         to be the most important for the results of a team.
-    \item As mentioned before, we could elaborate individual player information and visualize their statistics relevant to their field position.
+    \item As mentioned before, we could elaborate more on individual player 
+        information and visualize their statistics relevant to their field position.
 
@@
 \section{Conclusion}\label{sec:conclusion}
-We believe the visualization can be of tremendous value for fans
+We believe the visualization can be of tremendous value for fans of the NBA 
 and fantasy basketball players. Personally we have gained several insights into
@@
 individual players in a basketball team. We were able to find strong
-correlations between having two to four very strong players and the team's
-performance. It seems having one star player is not enough. On the other hand,
-having more than 5 good players no longer has much influence either.
+correlations between having two to five very strong players and the team's
+performance. It seems that just having one star player is not enough. On the other hand, having more than 5 good players no longer has much influence either.
 
@@
 
-\paragraph{NBA} National Basketball Association - The largest, most well-known
-basketball of the USA.
-
 \paragraph{PER} Player Efficiency Rating - An all-in-one basketball rating,
-boiling down all of a player's contributions into one number\cite{per}.
+boiling down all of a player's contributions into one number~\cite{per}.
 
@@
 point differential and strength of schedule. The rating is denominated in points
-above/below average, where zero is average\cite{srs}.
+above/below average, where zero is average~\cite{srs}.
 
-\paragraph{Playoffs} The National Basketball Association (NBA) playoffs are a
+\paragraph{Play-offs} The National Basketball Association (NBA) play-offs are a
 best-of-seven elimination tournament among 16 teams in the Eastern Conference
-and Western Conference (called divisions, pre-1970), ultimately deciding the
-winner of the NBA Finals\cite{playoffs}.
+and Western Conference, ultimately deciding the winner of the NBA Finals~\cite{playoffs}.
 
 \paragraph{Jersey} The shirt worn by basketball players.
+
+\paragraph{Three-point line} The line that separates the two-point area and the three-point area.
+
+%\todo{PER value, the age, the effective field goal percentage, the offensive rating and the defensive rating}
 

### 2016-05-30 paper/paper.tex
correction reference figure

@@
   \includegraphics[width=0.9\columnwidth]{figures/playoffviewfirstimplemantation}
-  \caption{First implementation of the play-off view.}~\label{fig:firstimplementationplayoffview}
+  \caption{First implementation of the play-off view.}~	
+  \label{fig:playoffviewfirstimplemantation}
 \end{figure}
@@
 But the combination of the teams' logo's with the different hues and ligthnesses resulted in an overload of information and made it more difficult to compare different teams based on their end ranking.
-Our initial design is shown in Figure~\ref{fig:firstdesignplayoffview}
+Our initial design is shown in Figure~\ref{fig:firstdesignplayoffview}.
 
-As an alternative we tried the same view without the logo's, , as shown in Figure~\ref{fig:playoffviewfirstimplemantation}.
+As an alternative we tried the same view without the logo's, as shown in Figure~\ref{fig:playoffviewfirstimplemantation}.
 This turned out to be inconvenient for the user as one could not instantly see which team is represented by which circle.
